\book[The Gospel According to Saint Matthew]{Matthew}



\chapter
\verse The book of the generation of Jesus Christ, the son of David, the son of Abraham.

\verse Abraham begat Isaac; and Isaac begat Jacob; and Jacob begat Judas and his brethren; \verse And Judas begat Phares and Zara of Thamar; and Phares begat Esrom; and Esrom begat Aram; \verse And Aram begat Aminadab; and Aminadab begat Naasson; and Naasson begat Salmon; \verse And Salmon begat Booz of Rachab; and Booz begat Obed of Ruth; and Obed begat Jesse; \verse And Jesse begat David the king; and David the king begat Solomon of her that had been the wife of Urias; \verse And Solomon begat Roboam; and Roboam begat Abia; and Abia begat Asa; \verse And Asa begat Josaphat; and Josaphat begat Joram; and Joram begat Ozias; \verse And Ozias begat Joatham; and Joatham begat Achaz; and Achaz begat Ezekias; \verse And Ezekias begat Manasses; and Manasses begat Amon; and Amon begat Josias; \verse And Josias begat Jechonias and his brethren, about the time they were carried away to Babylon: \verse And after they were brought to Babylon, Jechonias begat Salathiel; and Salathiel begat Zorobabel; \verse And Zorobabel begat Abiud; and Abiud begat Eliakim; and Eliakim begat Azor; \verse And Azor begat Sadoc; and Sadoc begat Achim; and Achim begat Eliud; \verse And Eliud begat Eleazar; and Eleazar begat Matthan; and Matthan begat Jacob; \verse And Jacob begat Joseph the husband of Mary, of whom was born Jesus, who is called Christ.

\verse So all the generations from Abraham to David are fourteen generations; and from David until the carrying away into Babylon are fourteen generations; and from the carrying away into Babylon unto Christ are fourteen generations.

\verse Now the birth of Jesus Christ was on this wise: When as his mother Mary was espoused to Joseph, before they came together, she was found with child of the Holy Ghost.

\verse Then Joseph her husband, being a just man, and not willing to make her a publick example, was minded to put her away privily.

\verse But while he thought on these things, behold, the angel of the \LORD appeared unto him in a dream, saying, Joseph, thou son of David, fear not to take unto thee Mary thy wife: for that which is conceived in her is of the Holy Ghost.

\verse And she shall bring forth a son, and thou shalt call his name JESUS: for he shall save his people from their sins.

\verse Now all this was done, that it might be fulfilled which was spoken of the Lord by the prophet, saying, \verse Behold, a virgin shall be with child, and shall bring forth a son, and they shall call his name Emmanuel, which being interpreted is, God with us.

\verse Then Joseph being raised from sleep did as the angel of the Lord had bidden him, and took unto him his wife: \verse And knew her not till she had brought forth her firstborn son: and he called his name JESUS.


\chapter
\verse Now when Jesus was born in Bethlehem of Judaea in the days of Herod the king, behold, there came wise men from the east to Jerusalem, \verse Saying, Where is he that is born King of the Jews? for we have seen his star in the east, and are come to worship him.

\verse When Herod the king had heard these things, he was troubled, and all Jerusalem with him.

\verse And when he had gathered all the chief priests and scribes of the people together, he demanded of them where Christ should be born.

\verse And they said unto him, In Bethlehem of Judaea: for thus it is written by the prophet, \verse And thou Bethlehem, in the land of Juda, art not the least among the princes of Juda: for out of thee shall come a Governor, that shall rule my people Israel.

\verse Then Herod, when he had privily called the wise men, enquired of them diligently what time the star appeared.

\verse And he sent them to Bethlehem, and said, Go and search diligently for the young child; and when ye have found him, bring me word again, that I may come and worship him also.

\verse When they had heard the king, they departed; and, lo, the star, which they saw in the east, went before them, till it came and stood over where the young child was.

\verse When they saw the star, they rejoiced with exceeding great joy.

\verse And when they were come into the house, they saw the young child with Mary his mother, and fell down, and worshipped him: and when they had opened their treasures, they presented unto him gifts; gold, and frankincense and myrrh.

\verse And being warned of God in a dream that they should not return to Herod, they departed into their own country another way.

\verse And when they were departed, behold, the angel of the Lord appeareth to Joseph in a dream, saying, Arise, and take the young child and his mother, and flee into Egypt, and be thou there until I bring thee word: for Herod will seek the young child to destroy him.

\verse When he arose, he took the young child and his mother by night, and departed into Egypt: \verse And was there until the death of Herod: that it might be fulfilled which was spoken of the Lord by the prophet, saying, Out of Egypt have I called my son.

\verse Then Herod, when he saw that he was mocked of the wise men, was exceeding wroth, and sent forth, and slew all the children that were in Bethlehem, and in all the coasts thereof, from two years old and under, according to the time which he had diligently enquired of the wise men.

\verse Then was fulfilled that which was spoken by Jeremy the prophet, saying, \verse In Rama was there a voice heard, lamentation, and weeping, and great mourning, Rachel weeping for her children, and would not be comforted, because they are not.

\verse But when Herod was dead, behold, an angel of the Lord appeareth in a dream to Joseph in Egypt, \verse Saying, Arise, and take the young child and his mother, and go into the land of Israel: for they are dead which sought the young child's life.

\verse And he arose, and took the young child and his mother, and came into the land of Israel.

\verse But when he heard that Archelaus did reign in Judaea in the room of his father Herod, he was afraid to go thither: notwithstanding, being warned of God in a dream, he turned aside into the parts of Galilee: \verse And he came and dwelt in a city called Nazareth: that it might be fulfilled which was spoken by the prophets, He shall be called a Nazarene.


\chapter
\verse In those days came John the Baptist, preaching in the wilderness of Judaea, \verse And saying, Repent ye: for the kingdom of heaven is at hand.

\verse For this is he that was spoken of by the prophet Esaias, saying, The voice of one crying in the wilderness, Prepare ye the way of the Lord, make his paths straight.

\verse And the same John had his raiment of camel's hair, and a leathern girdle about his loins; and his meat was locusts and wild honey.

\verse Then went out to him Jerusalem, and all Judaea, and all the region round about Jordan, \verse And were baptized of him in Jordan, confessing their sins.

\verse But when he saw many of the Pharisees and Sadducees come to his baptism, he said unto them, O generation of vipers, who hath warned you to flee from the wrath to come?  \verse Bring forth therefore fruits meet for repentance: \verse And think not to say within yourselves, We have Abraham to our father: for I say unto you, that God is able of these stones to raise up children unto Abraham.

\verse And now also the axe is laid unto the root of the trees: therefore every tree which bringeth not forth good fruit is hewn down, and cast into the fire.

\verse I indeed baptize you with water unto repentance. but he that cometh after me is mightier than I, whose shoes I am not worthy to bear: he shall baptize you with the Holy Ghost, and with fire: \verse Whose fan is in his hand, and he will throughly purge his floor, and gather his wheat into the garner; but he will burn up the chaff with unquenchable fire.

\verse Then cometh Jesus from Galilee to Jordan unto John, to be baptized of him.

\verse But John forbad him, saying, I have need to be baptized of thee, and comest thou to me?  \verse And Jesus answering said unto him, Suffer it to be so now: for thus it becometh us to fulfil all righteousness.  Then he suffered him.

\verse And Jesus, when he was baptized, went up straightway out of the water: and, lo, the heavens were opened unto him, and he saw the Spirit of God descending like a dove, and lighting upon him: \verse And lo a voice from heaven, saying, This is my beloved Son, in whom I am well pleased.


\chapter
\verse Then was Jesus led up of the spirit into the wilderness to be tempted of the devil.

\verse And when he had fasted forty days and forty nights, he was afterward an hungred.

\verse And when the tempter came to him, he said, If thou be the Son of God, command that these stones be made bread.

\verse But he answered and said, It is written, Man shall not live by bread alone, but by every word that proceedeth out of the mouth of God.

\verse Then the devil taketh him up into the holy city, and setteth him on a pinnacle of the temple, \verse And saith unto him, If thou be the Son of God, cast thyself down: for it is written, He shall give his angels charge concerning thee: and in their hands they shall bear thee up, lest at any time thou dash thy foot against a stone.

\verse Jesus said unto him, It is written again, Thou shalt not tempt the Lord thy God.

\verse Again, the devil taketh him up into an exceeding high mountain, and sheweth him all the kingdoms of the world, and the glory of them; \verse And saith unto him, All these things will I give thee, if thou wilt fall down and worship me.

\verse Then saith Jesus unto him, Get thee hence, Satan: for it is written, Thou shalt worship the Lord thy God, and him only shalt thou serve.

\verse Then the devil leaveth him, and, behold, angels came and ministered unto him.

\verse Now when Jesus had heard that John was cast into prison, he departed into Galilee; \verse And leaving Nazareth, he came and dwelt in Capernaum, which is upon the sea coast, in the borders of Zabulon and Nephthalim: \verse That it might be fulfilled which was spoken by Esaias the prophet, saying, \verse The land of Zabulon, and the land of Nephthalim, by the way of the sea, beyond Jordan, Galilee of the Gentiles; \verse The people which sat in darkness saw great light; and to them which sat in the region and shadow of death light is sprung up.

\verse From that time Jesus began to preach, and to say, Repent: for the kingdom of heaven is at hand.

\verse And Jesus, walking by the sea of Galilee, saw two brethren, Simon called Peter, and Andrew his brother, casting a net into the sea: for they were fishers.

\verse And he saith unto them, Follow me, and I will make you fishers of men.

\verse And they straightway left their nets, and followed him.

\verse And going on from thence, he saw other two brethren, James the son of Zebedee, and John his brother, in a ship with Zebedee their father, mending their nets; and he called them.

\verse And they immediately left the ship and their father, and followed him.

\verse And Jesus went about all Galilee, teaching in their synagogues, and preaching the gospel of the kingdom, and healing all manner of sickness and all manner of disease among the people.

\verse And his fame went throughout all Syria: and they brought unto him all sick people that were taken with divers diseases and torments, and those which were possessed with devils, and those which were lunatick, and those that had the palsy; and he healed them.

\verse And there followed him great multitudes of people from Galilee, and from Decapolis, and from Jerusalem, and from Judaea, and from beyond Jordan.


\chapter
\verse And seeing the multitudes, he went up into a mountain: and when he was set, his disciples came unto him: \verse And he opened his mouth, and taught them, saying, \verse Blessed are the poor in spirit: for theirs is the kingdom of heaven.

\verse Blessed are they that mourn: for they shall be comforted.

\verse Blessed are the meek: for they shall inherit the earth.

\verse Blessed are they which do hunger and thirst after righteousness: for they shall be filled.

\verse Blessed are the merciful: for they shall obtain mercy.

\verse Blessed are the pure in heart: for they shall see God.

\verse Blessed are the peacemakers: for they shall be called the children of God.

\verse Blessed are they which are persecuted for righteousness' sake: for theirs is the kingdom of heaven.

\verse Blessed are ye, when men shall revile you, and persecute you, and shall say all manner of evil against you falsely, for my sake.

\verse Rejoice, and be exceeding glad: for great is your reward in heaven: for so persecuted they the prophets which were before you.

\verse Ye are the salt of the earth: but if the salt have lost his savour, wherewith shall it be salted? it is thenceforth good for nothing, but to be cast out, and to be trodden under foot of men.

\verse Ye are the light of the world. A city that is set on an hill cannot be hid.

\verse Neither do men light a candle, and put it under a bushel, but on a candlestick; and it giveth light unto all that are in the house.

\verse Let your light so shine before men, that they may see your good works, and glorify your Father which is in heaven.

\verse Think not that I am come to destroy the law, or the prophets: I am not come to destroy, but to fulfil.

\verse For verily I say unto you, Till heaven and earth pass, one jot or one tittle shall in no wise pass from the law, till all be fulfilled.

\verse Whosoever therefore shall break one of these least commandments, and shall teach men so, he shall be called the least in the kingdom of heaven: but whosoever shall do and teach them, the same shall be called great in the kingdom of heaven.

\verse For I say unto you, That except your righteousness shall exceed the righteousness of the scribes and Pharisees, ye shall in no case enter into the kingdom of heaven.

\verse Ye have heard that it was said by them of old time, Thou shalt not kill; and whosoever shall kill shall be in danger of the judgment: \verse But I say unto you, That whosoever is angry with his brother without a cause shall be in danger of the judgment: and whosoever shall say to his brother, Raca, shall be in danger of the council: but whosoever shall say, Thou fool, shall be in danger of hell fire.

\verse Therefore if thou bring thy gift to the altar, and there rememberest that thy brother hath ought against thee; \verse Leave there thy gift before the altar, and go thy way; first be reconciled to thy brother, and then come and offer thy gift.

\verse Agree with thine adversary quickly, whiles thou art in the way with him; lest at any time the adversary deliver thee to the judge, and the judge deliver thee to the officer, and thou be cast into prison.

\verse Verily I say unto thee, Thou shalt by no means come out thence, till thou hast paid the uttermost farthing.

\verse Ye have heard that it was said by them of old time, Thou shalt not commit adultery: \verse But I say unto you, That whosoever looketh on a woman to lust after her hath committed adultery with her already in his heart.

\verse And if thy right eye offend thee, pluck it out, and cast it from thee: for it is profitable for thee that one of thy members should perish, and not that thy whole body should be cast into hell.

\verse And if thy right hand offend thee, cut it off, and cast it from thee: for it is profitable for thee that one of thy members should perish, and not that thy whole body should be cast into hell.

\verse It hath been said, Whosoever shall put away his wife, let him give her a writing of divorcement: \verse But I say unto you, That whosoever shall put away his wife, saving for the cause of fornication, causeth her to commit adultery: and whosoever shall marry her that is divorced committeth adultery.

\verse Again, ye have heard that it hath been said by them of old time, Thou shalt not forswear thyself, but shalt perform unto the Lord thine oaths: \verse But I say unto you, Swear not at all; neither by heaven; for it is God's throne: \verse Nor by the earth; for it is his footstool: neither by Jerusalem; for it is the city of the great King.

\verse Neither shalt thou swear by thy head, because thou canst not make one hair white or black.

\verse But let your communication be, Yea, yea; Nay, nay: for whatsoever is more than these cometh of evil.

\verse Ye have heard that it hath been said, An eye for an eye, and a tooth for a tooth: \verse But I say unto you, That ye resist not evil: but whosoever shall smite thee on thy right cheek, turn to him the other also.

\verse And if any man will sue thee at the law, and take away thy coat, let him have thy cloak also.

\verse And whosoever shall compel thee to go a mile, go with him twain.

\verse Give to him that asketh thee, and from him that would borrow of thee turn not thou away.

\verse Ye have heard that it hath been said, Thou shalt love thy neighbour, and hate thine enemy.

\verse But I say unto you, Love your enemies, bless them that curse you, do good to them that hate you, and pray for them which despitefully use you, and persecute you; \verse That ye may be the children of your Father which is in heaven: for he maketh his sun to rise on the evil and on the good, and sendeth rain on the just and on the unjust.

\verse For if ye love them which love you, what reward have ye? do not even the publicans the same?  \verse And if ye salute your brethren only, what do ye more than others? do not even the publicans so?  \verse Be ye therefore perfect, even as your Father which is in heaven is perfect.


\chapter
\verse Take heed that ye do not your alms before men, to be seen of them: otherwise ye have no reward of your Father which is in heaven.

\verse Therefore when thou doest thine alms, do not sound a trumpet before thee, as the hypocrites do in the synagogues and in the streets, that they may have glory of men. Verily I say unto you, They have their reward.

\verse But when thou doest alms, let not thy left hand know what thy right hand doeth: \verse That thine alms may be in secret: and thy Father which seeth in secret himself shall reward thee openly.

\verse And when thou prayest, thou shalt not be as the hypocrites are: for they love to pray standing in the synagogues and in the corners of the streets, that they may be seen of men. Verily I say unto you, They have their reward.

\verse But thou, when thou prayest, enter into thy closet, and when thou hast shut thy door, pray to thy Father which is in secret; and thy Father which seeth in secret shall reward thee openly.

\verse But when ye pray, use not vain repetitions, as the heathen do: for they think that they shall be heard for their much speaking.

\verse Be not ye therefore like unto them: for your Father knoweth what things ye have need of, before ye ask him.

\verse After this manner therefore pray ye: Our Father which art in heaven, Hallowed be thy name.

\verse Thy kingdom come, Thy will be done in earth, as it is in heaven.

\verse Give us this day our daily bread.

\verse And forgive us our debts, as we forgive our debtors.

\verse And lead us not into temptation, but deliver us from evil: For thine is the kingdom, and the power, and the glory, for ever. Amen.

\verse For if ye forgive men their trespasses, your heavenly Father will also forgive you: \verse But if ye forgive not men their trespasses, neither will your Father forgive your trespasses.

\verse Moreover when ye fast, be not, as the hypocrites, of a sad countenance: for they disfigure their faces, that they may appear unto men to fast. Verily I say unto you, They have their reward.

\verse But thou, when thou fastest, anoint thine head, and wash thy face; \verse That thou appear not unto men to fast, but unto thy Father which is in secret: and thy Father, which seeth in secret, shall reward thee openly.

\verse Lay not up for yourselves treasures upon earth, where moth and rust doth corrupt, and where thieves break through and steal: \verse But lay up for yourselves treasures in heaven, where neither moth nor rust doth corrupt, and where thieves do not break through nor steal: \verse For where your treasure is, there will your heart be also.

\verse The light of the body is the eye: if therefore thine eye be single, thy whole body shall be full of light.

\verse But if thine eye be evil, thy whole body shall be full of darkness.

If therefore the light that is in thee be darkness, how great is that darkness!  \verse No man can serve two masters: for either he will hate the one, and love the other; or else he will hold to the one, and despise the other. Ye cannot serve God and mammon.

\verse Therefore I say unto you, Take no thought for your life, what ye shall eat, or what ye shall drink; nor yet for your body, what ye shall put on. Is not the life more than meat, and the body than raiment?  \verse Behold the fowls of the air: for they sow not, neither do they reap, nor gather into barns; yet your heavenly Father feedeth them. Are ye not much better than they?  \verse Which of you by taking thought can add one cubit unto his stature?  \verse And why take ye thought for raiment? Consider the lilies of the field, how they grow; they toil not, neither do they spin: \verse And yet I say unto you, That even Solomon in all his glory was not arrayed like one of these.

\verse Wherefore, if God so clothe the grass of the field, which to day is, and to morrow is cast into the oven, shall he not much more clothe you, O ye of little faith?  \verse Therefore take no thought, saying, What shall we eat? or, What shall we drink? or, Wherewithal shall we be clothed?  \verse (For after all these things do the Gentiles seek:) for your heavenly Father knoweth that ye have need of all these things.

\verse But seek ye first the kingdom of God, and his righteousness; and all these things shall be added unto you.

\verse Take therefore no thought for the morrow: for the morrow shall take thought for the things of itself. Sufficient unto the day is the evil thereof.


\chapter
\verse Judge not, that ye be not judged.

\verse For with what judgment ye judge, ye shall be judged: and with what measure ye mete, it shall be measured to you again.

\verse And why beholdest thou the mote that is in thy brother's eye, but considerest not the beam that is in thine own eye?  \verse Or how wilt thou say to thy brother, Let me pull out the mote out of thine eye; and, behold, a beam is in thine own eye?  \verse Thou hypocrite, first cast out the beam out of thine own eye; and then shalt thou see clearly to cast out the mote out of thy brother's eye.

\verse Give not that which is holy unto the dogs, neither cast ye your pearls before swine, lest they trample them under their feet, and turn again and rend you.

\verse Ask, and it shall be given you; seek, and ye shall find; knock, and it shall be opened unto you: \verse For every one that asketh receiveth; and he that seeketh findeth; and to him that knocketh it shall be opened.

\verse Or what man is there of you, whom if his son ask bread, will he give him a stone?  \verse Or if he ask a fish, will he give him a serpent?  \verse If ye then, being evil, know how to give good gifts unto your children, how much more shall your Father which is in heaven give good things to them that ask him?  \verse Therefore all things whatsoever ye would that men should do to you, do ye even so to them: for this is the law and the prophets.

\verse Enter ye in at the strait gate: for wide is the gate, and broad is the way, that leadeth to destruction, and many there be which go in thereat: \verse Because strait is the gate, and narrow is the way, which leadeth unto life, and few there be that find it.

\verse Beware of false prophets, which come to you in sheep's clothing, but inwardly they are ravening wolves.

\verse Ye shall know them by their fruits. Do men gather grapes of thorns, or figs of thistles?  \verse Even so every good tree bringeth forth good fruit; but a corrupt tree bringeth forth evil fruit.

\verse A good tree cannot bring forth evil fruit, neither can a corrupt tree bring forth good fruit.

\verse Every tree that bringeth not forth good fruit is hewn down, and cast into the fire.

\verse Wherefore by their fruits ye shall know them.

\verse Not every one that saith unto me, Lord, Lord, shall enter into the kingdom of heaven; but he that doeth the will of my Father which is in heaven.

\verse Many will say to me in that day, Lord, Lord, have we not prophesied in thy name? and in thy name have cast out devils? and in thy name done many wonderful works?  \verse And then will I profess unto them, I never knew you: depart from me, ye that work iniquity.

\verse Therefore whosoever heareth these sayings of mine, and doeth them, I will liken him unto a wise man, which built his house upon a rock: \verse And the rain descended, and the floods came, and the winds blew, and beat upon that house; and it fell not: for it was founded upon a rock.

\verse And every one that heareth these sayings of mine, and doeth them not, shall be likened unto a foolish man, which built his house upon the sand: \verse And the rain descended, and the floods came, and the winds blew, and beat upon that house; and it fell: and great was the fall of it.

\verse And it came to pass, when Jesus had ended these sayings, the people were astonished at his doctrine: \verse For he taught them as one having authority, and not as the scribes.


\chapter
\verse When he was come down from the mountain, great multitudes followed him.

\verse And, behold, there came a leper and worshipped him, saying, Lord, if thou wilt, thou canst make me clean.

\verse And Jesus put forth his hand, and touched him, saying, I will; be thou clean. And immediately his leprosy was cleansed.

\verse And Jesus saith unto him, See thou tell no man; but go thy way, shew thyself to the priest, and offer the gift that Moses commanded, for a testimony unto them.

\verse And when Jesus was entered into Capernaum, there came unto him a centurion, beseeching him, \verse And saying, Lord, my servant lieth at home sick of the palsy, grievously tormented.

\verse And Jesus saith unto him, I will come and heal him.

\verse The centurion answered and said, Lord, I am not worthy that thou shouldest come under my roof: but speak the word only, and my servant shall be healed.

\verse For I am a man under authority, having soldiers under me: and I say to this man, Go, and he goeth; and to another, Come, and he cometh; and to my servant, Do this, and he doeth it.

\verse When Jesus heard it, he marvelled, and said to them that followed, Verily I say unto you, I have not found so great faith, no, not in Israel.

\verse And I say unto you, That many shall come from the east and west, and shall sit down with Abraham, and Isaac, and Jacob, in the kingdom of heaven.

\verse But the children of the kingdom shall be cast out into outer darkness: there shall be weeping and gnashing of teeth.

\verse And Jesus said unto the centurion, Go thy way; and as thou hast believed, so be it done unto thee. And his servant was healed in the selfsame hour.

\verse And when Jesus was come into Peter's house, he saw his wife's mother laid, and sick of a fever.

\verse And he touched her hand, and the fever left her: and she arose, and ministered unto them.

\verse When the even was come, they brought unto him many that were possessed with devils: and he cast out the spirits with his word, and healed all that were sick: \verse That it might be fulfilled which was spoken by Esaias the prophet, saying, Himself took our infirmities, and bare our sicknesses.

\verse Now when Jesus saw great multitudes about him, he gave commandment to depart unto the other side.

\verse And a certain scribe came, and said unto him, Master, I will follow thee whithersoever thou goest.

\verse And Jesus saith unto him, The foxes have holes, and the birds of the air have nests; but the Son of man hath not where to lay his head.

\verse And another of his disciples said unto him, Lord, suffer me first to go and bury my father.

\verse But Jesus said unto him, Follow me; and let the dead bury their dead.

\verse And when he was entered into a ship, his disciples followed him.

\verse And, behold, there arose a great tempest in the sea, insomuch that the ship was covered with the waves: but he was asleep.

\verse And his disciples came to him, and awoke him, saying, Lord, save us: we perish.

\verse And he saith unto them, Why are ye fearful, O ye of little faith?  Then he arose, and rebuked the winds and the sea; and there was a great calm.

\verse But the men marvelled, saying, What manner of man is this, that even the winds and the sea obey him!  \verse And when he was come to the other side into the country of the Gergesenes, there met him two possessed with devils, coming out of the tombs, exceeding fierce, so that no man might pass by that way.

\verse And, behold, they cried out, saying, What have we to do with thee, Jesus, thou Son of God? art thou come hither to torment us before the time?  \verse And there was a good way off from them an herd of many swine feeding.

\verse So the devils besought him, saying, If thou cast us out, suffer us to go away into the herd of swine.

\verse And he said unto them, Go. And when they were come out, they went into the herd of swine: and, behold, the whole herd of swine ran violently down a steep place into the sea, and perished in the waters.

\verse And they that kept them fled, and went their ways into the city, and told every thing, and what was befallen to the possessed of the devils.

\verse And, behold, the whole city came out to meet Jesus: and when they saw him, they besought him that he would depart out of their coasts.


\chapter
\verse And he entered into a ship, and passed over, and came into his own city.

\verse And, behold, they brought to him a man sick of the palsy, lying on a bed: and Jesus seeing their faith said unto the sick of the palsy; Son, be of good cheer; thy sins be forgiven thee.

\verse And, behold, certain of the scribes said within themselves, This man blasphemeth.

\verse And Jesus knowing their thoughts said, Wherefore think ye evil in your hearts?  \verse For whether is easier, to say, Thy sins be forgiven thee; or to say, Arise, and walk?  \verse But that ye may know that the Son of man hath power on earth to forgive sins, (then saith he to the sick of the palsy,) Arise, take up thy bed, and go unto thine house.

\verse And he arose, and departed to his house.

\verse But when the multitudes saw it, they marvelled, and glorified God, which had given such power unto men.

\verse And as Jesus passed forth from thence, he saw a man, named Matthew, sitting at the receipt of custom: and he saith unto him, Follow me. And he arose, and followed him.

\verse And it came to pass, as Jesus sat at meat in the house, behold, many publicans and sinners came and sat down with him and his disciples.

\verse And when the Pharisees saw it, they said unto his disciples, Why eateth your Master with publicans and sinners?  \verse But when Jesus heard that, he said unto them, They that be whole need not a physician, but they that are sick.

\verse But go ye and learn what that meaneth, I will have mercy, and not sacrifice: for I am not come to call the righteous, but sinners to repentance.

\verse Then came to him the disciples of John, saying, Why do we and the Pharisees fast oft, but thy disciples fast not?  \verse And Jesus said unto them, Can the children of the bridechamber mourn, as long as the bridegroom is with them? but the days will come, when the bridegroom shall be taken from them, and then shall they fast.

\verse No man putteth a piece of new cloth unto an old garment, for that which is put in to fill it up taketh from the garment, and the rent is made worse.

\verse Neither do men put new wine into old bottles: else the bottles break, and the wine runneth out, and the bottles perish: but they put new wine into new bottles, and both are preserved.

\verse While he spake these things unto them, behold, there came a certain ruler, and worshipped him, saying, My daughter is even now dead: but come and lay thy hand upon her, and she shall live.

\verse And Jesus arose, and followed him, and so did his disciples.

\verse And, behold, a woman, which was diseased with an issue of blood twelve years, came behind him, and touched the hem of his garment: \verse For she said within herself, If I may but touch his garment, I shall be whole.

\verse But Jesus turned him about, and when he saw her, he said, Daughter, be of good comfort; thy faith hath made thee whole. And the woman was made whole from that hour.

\verse And when Jesus came into the ruler's house, and saw the minstrels and the people making a noise, \verse He said unto them, Give place: for the maid is not dead, but sleepeth. And they laughed him to scorn.

\verse But when the people were put forth, he went in, and took her by the hand, and the maid arose.

\verse And the fame hereof went abroad into all that land.

\verse And when Jesus departed thence, two blind men followed him, crying, and saying, Thou son of David, have mercy on us.

\verse And when he was come into the house, the blind men came to him: and Jesus saith unto them, Believe ye that I am able to do this? They said unto him, Yea, Lord.

\verse Then touched he their eyes, saying, According to your faith be it unto you.

\verse And their eyes were opened; and Jesus straitly charged them, saying, See that no man know it.

\verse But they, when they were departed, spread abroad his fame in all that country.

\verse As they went out, behold, they brought to him a dumb man possessed with a devil.

\verse And when the devil was cast out, the dumb spake: and the multitudes marvelled, saying, It was never so seen in Israel.

\verse But the Pharisees said, He casteth out devils through the prince of the devils.

\verse And Jesus went about all the cities and villages, teaching in their synagogues, and preaching the gospel of the kingdom, and healing every sickness and every disease among the people.

\verse But when he saw the multitudes, he was moved with compassion on them, because they fainted, and were scattered abroad, as sheep having no shepherd.

\verse Then saith he unto his disciples, The harvest truly is plenteous, but the labourers are few; \verse Pray ye therefore the Lord of the harvest, that he will send forth labourers into his harvest.


\chapter
\verse And when he had called unto him his twelve disciples, he gave them power against unclean spirits, to cast them out, and to heal all manner of sickness and all manner of disease.

\verse Now the names of the twelve apostles are these; The first, Simon, who is called Peter, and Andrew his brother; James the son of Zebedee, and John his brother; \verse Philip, and Bartholomew; Thomas, and Matthew the publican; James the son of Alphaeus, and Lebbaeus, whose surname was Thaddaeus; \verse Simon the Canaanite, and Judas Iscariot, who also betrayed him.

\verse These twelve Jesus sent forth, and commanded them, saying, Go not into the way of the Gentiles, and into any city of the Samaritans enter ye not: \verse But go rather to the lost sheep of the house of Israel.

\verse And as ye go, preach, saying, The kingdom of heaven is at hand.

\verse Heal the sick, cleanse the lepers, raise the dead, cast out devils: freely ye have received, freely give.

\verse Provide neither gold, nor silver, nor brass in your purses, \verse Nor scrip for your journey, neither two coats, neither shoes, nor yet staves: for the workman is worthy of his meat.

\verse And into whatsoever city or town ye shall enter, enquire who in it is worthy; and there abide till ye go thence.

\verse And when ye come into an house, salute it.

\verse And if the house be worthy, let your peace come upon it: but if it be not worthy, let your peace return to you.

\verse And whosoever shall not receive you, nor hear your words, when ye depart out of that house or city, shake off the dust of your feet.

\verse Verily I say unto you, It shall be more tolerable for the land of Sodom and Gomorrha in the day of judgment, than for that city.

\verse Behold, I send you forth as sheep in the midst of wolves: be ye therefore wise as serpents, and harmless as doves.

\verse But beware of men: for they will deliver you up to the councils, and they will scourge you in their synagogues; \verse And ye shall be brought before governors and kings for my sake, for a testimony against them and the Gentiles.

\verse But when they deliver you up, take no thought how or what ye shall speak: for it shall be given you in that same hour what ye shall speak.

\verse For it is not ye that speak, but the Spirit of your Father which speaketh in you.

\verse And the brother shall deliver up the brother to death, and the father the child: and the children shall rise up against their parents, and cause them to be put to death.

\verse And ye shall be hated of all men for my name's sake: but he that endureth to the end shall be saved.

\verse But when they persecute you in this city, flee ye into another: for verily I say unto you, Ye shall not have gone over the cities of Israel, till the Son of man be come.

\verse The disciple is not above his master, nor the servant above his lord.

\verse It is enough for the disciple that he be as his master, and the servant as his lord. If they have called the master of the house Beelzebub, how much more shall they call them of his household?  \verse Fear them not therefore: for there is nothing covered, that shall not be revealed; and hid, that shall not be known.

\verse What I tell you in darkness, that speak ye in light: and what ye hear in the ear, that preach ye upon the housetops.

\verse And fear not them which kill the body, but are not able to kill the soul: but rather fear him which is able to destroy both soul and body in hell.

\verse Are not two sparrows sold for a farthing? and one of them shall not fall on the ground without your Father.

\verse But the very hairs of your head are all numbered.

\verse Fear ye not therefore, ye are of more value than many sparrows.

\verse Whosoever therefore shall confess me before men, him will I confess also before my Father which is in heaven.

\verse But whosoever shall deny me before men, him will I also deny before my Father which is in heaven.

\verse Think not that I am come to send peace on earth: I came not to send peace, but a sword.

\verse For I am come to set a man at variance against his father, and the daughter against her mother, and the daughter in law against her mother in law.

\verse And a man's foes shall be they of his own household.

\verse He that loveth father or mother more than me is not worthy of me: and he that loveth son or daughter more than me is not worthy of me.

\verse And he that taketh not his cross, and followeth after me, is not worthy of me.

\verse He that findeth his life shall lose it: and he that loseth his life for my sake shall find it.

\verse He that receiveth you receiveth me, and he that receiveth me receiveth him that sent me.

\verse He that receiveth a prophet in the name of a prophet shall receive a prophet's reward; and he that receiveth a righteous man in the name of a righteous man shall receive a righteous man's reward.

\verse And whosoever shall give to drink unto one of these little ones a cup of cold water only in the name of a disciple, verily I say unto you, he shall in no wise lose his reward.


\chapter
\verse And it came to pass, when Jesus had made an end of commanding his twelve disciples, he departed thence to teach and to preach in their cities.

\verse Now when John had heard in the prison the works of Christ, he sent two of his disciples, \verse And said unto him, Art thou he that should come, or do we look for another?  \verse Jesus answered and said unto them, Go and shew John again those things which ye do hear and see: \verse The blind receive their sight, and the lame walk, the lepers are cleansed, and the deaf hear, the dead are raised up, and the poor have the gospel preached to them.

\verse And blessed is he, whosoever shall not be offended in me.

\verse And as they departed, Jesus began to say unto the multitudes concerning John, What went ye out into the wilderness to see? A reed shaken with the wind?  \verse But what went ye out for to see? A man clothed in soft raiment?  behold, they that wear soft clothing are in kings' houses.

\verse But what went ye out for to see? A prophet? yea, I say unto you, and more than a prophet.

\verse For this is he, of whom it is written, Behold, I send my messenger before thy face, which shall prepare thy way before thee.

\verse Verily I say unto you, Among them that are born of women there hath not risen a greater than John the Baptist: notwithstanding he that is least in the kingdom of heaven is greater than he.

\verse And from the days of John the Baptist until now the kingdom of heaven suffereth violence, and the violent take it by force.

\verse For all the prophets and the law prophesied until John.

\verse And if ye will receive it, this is Elias, which was for to come.

\verse He that hath ears to hear, let him hear.

\verse But whereunto shall I liken this generation? It is like unto children sitting in the markets, and calling unto their fellows, \verse And saying, We have piped unto you, and ye have not danced; we have mourned unto you, and ye have not lamented.

\verse For John came neither eating nor drinking, and they say, He hath a devil.

\verse The Son of man came eating and drinking, and they say, Behold a man gluttonous, and a winebibber, a friend of publicans and sinners.  But wisdom is justified of her children.

\verse Then began he to upbraid the cities wherein most of his mighty works were done, because they repented not: \verse Woe unto thee, Chorazin! woe unto thee, Bethsaida! for if the mighty works, which were done in you, had been done in Tyre and Sidon, they would have repented long ago in sackcloth and ashes.

\verse But I say unto you, It shall be more tolerable for Tyre and Sidon at the day of judgment, than for you.

\verse And thou, Capernaum, which art exalted unto heaven, shalt be brought down to hell: for if the mighty works, which have been done in thee, had been done in Sodom, it would have remained until this day.

\verse But I say unto you, That it shall be more tolerable for the land of Sodom in the day of judgment, than for thee.

\verse At that time Jesus answered and said, I thank thee, O Father, Lord of heaven and earth, because thou hast hid these things from the wise and prudent, and hast revealed them unto babes.

\verse Even so, Father: for so it seemed good in thy sight.

\verse All things are delivered unto me of my Father: and no man knoweth the Son, but the Father; neither knoweth any man the Father, save the Son, and he to whomsoever the Son will reveal him.

\verse Come unto me, all ye that labour and are heavy laden, and I will give you rest.

\verse Take my yoke upon you, and learn of me; for I am meek and lowly in heart: and ye shall find rest unto your souls.

\verse For my yoke is easy, and my burden is light.


\chapter
\verse At that time Jesus went on the sabbath day through the corn; and his disciples were an hungred, and began to pluck the ears of corn and to eat.

\verse But when the Pharisees saw it, they said unto him, Behold, thy disciples do that which is not lawful to do upon the sabbath day.

\verse But he said unto them, Have ye not read what David did, when he was an hungred, and they that were with him; \verse How he entered into the house of God, and did eat the shewbread, which was not lawful for him to eat, neither for them which were with him, but only for the priests?  \verse Or have ye not read in the law, how that on the sabbath days the priests in the temple profane the sabbath, and are blameless?  \verse But I say unto you, That in this place is one greater than the temple.

\verse But if ye had known what this meaneth, I will have mercy, and not sacrifice, ye would not have condemned the guiltless.

\verse For the Son of man is Lord even of the sabbath day.

\verse And when he was departed thence, he went into their synagogue: \verse And, behold, there was a man which had his hand withered. And they asked him, saying, Is it lawful to heal on the sabbath days? that they might accuse him.

\verse And he said unto them, What man shall there be among you, that shall have one sheep, and if it fall into a pit on the sabbath day, will he not lay hold on it, and lift it out?  \verse How much then is a man better than a sheep? Wherefore it is lawful to do well on the sabbath days.

\verse Then saith he to the man, Stretch forth thine hand. And he stretched it forth; and it was restored whole, like as the other.

\verse Then the Pharisees went out, and held a council against him, how they might destroy him.

\verse But when Jesus knew it, he withdrew himself from thence: and great multitudes followed him, and he healed them all; \verse And charged them that they should not make him known: \verse That it might be fulfilled which was spoken by Esaias the prophet, saying, \verse Behold my servant, whom I have chosen; my beloved, in whom my soul is well pleased: I will put my spirit upon him, and he shall shew judgment to the Gentiles.

\verse He shall not strive, nor cry; neither shall any man hear his voice in the streets.

\verse A bruised reed shall he not break, and smoking flax shall he not quench, till he send forth judgment unto victory.

\verse And in his name shall the Gentiles trust.

\verse Then was brought unto him one possessed with a devil, blind, and dumb: and he healed him, insomuch that the blind and dumb both spake and saw.

\verse And all the people were amazed, and said, Is not this the son of David?  \verse But when the Pharisees heard it, they said, This fellow doth not cast out devils, but by Beelzebub the prince of the devils.

\verse And Jesus knew their thoughts, and said unto them, Every kingdom divided against itself is brought to desolation; and every city or house divided against itself shall not stand: \verse And if Satan cast out Satan, he is divided against himself; how shall then his kingdom stand?  \verse And if I by Beelzebub cast out devils, by whom do your children cast them out? therefore they shall be your judges.

\verse But if I cast out devils by the Spirit of God, then the kingdom of God is come unto you.

\verse Or else how can one enter into a strong man's house, and spoil his goods, except he first bind the strong man? and then he will spoil his house.

\verse He that is not with me is against me; and he that gathereth not with me scattereth abroad.

\verse Wherefore I say unto you, All manner of sin and blasphemy shall be forgiven unto men: but the blasphemy against the Holy Ghost shall not be forgiven unto men.

\verse And whosoever speaketh a word against the Son of man, it shall be forgiven him: but whosoever speaketh against the Holy Ghost, it shall not be forgiven him, neither in this world, neither in the world to come.

\verse Either make the tree good, and his fruit good; or else make the tree corrupt, and his fruit corrupt: for the tree is known by his fruit.

\verse O generation of vipers, how can ye, being evil, speak good things?  for out of the abundance of the heart the mouth speaketh.

\verse A good man out of the good treasure of the heart bringeth forth good things: and an evil man out of the evil treasure bringeth forth evil things.

\verse But I say unto you, That every idle word that men shall speak, they shall give account thereof in the day of judgment.

\verse For by thy words thou shalt be justified, and by thy words thou shalt be condemned.

\verse Then certain of the scribes and of the Pharisees answered, saying, Master, we would see a sign from thee.

\verse But he answered and said unto them, An evil and adulterous generation seeketh after a sign; and there shall no sign be given to it, but the sign of the prophet Jonas: \verse For as Jonas was three days and three nights in the whale's belly; so shall the Son of man be three days and three nights in the heart of the earth.

\verse The men of Nineveh shall rise in judgment with this generation, and shall condemn it: because they repented at the preaching of Jonas; and, behold, a greater than Jonas is here.

\verse The queen of the south shall rise up in the judgment with this generation, and shall condemn it: for she came from the uttermost parts of the earth to hear the wisdom of Solomon; and, behold, a greater than Solomon is here.

\verse When the unclean spirit is gone out of a man, he walketh through dry places, seeking rest, and findeth none.

\verse Then he saith, I will return into my house from whence I came out; and when he is come, he findeth it empty, swept, and garnished.

\verse Then goeth he, and taketh with himself seven other spirits more wicked than himself, and they enter in and dwell there: and the last state of that man is worse than the first. Even so shall it be also unto this wicked generation.

\verse While he yet talked to the people, behold, his mother and his brethren stood without, desiring to speak with him.

\verse Then one said unto him, Behold, thy mother and thy brethren stand without, desiring to speak with thee.

\verse But he answered and said unto him that told him, Who is my mother?  and who are my brethren?  \verse And he stretched forth his hand toward his disciples, and said, Behold my mother and my brethren!  \verse For whosoever shall do the will of my Father which is in heaven, the same is my brother, and sister, and mother.


\chapter
\verse The same day went Jesus out of the house, and sat by the sea side.

\verse And great multitudes were gathered together unto him, so that he went into a ship, and sat; and the whole multitude stood on the shore.

\verse And he spake many things unto them in parables, saying, Behold, a sower went forth to sow; \verse And when he sowed, some seeds fell by the way side, and the fowls came and devoured them up: \verse Some fell upon stony places, where they had not much earth: and forthwith they sprung up, because they had no deepness of earth: \verse And when the sun was up, they were scorched; and because they had no root, they withered away.

\verse And some fell among thorns; and the thorns sprung up, and choked them: \verse But other fell into good ground, and brought forth fruit, some an hundredfold, some sixtyfold, some thirtyfold.

\verse Who hath ears to hear, let him hear.

\verse And the disciples came, and said unto him, Why speakest thou unto them in parables?  \verse He answered and said unto them, Because it is given unto you to know the mysteries of the kingdom of heaven, but to them it is not given.

\verse For whosoever hath, to him shall be given, and he shall have more abundance: but whosoever hath not, from him shall be taken away even that he hath.

\verse Therefore speak I to them in parables: because they seeing see not; and hearing they hear not, neither do they understand.

\verse And in them is fulfilled the prophecy of Esaias, which saith, By hearing ye shall hear, and shall not understand; and seeing ye shall see, and shall not perceive: \verse For this people's heart is waxed gross, and their ears are dull of hearing, and their eyes they have closed; lest at any time they should see with their eyes and hear with their ears, and should understand with their heart, and should be converted, and I should heal them.

\verse But blessed are your eyes, for they see: and your ears, for they hear.

\verse For verily I say unto you, That many prophets and righteous men have desired to see those things which ye see, and have not seen them; and to hear those things which ye hear, and have not heard them.

\verse Hear ye therefore the parable of the sower.

\verse When any one heareth the word of the kingdom, and understandeth it not, then cometh the wicked one, and catcheth away that which was sown in his heart. This is he which received seed by the way side.

\verse But he that received the seed into stony places, the same is he that heareth the word, and anon with joy receiveth it; \verse Yet hath he not root in himself, but dureth for a while: for when tribulation or persecution ariseth because of the word, by and by he is offended.

\verse He also that received seed among the thorns is he that heareth the word; and the care of this world, and the deceitfulness of riches, choke the word, and he becometh unfruitful.

\verse But he that received seed into the good ground is he that heareth the word, and understandeth it; which also beareth fruit, and bringeth forth, some an hundredfold, some sixty, some thirty.

\verse Another parable put he forth unto them, saying, The kingdom of heaven is likened unto a man which sowed good seed in his field: \verse But while men slept, his enemy came and sowed tares among the wheat, and went his way.

\verse But when the blade was sprung up, and brought forth fruit, then appeared the tares also.

\verse So the servants of the householder came and said unto him, Sir, didst not thou sow good seed in thy field? from whence then hath it tares?  \verse He said unto them, An enemy hath done this. The servants said unto him, Wilt thou then that we go and gather them up?  \verse But he said, Nay; lest while ye gather up the tares, ye root up also the wheat with them.

\verse Let both grow together until the harvest: and in the time of harvest I will say to the reapers, Gather ye together first the tares, and bind them in bundles to burn them: but gather the wheat into my barn.

\verse Another parable put he forth unto them, saying, The kingdom of heaven is like to a grain of mustard seed, which a man took, and sowed in his field: \verse Which indeed is the least of all seeds: but when it is grown, it is the greatest among herbs, and becometh a tree, so that the birds of the air come and lodge in the branches thereof.

\verse Another parable spake he unto them; The kingdom of heaven is like unto leaven, which a woman took, and hid in three measures of meal, till the whole was leavened.

\verse All these things spake Jesus unto the multitude in parables; and without a parable spake he not unto them: \verse That it might be fulfilled which was spoken by the prophet, saying, I will open my mouth in parables; I will utter things which have been kept secret from the foundation of the world.

\verse Then Jesus sent the multitude away, and went into the house: and his disciples came unto him, saying, Declare unto us the parable of the tares of the field.

\verse He answered and said unto them, He that soweth the good seed is the Son of man; \verse The field is the world; the good seed are the children of the kingdom; but the tares are the children of the wicked one; \verse The enemy that sowed them is the devil; the harvest is the end of the world; and the reapers are the angels.

\verse As therefore the tares are gathered and burned in the fire; so shall it be in the end of this world.

\verse The Son of man shall send forth his angels, and they shall gather out of his kingdom all things that offend, and them which do iniquity; \verse And shall cast them into a furnace of fire: there shall be wailing and gnashing of teeth.

\verse Then shall the righteous shine forth as the sun in the kingdom of their Father. Who hath ears to hear, let him hear.

\verse Again, the kingdom of heaven is like unto treasure hid in a field; the which when a man hath found, he hideth, and for joy thereof goeth and selleth all that he hath, and buyeth that field.

\verse Again, the kingdom of heaven is like unto a merchant man, seeking goodly pearls: \verse Who, when he had found one pearl of great price, went and sold all that he had, and bought it.

\verse Again, the kingdom of heaven is like unto a net, that was cast into the sea, and gathered of every kind: \verse Which, when it was full, they drew to shore, and sat down, and gathered the good into vessels, but cast the bad away.

\verse So shall it be at the end of the world: the angels shall come forth, and sever the wicked from among the just, \verse And shall cast them into the furnace of fire: there shall be wailing and gnashing of teeth.

\verse Jesus saith unto them, Have ye understood all these things? They say unto him, Yea, Lord.

\verse Then said he unto them, Therefore every scribe which is instructed unto the kingdom of heaven is like unto a man that is an householder, which bringeth forth out of his treasure things new and old.

\verse And it came to pass, that when Jesus had finished these parables, he departed thence.

\verse And when he was come into his own country, he taught them in their synagogue, insomuch that they were astonished, and said, Whence hath this man this wisdom, and these mighty works?  \verse Is not this the carpenter's son? is not his mother called Mary? and his brethren, James, and Joses, and Simon, and Judas?  \verse And his sisters, are they not all with us? Whence then hath this man all these things?  \verse And they were offended in him. But Jesus said unto them, A prophet is not without honour, save in his own country, and in his own house.

\verse And he did not many mighty works there because of their unbelief.


\chapter
\verse At that time Herod the tetrarch heard of the fame of Jesus, \verse And said unto his servants, This is John the Baptist; he is risen from the dead; and therefore mighty works do shew forth themselves in him.

\verse For Herod had laid hold on John, and bound him, and put him in prison for Herodias' sake, his brother Philip's wife.

\verse For John said unto him, It is not lawful for thee to have her.

\verse And when he would have put him to death, he feared the multitude, because they counted him as a prophet.

\verse But when Herod's birthday was kept, the daughter of Herodias danced before them, and pleased Herod.

\verse Whereupon he promised with an oath to give her whatsoever she would ask.

\verse And she, being before instructed of her mother, said, Give me here John Baptist's head in a charger.

\verse And the king was sorry: nevertheless for the oath's sake, and them which sat with him at meat, he commanded it to be given her.

\verse And he sent, and beheaded John in the prison.

\verse And his head was brought in a charger, and given to the damsel: and she brought it to her mother.

\verse And his disciples came, and took up the body, and buried it, and went and told Jesus.

\verse When Jesus heard of it, he departed thence by ship into a desert place apart: and when the people had heard thereof, they followed him on foot out of the cities.

\verse And Jesus went forth, and saw a great multitude, and was moved with compassion toward them, and he healed their sick.

\verse And when it was evening, his disciples came to him, saying, This is a desert place, and the time is now past; send the multitude away, that they may go into the villages, and buy themselves victuals.

\verse But Jesus said unto them, They need not depart; give ye them to eat.

\verse And they say unto him, We have here but five loaves, and two fishes.

\verse He said, Bring them hither to me.

\verse And he commanded the multitude to sit down on the grass, and took the five loaves, and the two fishes, and looking up to heaven, he blessed, and brake, and gave the loaves to his disciples, and the disciples to the multitude.

\verse And they did all eat, and were filled: and they took up of the fragments that remained twelve baskets full.

\verse And they that had eaten were about five thousand men, beside women and children.

\verse And straightway Jesus constrained his disciples to get into a ship, and to go before him unto the other side, while he sent the multitudes away.

\verse And when he had sent the multitudes away, he went up into a mountain apart to pray: and when the evening was come, he was there alone.

\verse But the ship was now in the midst of the sea, tossed with waves: for the wind was contrary.

\verse And in the fourth watch of the night Jesus went unto them, walking on the sea.

\verse And when the disciples saw him walking on the sea, they were troubled, saying, It is a spirit; and they cried out for fear.

\verse But straightway Jesus spake unto them, saying, Be of good cheer; it is I; be not afraid.

\verse And Peter answered him and said, Lord, if it be thou, bid me come unto thee on the water.

\verse And he said, Come. And when Peter was come down out of the ship, he walked on the water, to go to Jesus.

\verse But when he saw the wind boisterous, he was afraid; and beginning to sink, he cried, saying, Lord, save me.

\verse And immediately Jesus stretched forth his hand, and caught him, and said unto him, O thou of little faith, wherefore didst thou doubt?  \verse And when they were come into the ship, the wind ceased.

\verse Then they that were in the ship came and worshipped him, saying, Of a truth thou art the Son of God.

\verse And when they were gone over, they came into the land of Gennesaret.

\verse And when the men of that place had knowledge of him, they sent out into all that country round about, and brought unto him all that were diseased; \verse And besought him that they might only touch the hem of his garment: and as many as touched were made perfectly whole.


\chapter
\verse Then came to Jesus scribes and Pharisees, which were of Jerusalem, saying, \verse Why do thy disciples transgress the tradition of the elders? for they wash not their hands when they eat bread.

\verse But he answered and said unto them, Why do ye also transgress the commandment of God by your tradition?  \verse For God commanded, saying, Honour thy father and mother: and, He that curseth father or mother, let him die the death.

\verse But ye say, Whosoever shall say to his father or his mother, It is a gift, by whatsoever thou mightest be profited by me; \verse And honour not his father or his mother, he shall be free. Thus have ye made the commandment of God of none effect by your tradition.

\verse Ye hypocrites, well did Esaias prophesy of you, saying, \verse This people draweth nigh unto me with their mouth, and honoureth me with their lips; but their heart is far from me.

\verse But in vain they do worship me, teaching for doctrines the commandments of men.

\verse And he called the multitude, and said unto them, Hear, and understand: \verse Not that which goeth into the mouth defileth a man; but that which cometh out of the mouth, this defileth a man.

\verse Then came his disciples, and said unto him, Knowest thou that the Pharisees were offended, after they heard this saying?  \verse But he answered and said, Every plant, which my heavenly Father hath not planted, shall be rooted up.

\verse Let them alone: they be blind leaders of the blind. And if the blind lead the blind, both shall fall into the ditch.

\verse Then answered Peter and said unto him, Declare unto us this parable.

\verse And Jesus said, Are ye also yet without understanding?  \verse Do not ye yet understand, that whatsoever entereth in at the mouth goeth into the belly, and is cast out into the draught?  \verse But those things which proceed out of the mouth come forth from the heart; and they defile the man.

\verse For out of the heart proceed evil thoughts, murders, adulteries, fornications, thefts, false witness, blasphemies: \verse These are the things which defile a man: but to eat with unwashen hands defileth not a man.

\verse Then Jesus went thence, and departed into the coasts of Tyre and Sidon.

\verse And, behold, a woman of Canaan came out of the same coasts, and cried unto him, saying, Have mercy on me, O Lord, thou son of David; my daughter is grievously vexed with a devil.

\verse But he answered her not a word. And his disciples came and besought him, saying, Send her away; for she crieth after us.

\verse But he answered and said, I am not sent but unto the lost sheep of the house of Israel.

\verse Then came she and worshipped him, saying, Lord, help me.

\verse But he answered and said, It is not meet to take the children's bread, and to cast it to dogs.

\verse And she said, Truth, Lord: yet the dogs eat of the crumbs which fall from their masters' table.

\verse Then Jesus answered and said unto her, O woman, great is thy faith: be it unto thee even as thou wilt. And her daughter was made whole from that very hour.

\verse And Jesus departed from thence, and came nigh unto the sea of Galilee; and went up into a mountain, and sat down there.

\verse And great multitudes came unto him, having with them those that were lame, blind, dumb, maimed, and many others, and cast them down at Jesus' feet; and he healed them: \verse Insomuch that the multitude wondered, when they saw the dumb to speak, the maimed to be whole, the lame to walk, and the blind to see: and they glorified the God of Israel.

\verse Then Jesus called his disciples unto him, and said, I have compassion on the multitude, because they continue with me now three days, and have nothing to eat: and I will not send them away fasting, lest they faint in the way.

\verse And his disciples say unto him, Whence should we have so much bread in the wilderness, as to fill so great a multitude?  \verse And Jesus saith unto them, How many loaves have ye? And they said, Seven, and a few little fishes.

\verse And he commanded the multitude to sit down on the ground.

\verse And he took the seven loaves and the fishes, and gave thanks, and brake them, and gave to his disciples, and the disciples to the multitude.

\verse And they did all eat, and were filled: and they took up of the broken meat that was left seven baskets full.

\verse And they that did eat were four thousand men, beside women and children.

\verse And he sent away the multitude, and took ship, and came into the coasts of Magdala.


\chapter
\verse The Pharisees also with the Sadducees came, and tempting desired him that he would shew them a sign from heaven.

\verse He answered and said unto them, When it is evening, ye say, It will be fair weather: for the sky is red.

\verse And in the morning, It will be foul weather to day: for the sky is red and lowering. O ye hypocrites, ye can discern the face of the sky; but can ye not discern the signs of the times?  \verse A wicked and adulterous generation seeketh after a sign; and there shall no sign be given unto it, but the sign of the prophet Jonas. And he left them, and departed.

\verse And when his disciples were come to the other side, they had forgotten to take bread.

\verse Then Jesus said unto them, Take heed and beware of the leaven of the Pharisees and of the Sadducees.

\verse And they reasoned among themselves, saying, It is because we have taken no bread.

\verse Which when Jesus perceived, he said unto them, O ye of little faith, why reason ye among yourselves, because ye have brought no bread?  \verse Do ye not yet understand, neither remember the five loaves of the five thousand, and how many baskets ye took up?  \verse Neither the seven loaves of the four thousand, and how many baskets ye took up?  \verse How is it that ye do not understand that I spake it not to you concerning bread, that ye should beware of the leaven of the Pharisees and of the Sadducees?  \verse Then understood they how that he bade them not beware of the leaven of bread, but of the doctrine of the Pharisees and of the Sadducees.

\verse When Jesus came into the coasts of Caesarea Philippi, he asked his disciples, saying, Whom do men say that I the Son of man am?  \verse And they said, Some say that thou art John the Baptist: some, Elias; and others, Jeremias, or one of the prophets.

\verse He saith unto them, But whom say ye that I am?  \verse And Simon Peter answered and said, Thou art the Christ, the Son of the living God.

\verse And Jesus answered and said unto him, Blessed art thou, Simon Barjona: for flesh and blood hath not revealed it unto thee, but my Father which is in heaven.

\verse And I say also unto thee, That thou art Peter, and upon this rock I will build my church; and the gates of hell shall not prevail against it.

\verse And I will give unto thee the keys of the kingdom of heaven: and whatsoever thou shalt bind on earth shall be bound in heaven: and whatsoever thou shalt loose on earth shall be loosed in heaven.

\verse Then charged he his disciples that they should tell no man that he was Jesus the Christ.

\verse From that time forth began Jesus to shew unto his disciples, how that he must go unto Jerusalem, and suffer many things of the elders and chief priests and scribes, and be killed, and be raised again the third day.

\verse Then Peter took him, and began to rebuke him, saying, Be it far from thee, Lord: this shall not be unto thee.

\verse But he turned, and said unto Peter, Get thee behind me, Satan: thou art an offence unto me: for thou savourest not the things that be of God, but those that be of men.

\verse Then said Jesus unto his disciples, If any man will come after me, let him deny himself, and take up his cross, and follow me.

\verse For whosoever will save his life shall lose it: and whosoever will lose his life for my sake shall find it.

\verse For what is a man profited, if he shall gain the whole world, and lose his own soul? or what shall a man give in exchange for his soul?  \verse For the Son of man shall come in the glory of his Father with his angels; and then he shall reward every man according to his works.

\verse Verily I say unto you, There be some standing here, which shall not taste of death, till they see the Son of man coming in his kingdom.


\chapter
\verse And after six days Jesus taketh Peter, James, and John his brother, and bringeth them up into an high mountain apart, \verse And was transfigured before them: and his face did shine as the sun, and his raiment was white as the light.

\verse And, behold, there appeared unto them Moses and Elias talking with him.

\verse Then answered Peter, and said unto Jesus, Lord, it is good for us to be here: if thou wilt, let us make here three tabernacles; one for thee, and one for Moses, and one for Elias.

\verse While he yet spake, behold, a bright cloud overshadowed them: and behold a voice out of the cloud, which said, This is my beloved Son, in whom I am well pleased; hear ye him.

\verse And when the disciples heard it, they fell on their face, and were sore afraid.

\verse And Jesus came and touched them, and said, Arise, and be not afraid.

\verse And when they had lifted up their eyes, they saw no man, save Jesus only.

\verse And as they came down from the mountain, Jesus charged them, saying, Tell the vision to no man, until the Son of man be risen again from the dead.

\verse And his disciples asked him, saying, Why then say the scribes that Elias must first come?  \verse And Jesus answered and said unto them, Elias truly shall first come, and restore all things.

\verse But I say unto you, That Elias is come already, and they knew him not, but have done unto him whatsoever they listed. Likewise shall also the Son of man suffer of them.

\verse Then the disciples understood that he spake unto them of John the Baptist.

\verse And when they were come to the multitude, there came to him a certain man, kneeling down to him, and saying, \verse Lord, have mercy on my son: for he is lunatick, and sore vexed: for ofttimes he falleth into the fire, and oft into the water.

\verse And I brought him to thy disciples, and they could not cure him.

\verse Then Jesus answered and said, O faithless and perverse generation, how long shall I be with you? how long shall I suffer you?  bring him hither to me.

\verse And Jesus rebuked the devil; and he departed out of him: and the child was cured from that very hour.

\verse Then came the disciples to Jesus apart, and said, Why could not we cast him out?  \verse And Jesus said unto them, Because of your unbelief: for verily I say unto you, If ye have faith as a grain of mustard seed, ye shall say unto this mountain, Remove hence to yonder place; and it shall remove; and nothing shall be impossible unto you.

\verse Howbeit this kind goeth not out but by prayer and fasting.

\verse And while they abode in Galilee, Jesus said unto them, The Son of man shall be betrayed into the hands of men: \verse And they shall kill him, and the third day he shall be raised again.

And they were exceeding sorry.

\verse And when they were come to Capernaum, they that received tribute money came to Peter, and said, Doth not your master pay tribute?  \verse He saith, Yes. And when he was come into the house, Jesus prevented him, saying, What thinkest thou, Simon? of whom do the kings of the earth take custom or tribute? of their own children, or of strangers?  \verse Peter saith unto him, Of strangers. Jesus saith unto him, Then are the children free.

\verse Notwithstanding, lest we should offend them, go thou to the sea, and cast an hook, and take up the fish that first cometh up; and when thou hast opened his mouth, thou shalt find a piece of money: that take, and give unto them for me and thee.


\chapter
\verse At the same time came the disciples unto Jesus, saying, Who is the greatest in the kingdom of heaven?  \verse And Jesus called a little child unto him, and set him in the midst of them, \verse And said, Verily I say unto you, Except ye be converted, and become as little children, ye shall not enter into the kingdom of heaven.

\verse Whosoever therefore shall humble himself as this little child, the same is greatest in the kingdom of heaven.

\verse And whoso shall receive one such little child in my name receiveth me.

\verse But whoso shall offend one of these little ones which believe in me, it were better for him that a millstone were hanged about his neck, and that he were drowned in the depth of the sea.

\verse Woe unto the world because of offences! for it must needs be that offences come; but woe to that man by whom the offence cometh!  \verse Wherefore if thy hand or thy foot offend thee, cut them off, and cast them from thee: it is better for thee to enter into life halt or maimed, rather than having two hands or two feet to be cast into everlasting fire.

\verse And if thine eye offend thee, pluck it out, and cast it from thee: it is better for thee to enter into life with one eye, rather than having two eyes to be cast into hell fire.

\verse Take heed that ye despise not one of these little ones; for I say unto you, That in heaven their angels do always behold the face of my Father which is in heaven.

\verse For the Son of man is come to save that which was lost.

\verse How think ye? if a man have an hundred sheep, and one of them be gone astray, doth he not leave the ninety and nine, and goeth into the mountains, and seeketh that which is gone astray?  \verse And if so be that he find it, verily I say unto you, he rejoiceth more of that sheep, than of the ninety and nine which went not astray.

\verse Even so it is not the will of your Father which is in heaven, that one of these little ones should perish.

\verse Moreover if thy brother shall trespass against thee, go and tell him his fault between thee and him alone: if he shall hear thee, thou hast gained thy brother.

\verse But if he will not hear thee, then take with thee one or two more, that in the mouth of two or three witnesses every word may be established.

\verse And if he shall neglect to hear them, tell it unto the church: but if he neglect to hear the church, let him be unto thee as an heathen man and a publican.

\verse Verily I say unto you, Whatsoever ye shall bind on earth shall be bound in heaven: and whatsoever ye shall loose on earth shall be loosed in heaven.

\verse Again I say unto you, That if two of you shall agree on earth as touching any thing that they shall ask, it shall be done for them of my Father which is in heaven.

\verse For where two or three are gathered together in my name, there am I in the midst of them.

\verse Then came Peter to him, and said, Lord, how oft shall my brother sin against me, and I forgive him? till seven times?  \verse Jesus saith unto him, I say not unto thee, Until seven times: but, Until seventy times seven.

\verse Therefore is the kingdom of heaven likened unto a certain king, which would take account of his servants.

\verse And when he had begun to reckon, one was brought unto him, which owed him ten thousand talents.

\verse But forasmuch as he had not to pay, his lord commanded him to be sold, and his wife, and children, and all that he had, and payment to be made.

\verse The servant therefore fell down, and worshipped him, saying, Lord, have patience with me, and I will pay thee all.

\verse Then the lord of that servant was moved with compassion, and loosed him, and forgave him the debt.

\verse But the same servant went out, and found one of his fellowservants, which owed him an hundred pence: and he laid hands on him, and took him by the throat, saying, Pay me that thou owest.

\verse And his fellowservant fell down at his feet, and besought him, saying, Have patience with me, and I will pay thee all.

\verse And he would not: but went and cast him into prison, till he should pay the debt.

\verse So when his fellowservants saw what was done, they were very sorry, and came and told unto their lord all that was done.

\verse Then his lord, after that he had called him, said unto him, O thou wicked servant, I forgave thee all that debt, because thou desiredst me: \verse Shouldest not thou also have had compassion on thy fellowservant, even as I had pity on thee?  \verse And his lord was wroth, and delivered him to the tormentors, till he should pay all that was due unto him.

\verse So likewise shall my heavenly Father do also unto you, if ye from your hearts forgive not every one his brother their trespasses.


\chapter
\verse And it came to pass, that when Jesus had finished these sayings, he departed from Galilee, and came into the coasts of Judaea beyond Jordan; \verse And great multitudes followed him; and he healed them there.

\verse The Pharisees also came unto him, tempting him, and saying unto him, Is it lawful for a man to put away his wife for every cause?  \verse And he answered and said unto them, Have ye not read, that he which made them at the beginning made them male and female, \verse And said, For this cause shall a man leave father and mother, and shall cleave to his wife: and they twain shall be one flesh?  \verse Wherefore they are no more twain, but one flesh. What therefore God hath joined together, let not man put asunder.

\verse They say unto him, Why did Moses then command to give a writing of divorcement, and to put her away?  \verse He saith unto them, Moses because of the hardness of your hearts suffered you to put away your wives: but from the beginning it was not so.

\verse And I say unto you, Whosoever shall put away his wife, except it be for fornication, and shall marry another, committeth adultery: and whoso marrieth her which is put away doth commit adultery.

\verse His disciples say unto him, If the case of the man be so with his wife, it is not good to marry.

\verse But he said unto them, All men cannot receive this saying, save they to whom it is given.

\verse For there are some eunuchs, which were so born from their mother's womb: and there are some eunuchs, which were made eunuchs of men: and there be eunuchs, which have made themselves eunuchs for the kingdom of heaven's sake. He that is able to receive it, let him receive it.

\verse Then were there brought unto him little children, that he should put his hands on them, and pray: and the disciples rebuked them.

\verse But Jesus said, Suffer little children, and forbid them not, to come unto me: for of such is the kingdom of heaven.

\verse And he laid his hands on them, and departed thence.

\verse And, behold, one came and said unto him, Good Master, what good thing shall I do, that I may have eternal life?  \verse And he said unto him, Why callest thou me good? there is none good but one, that is, God: but if thou wilt enter into life, keep the commandments.

\verse He saith unto him, Which? Jesus said, Thou shalt do no murder, Thou shalt not commit adultery, Thou shalt not steal, Thou shalt not bear false witness, \verse Honour thy father and thy mother: and, Thou shalt love thy neighbour as thyself.

\verse The young man saith unto him, All these things have I kept from my youth up: what lack I yet?  \verse Jesus said unto him, If thou wilt be perfect, go and sell that thou hast, and give to the poor, and thou shalt have treasure in heaven: and come and follow me.

\verse But when the young man heard that saying, he went away sorrowful: for he had great possessions.

\verse Then said Jesus unto his disciples, Verily I say unto you, That a rich man shall hardly enter into the kingdom of heaven.

\verse And again I say unto you, It is easier for a camel to go through the eye of a needle, than for a rich man to enter into the kingdom of God.

\verse When his disciples heard it, they were exceedingly amazed, saying, Who then can be saved?  \verse But Jesus beheld them, and said unto them, With men this is impossible; but with God all things are possible.

\verse Then answered Peter and said unto him, Behold, we have forsaken all, and followed thee; what shall we have therefore?  \verse And Jesus said unto them, Verily I say unto you, That ye which have followed me, in the regeneration when the Son of man shall sit in the throne of his glory, ye also shall sit upon twelve thrones, judging the twelve tribes of Israel.

\verse And every one that hath forsaken houses, or brethren, or sisters, or father, or mother, or wife, or children, or lands, for my name's sake, shall receive an hundredfold, and shall inherit everlasting life.

\verse But many that are first shall be last; and the last shall be first.


\chapter
\verse For the kingdom of heaven is like unto a man that is an householder, which went out early in the morning to hire labourers into his vineyard.

\verse And when he had agreed with the labourers for a penny a day, he sent them into his vineyard.

\verse And he went out about the third hour, and saw others standing idle in the marketplace, \verse And said unto them; Go ye also into the vineyard, and whatsoever is right I will give you. And they went their way.

\verse Again he went out about the sixth and ninth hour, and did likewise.

\verse And about the eleventh hour he went out, and found others standing idle, and saith unto them, Why stand ye here all the day idle?  \verse They say unto him, Because no man hath hired us. He saith unto them, Go ye also into the vineyard; and whatsoever is right, that shall ye receive.

\verse So when even was come, the lord of the vineyard saith unto his steward, Call the labourers, and give them their hire, beginning from the last unto the first.

\verse And when they came that were hired about the eleventh hour, they received every man a penny.

\verse But when the first came, they supposed that they should have received more; and they likewise received every man a penny.

\verse And when they had received it, they murmured against the goodman of the house, \verse Saying, These last have wrought but one hour, and thou hast made them equal unto us, which have borne the burden and heat of the day.

\verse But he answered one of them, and said, Friend, I do thee no wrong: didst not thou agree with me for a penny?  \verse Take that thine is, and go thy way: I will give unto this last, even as unto thee.

\verse Is it not lawful for me to do what I will with mine own? Is thine eye evil, because I am good?  \verse So the last shall be first, and the first last: for many be called, but few chosen.

\verse And Jesus going up to Jerusalem took the twelve disciples apart in the way, and said unto them, \verse Behold, we go up to Jerusalem; and the Son of man shall be betrayed unto the chief priests and unto the scribes, and they shall condemn him to death, \verse And shall deliver him to the Gentiles to mock, and to scourge, and to crucify him: and the third day he shall rise again.

\verse Then came to him the mother of Zebedees children with her sons, worshipping him, and desiring a certain thing of him.

\verse And he said unto her, What wilt thou? She saith unto him, Grant that these my two sons may sit, the one on thy right hand, and the other on the left, in thy kingdom.

\verse But Jesus answered and said, Ye know not what ye ask. Are ye able to drink of the cup that I shall drink of, and to be baptized with the baptism that I am baptized with? They say unto him, We are able.

\verse And he saith unto them, Ye shall drink indeed of my cup, and be baptized with the baptism that I am baptized with: but to sit on my right hand, and on my left, is not mine to give, but it shall be given to them for whom it is prepared of my Father.

\verse And when the ten heard it, they were moved with indignation against the two brethren.

\verse But Jesus called them unto him, and said, Ye know that the princes of the Gentiles exercise dominion over them, and they that are great exercise authority upon them.

\verse But it shall not be so among you: but whosoever will be great among you, let him be your minister; \verse And whosoever will be chief among you, let him be your servant: \verse Even as the Son of man came not to be ministered unto, but to minister, and to give his life a ransom for many.

\verse And as they departed from Jericho, a great multitude followed him.

\verse And, behold, two blind men sitting by the way side, when they heard that Jesus passed by, cried out, saying, Have mercy on us, O Lord, thou son of David.

\verse And the multitude rebuked them, because they should hold their peace: but they cried the more, saying, Have mercy on us, O Lord, thou son of David.

\verse And Jesus stood still, and called them, and said, What will ye that I shall do unto you?  \verse They say unto him, Lord, that our eyes may be opened.

\verse So Jesus had compassion on them, and touched their eyes: and immediately their eyes received sight, and they followed him.


\chapter
\verse And when they drew nigh unto Jerusalem, and were come to Bethphage, unto the mount of Olives, then sent Jesus two disciples, \verse Saying unto them, Go into the village over against you, and straightway ye shall find an ass tied, and a colt with her: loose them, and bring them unto me.

\verse And if any man say ought unto you, ye shall say, The Lord hath need of them; and straightway he will send them.

\verse All this was done, that it might be fulfilled which was spoken by the prophet, saying, \verse Tell ye the daughter of Sion, Behold, thy King cometh unto thee, meek, and sitting upon an ass, and a colt the foal of an ass.

\verse And the disciples went, and did as Jesus commanded them, \verse And brought the ass, and the colt, and put on them their clothes, and they set him thereon.

\verse And a very great multitude spread their garments in the way; others cut down branches from the trees, and strawed them in the way.

\verse And the multitudes that went before, and that followed, cried, saying, Hosanna to the son of David: Blessed is he that cometh in the name of the Lord; Hosanna in the highest.

\verse And when he was come into Jerusalem, all the city was moved, saying, Who is this?  \verse And the multitude said, This is Jesus the prophet of Nazareth of Galilee.

\verse And Jesus went into the temple of God, and cast out all them that sold and bought in the temple, and overthrew the tables of the moneychangers, and the seats of them that sold doves, \verse And said unto them, It is written, My house shall be called the house of prayer; but ye have made it a den of thieves.

\verse And the blind and the lame came to him in the temple; and he healed them.

\verse And when the chief priests and scribes saw the wonderful things that he did, and the children crying in the temple, and saying, Hosanna to the son of David; they were sore displeased, \verse And said unto him, Hearest thou what these say? And Jesus saith unto them, Yea; have ye never read, Out of the mouth of babes and sucklings thou hast perfected praise?  \verse And he left them, and went out of the city into Bethany; and he lodged there.

\verse Now in the morning as he returned into the city, he hungered.

\verse And when he saw a fig tree in the way, he came to it, and found nothing thereon, but leaves only, and said unto it, Let no fruit grow on thee henceforward for ever. And presently the fig tree withered away.

\verse And when the disciples saw it, they marvelled, saying, How soon is the fig tree withered away!  \verse Jesus answered and said unto them, Verily I say unto you, If ye have faith, and doubt not, ye shall not only do this which is done to the fig tree, but also if ye shall say unto this mountain, Be thou removed, and be thou cast into the sea; it shall be done.

\verse And all things, whatsoever ye shall ask in prayer, believing, ye shall receive.

\verse And when he was come into the temple, the chief priests and the elders of the people came unto him as he was teaching, and said, By what authority doest thou these things? and who gave thee this authority?  \verse And Jesus answered and said unto them, I also will ask you one thing, which if ye tell me, I in like wise will tell you by what authority I do these things.

\verse The baptism of John, whence was it? from heaven, or of men? And they reasoned with themselves, saying, If we shall say, From heaven; he will say unto us, Why did ye not then believe him?  \verse But if we shall say, Of men; we fear the people; for all hold John as a prophet.

\verse And they answered Jesus, and said, We cannot tell. And he said unto them, Neither tell I you by what authority I do these things.

\verse But what think ye? A certain man had two sons; and he came to the first, and said, Son, go work to day in my vineyard.

\verse He answered and said, I will not: but afterward he repented, and went.

\verse And he came to the second, and said likewise. And he answered and said, I go, sir: and went not.

\verse Whether of them twain did the will of his father? They say unto him, The first. Jesus saith unto them, Verily I say unto you, That the publicans and the harlots go into the kingdom of God before you.

\verse For John came unto you in the way of righteousness, and ye believed him not: but the publicans and the harlots believed him: and ye, when ye had seen it, repented not afterward, that ye might believe him.

\verse Hear another parable: There was a certain householder, which planted a vineyard, and hedged it round about, and digged a winepress in it, and built a tower, and let it out to husbandmen, and went into a far country: \verse And when the time of the fruit drew near, he sent his servants to the husbandmen, that they might receive the fruits of it.

\verse And the husbandmen took his servants, and beat one, and killed another, and stoned another.

\verse Again, he sent other servants more than the first: and they did unto them likewise.

\verse But last of all he sent unto them his son, saying, They will reverence my son.

\verse But when the husbandmen saw the son, they said among themselves, This is the heir; come, let us kill him, and let us seize on his inheritance.

\verse And they caught him, and cast him out of the vineyard, and slew him.

\verse When the lord therefore of the vineyard cometh, what will he do unto those husbandmen?  \verse They say unto him, He will miserably destroy those wicked men, and will let out his vineyard unto other husbandmen, which shall render him the fruits in their seasons.

\verse Jesus saith unto them, Did ye never read in the scriptures, The stone which the builders rejected, the same is become the head of the corner: this is the Lord's doing, and it is marvellous in our eyes?  \verse Therefore say I unto you, The kingdom of God shall be taken from you, and given to a nation bringing forth the fruits thereof.

\verse And whosoever shall fall on this stone shall be broken: but on whomsoever it shall fall, it will grind him to powder.

\verse And when the chief priests and Pharisees had heard his parables, they perceived that he spake of them.

\verse But when they sought to lay hands on him, they feared the multitude, because they took him for a prophet.


\chapter
\verse And Jesus answered and spake unto them again by parables, and said, \verse The kingdom of heaven is like unto a certain king, which made a marriage for his son, \verse And sent forth his servants to call them that were bidden to the wedding: and they would not come.

\verse Again, he sent forth other servants, saying, Tell them which are bidden, Behold, I have prepared my dinner: my oxen and my fatlings are killed, and all things are ready: come unto the marriage.

\verse But they made light of it, and went their ways, one to his farm, another to his merchandise: \verse And the remnant took his servants, and entreated them spitefully, and slew them.

\verse But when the king heard thereof, he was wroth: and he sent forth his armies, and destroyed those murderers, and burned up their city.

\verse Then saith he to his servants, The wedding is ready, but they which were bidden were not worthy.

\verse Go ye therefore into the highways, and as many as ye shall find, bid to the marriage.

\verse So those servants went out into the highways, and gathered together all as many as they found, both bad and good: and the wedding was furnished with guests.

\verse And when the king came in to see the guests, he saw there a man which had not on a wedding garment: \verse And he saith unto him, Friend, how camest thou in hither not having a wedding garment? And he was speechless.

\verse Then said the king to the servants, Bind him hand and foot, and take him away, and cast him into outer darkness, there shall be weeping and gnashing of teeth.

\verse For many are called, but few are chosen.

\verse Then went the Pharisees, and took counsel how they might entangle him in his talk.

\verse And they sent out unto him their disciples with the Herodians, saying, Master, we know that thou art true, and teachest the way of God in truth, neither carest thou for any man: for thou regardest not the person of men.

\verse Tell us therefore, What thinkest thou? Is it lawful to give tribute unto Caesar, or not?  \verse But Jesus perceived their wickedness, and said, Why tempt ye me, ye hypocrites?  \verse Shew me the tribute money. And they brought unto him a penny.

\verse And he saith unto them, Whose is this image and superscription?  \verse They say unto him, Caesar's. Then saith he unto them, Render therefore unto Caesar the things which are Caesar's; and unto God the things that are God's.

\verse When they had heard these words, they marvelled, and left him, and went their way.

\verse The same day came to him the Sadducees, which say that there is no resurrection, and asked him, \verse Saying, Master, Moses said, If a man die, having no children, his brother shall marry his wife, and raise up seed unto his brother.

\verse Now there were with us seven brethren: and the first, when he had married a wife, deceased, and, having no issue, left his wife unto his brother: \verse Likewise the second also, and the third, unto the seventh.

\verse And last of all the woman died also.

\verse Therefore in the resurrection whose wife shall she be of the seven?  for they all had her.

\verse Jesus answered and said unto them, Ye do err, not knowing the scriptures, nor the power of God.

\verse For in the resurrection they neither marry, nor are given in marriage, but are as the angels of God in heaven.

\verse But as touching the resurrection of the dead, have ye not read that which was spoken unto you by God, saying, \verse I am the God of Abraham, and the God of Isaac, and the God of Jacob?  God is not the God of the dead, but of the living.

\verse And when the multitude heard this, they were astonished at his doctrine.

\verse But when the Pharisees had heard that he had put the Sadducees to silence, they were gathered together.

\verse Then one of them, which was a lawyer, asked him a question, tempting him, and saying, \verse Master, which is the great commandment in the law?  \verse Jesus said unto him, Thou shalt love the Lord thy God with all thy heart, and with all thy soul, and with all thy mind.

\verse This is the first and great commandment.

\verse And the second is like unto it, Thou shalt love thy neighbour as thyself.

\verse On these two commandments hang all the law and the prophets.

\verse While the Pharisees were gathered together, Jesus asked them, \verse Saying, What think ye of Christ? whose son is he? They say unto him, The son of David.

\verse He saith unto them, How then doth David in spirit call him Lord, saying, \verse The \LORD said unto my Lord, Sit thou on my right hand, till I make thine enemies thy footstool?  \verse If David then call him Lord, how is he his son?  \verse And no man was able to answer him a word, neither durst any man from that day forth ask him any more questions.


\chapter
\verse Then spake Jesus to the multitude, and to his disciples, \verse Saying The scribes and the Pharisees sit in Moses' seat: \verse All therefore whatsoever they bid you observe, that observe and do; but do not ye after their works: for they say, and do not.

\verse For they bind heavy burdens and grievous to be borne, and lay them on men's shoulders; but they themselves will not move them with one of their fingers.

\verse But all their works they do for to be seen of men: they make broad their phylacteries, and enlarge the borders of their garments, \verse And love the uppermost rooms at feasts, and the chief seats in the synagogues, \verse And greetings in the markets, and to be called of men, Rabbi, Rabbi.

\verse But be not ye called Rabbi: for one is your Master, even Christ; and all ye are brethren.

\verse And call no man your father upon the earth: for one is your Father, which is in heaven.

\verse Neither be ye called masters: for one is your Master, even Christ.

\verse But he that is greatest among you shall be your servant.

\verse And whosoever shall exalt himself shall be abased; and he that shall humble himself shall be exalted.

\verse But woe unto you, scribes and Pharisees, hypocrites! for ye shut up the kingdom of heaven against men: for ye neither go in yourselves, neither suffer ye them that are entering to go in.

\verse Woe unto you, scribes and Pharisees, hypocrites! for ye devour widows' houses, and for a pretence make long prayer: therefore ye shall receive the greater damnation.

\verse Woe unto you, scribes and Pharisees, hypocrites! for ye compass sea and land to make one proselyte, and when he is made, ye make him twofold more the child of hell than yourselves.

\verse Woe unto you, ye blind guides, which say, Whosoever shall swear by the temple, it is nothing; but whosoever shall swear by the gold of the temple, he is a debtor!  \verse Ye fools and blind: for whether is greater, the gold, or the temple that sanctifieth the gold?  \verse And, Whosoever shall swear by the altar, it is nothing; but whosoever sweareth by the gift that is upon it, he is guilty.

\verse Ye fools and blind: for whether is greater, the gift, or the altar that sanctifieth the gift?  \verse Whoso therefore shall swear by the altar, sweareth by it, and by all things thereon.

\verse And whoso shall swear by the temple, sweareth by it, and by him that dwelleth therein.

\verse And he that shall swear by heaven, sweareth by the throne of God, and by him that sitteth thereon.

\verse Woe unto you, scribes and Pharisees, hypocrites! for ye pay tithe of mint and anise and cummin, and have omitted the weightier matters of the law, judgment, mercy, and faith: these ought ye to have done, and not to leave the other undone.

\verse Ye blind guides, which strain at a gnat, and swallow a camel.

\verse Woe unto you, scribes and Pharisees, hypocrites! for ye make clean the outside of the cup and of the platter, but within they are full of extortion and excess.

\verse Thou blind Pharisee, cleanse first that which is within the cup and platter, that the outside of them may be clean also.

\verse Woe unto you, scribes and Pharisees, hypocrites! for ye are like unto whited sepulchres, which indeed appear beautiful outward, but are within full of dead men's bones, and of all uncleanness.

\verse Even so ye also outwardly appear righteous unto men, but within ye are full of hypocrisy and iniquity.

\verse Woe unto you, scribes and Pharisees, hypocrites! because ye build the tombs of the prophets, and garnish the sepulchres of the righteous, \verse And say, If we had been in the days of our fathers, we would not have been partakers with them in the blood of the prophets.

\verse Wherefore ye be witnesses unto yourselves, that ye are the children of them which killed the prophets.

\verse Fill ye up then the measure of your fathers.

\verse Ye serpents, ye generation of vipers, how can ye escape the damnation of hell?  \verse Wherefore, behold, I send unto you prophets, and wise men, and scribes: and some of them ye shall kill and crucify; and some of them shall ye scourge in your synagogues, and persecute them from city to city: \verse That upon you may come all the righteous blood shed upon the earth, from the blood of righteous Abel unto the blood of Zacharias son of Barachias, whom ye slew between the temple and the altar.

\verse Verily I say unto you, All these things shall come upon this generation.

\verse O Jerusalem, Jerusalem, thou that killest the prophets, and stonest them which are sent unto thee, how often would I have gathered thy children together, even as a hen gathereth her chickens under her wings, and ye would not!  \verse Behold, your house is left unto you desolate.

\verse For I say unto you, Ye shall not see me henceforth, till ye shall say, Blessed is he that cometh in the name of the Lord.


\chapter
\verse And Jesus went out, and departed from the temple: and his disciples came to him for to shew him the buildings of the temple.

\verse And Jesus said unto them, See ye not all these things? verily I say unto you, There shall not be left here one stone upon another, that shall not be thrown down.

\verse And as he sat upon the mount of Olives, the disciples came unto him privately, saying, Tell us, when shall these things be? and what shall be the sign of thy coming, and of the end of the world?  \verse And Jesus answered and said unto them, Take heed that no man deceive you.

\verse For many shall come in my name, saying, I am Christ; and shall deceive many.

\verse And ye shall hear of wars and rumours of wars: see that ye be not troubled: for all these things must come to pass, but the end is not yet.

\verse For nation shall rise against nation, and kingdom against kingdom: and there shall be famines, and pestilences, and earthquakes, in divers places.

\verse All these are the beginning of sorrows.

\verse Then shall they deliver you up to be afflicted, and shall kill you: and ye shall be hated of all nations for my name's sake.

\verse And then shall many be offended, and shall betray one another, and shall hate one another.

\verse And many false prophets shall rise, and shall deceive many.

\verse And because iniquity shall abound, the love of many shall wax cold.

\verse But he that shall endure unto the end, the same shall be saved.

\verse And this gospel of the kingdom shall be preached in all the world for a witness unto all nations; and then shall the end come.

\verse When ye therefore shall see the abomination of desolation, spoken of by Daniel the prophet, stand in the holy place, (whoso readeth, let him understand:) \verse Then let them which be in Judaea flee into the mountains: \verse Let him which is on the housetop not come down to take any thing out of his house: \verse Neither let him which is in the field return back to take his clothes.

\verse And woe unto them that are with child, and to them that give suck in those days!  \verse But pray ye that your flight be not in the winter, neither on the sabbath day: \verse For then shall be great tribulation, such as was not since the beginning of the world to this time, no, nor ever shall be.

\verse And except those days should be shortened, there should no flesh be saved: but for the elect's sake those days shall be shortened.

\verse Then if any man shall say unto you, Lo, here is Christ, or there; believe it not.

\verse For there shall arise false Christs, and false prophets, and shall shew great signs and wonders; insomuch that, if it were possible, they shall deceive the very elect.

\verse Behold, I have told you before.

\verse Wherefore if they shall say unto you, Behold, he is in the desert; go not forth: behold, he is in the secret chambers; believe it not.

\verse For as the lightning cometh out of the east, and shineth even unto the west; so shall also the coming of the Son of man be.

\verse For wheresoever the carcase is, there will the eagles be gathered together.

\verse Immediately after the tribulation of those days shall the sun be darkened, and the moon shall not give her light, and the stars shall fall from heaven, and the powers of the heavens shall be shaken: \verse And then shall appear the sign of the Son of man in heaven: and then shall all the tribes of the earth mourn, and they shall see the Son of man coming in the clouds of heaven with power and great glory.

\verse And he shall send his angels with a great sound of a trumpet, and they shall gather together his elect from the four winds, from one end of heaven to the other.

\verse Now learn a parable of the fig tree; When his branch is yet tender, and putteth forth leaves, ye know that summer is nigh: \verse So likewise ye, when ye shall see all these things, know that it is near, even at the doors.

\verse Verily I say unto you, This generation shall not pass, till all these things be fulfilled.

\verse Heaven and earth shall pass away, but my words shall not pass away.

\verse But of that day and hour knoweth no man, no, not the angels of heaven, but my Father only.

\verse But as the days of Noe were, so shall also the coming of the Son of man be.

\verse For as in the days that were before the flood they were eating and drinking, marrying and giving in marriage, until the day that Noe entered into the ark, \verse And knew not until the flood came, and took them all away; so shall also the coming of the Son of man be.

\verse Then shall two be in the field; the one shall be taken, and the other left.

\verse Two women shall be grinding at the mill; the one shall be taken, and the other left.

\verse Watch therefore: for ye know not what hour your Lord doth come.

\verse But know this, that if the goodman of the house had known in what watch the thief would come, he would have watched, and would not have suffered his house to be broken up.

\verse Therefore be ye also ready: for in such an hour as ye think not the Son of man cometh.

\verse Who then is a faithful and wise servant, whom his lord hath made ruler over his household, to give them meat in due season?  \verse Blessed is that servant, whom his lord when he cometh shall find so doing.

\verse Verily I say unto you, That he shall make him ruler over all his goods.

\verse But and if that evil servant shall say in his heart, My lord delayeth his coming; \verse And shall begin to smite his fellowservants, and to eat and drink with the drunken; \verse The lord of that servant shall come in a day when he looketh not for him, and in an hour that he is not aware of, \verse And shall cut him asunder, and appoint him his portion with the hypocrites: there shall be weeping and gnashing of teeth.


\chapter
\verse Then shall the kingdom of heaven be likened unto ten virgins, which took their lamps, and went forth to meet the bridegroom.

\verse And five of them were wise, and five were foolish.

\verse They that were foolish took their lamps, and took no oil with them: \verse But the wise took oil in their vessels with their lamps.

\verse While the bridegroom tarried, they all slumbered and slept.

\verse And at midnight there was a cry made, Behold, the bridegroom cometh; go ye out to meet him.

\verse Then all those virgins arose, and trimmed their lamps.

\verse And the foolish said unto the wise, Give us of your oil; for our lamps are gone out.

\verse But the wise answered, saying, Not so; lest there be not enough for us and you: but go ye rather to them that sell, and buy for yourselves.

\verse And while they went to buy, the bridegroom came; and they that were ready went in with him to the marriage: and the door was shut.

\verse Afterward came also the other virgins, saying, Lord, Lord, open to us.

\verse But he answered and said, Verily I say unto you, I know you not.

\verse Watch therefore, for ye know neither the day nor the hour wherein the Son of man cometh.

\verse For the kingdom of heaven is as a man travelling into a far country, who called his own servants, and delivered unto them his goods.

\verse And unto one he gave five talents, to another two, and to another one; to every man according to his several ability; and straightway took his journey.

\verse Then he that had received the five talents went and traded with the same, and made them other five talents.

\verse And likewise he that had received two, he also gained other two.

\verse But he that had received one went and digged in the earth, and hid his lord's money.

\verse After a long time the lord of those servants cometh, and reckoneth with them.

\verse And so he that had received five talents came and brought other five talents, saying, Lord, thou deliveredst unto me five talents: behold, I have gained beside them five talents more.

\verse His lord said unto him, Well done, thou good and faithful servant: thou hast been faithful over a few things, I will make thee ruler over many things: enter thou into the joy of thy lord.

\verse He also that had received two talents came and said, Lord, thou deliveredst unto me two talents: behold, I have gained two other talents beside them.

\verse His lord said unto him, Well done, good and faithful servant; thou hast been faithful over a few things, I will make thee ruler over many things: enter thou into the joy of thy lord.

\verse Then he which had received the one talent came and said, Lord, I knew thee that thou art an hard man, reaping where thou hast not sown, and gathering where thou hast not strawed: \verse And I was afraid, and went and hid thy talent in the earth: lo, there thou hast that is thine.

\verse His lord answered and said unto him, Thou wicked and slothful servant, thou knewest that I reap where I sowed not, and gather where I have not strawed: \verse Thou oughtest therefore to have put my money to the exchangers, and then at my coming I should have received mine own with usury.

\verse Take therefore the talent from him, and give it unto him which hath ten talents.

\verse For unto every one that hath shall be given, and he shall have abundance: but from him that hath not shall be taken away even that which he hath.

\verse And cast ye the unprofitable servant into outer darkness: there shall be weeping and gnashing of teeth.

\verse When the Son of man shall come in his glory, and all the holy angels with him, then shall he sit upon the throne of his glory: \verse And before him shall be gathered all nations: and he shall separate them one from another, as a shepherd divideth his sheep from the goats: \verse And he shall set the sheep on his right hand, but the goats on the left.

\verse Then shall the King say unto them on his right hand, Come, ye blessed of my Father, inherit the kingdom prepared for you from the foundation of the world: \verse For I was an hungred, and ye gave me meat: I was thirsty, and ye gave me drink: I was a stranger, and ye took me in: \verse Naked, and ye clothed me: I was sick, and ye visited me: I was in prison, and ye came unto me.

\verse Then shall the righteous answer him, saying, Lord, when saw we thee an hungred, and fed thee? or thirsty, and gave thee drink?  \verse When saw we thee a stranger, and took thee in? or naked, and clothed thee?  \verse Or when saw we thee sick, or in prison, and came unto thee?  \verse And the King shall answer and say unto them, Verily I say unto you, Inasmuch as ye have done it unto one of the least of these my brethren, ye have done it unto me.

\verse Then shall he say also unto them on the left hand, Depart from me, ye cursed, into everlasting fire, prepared for the devil and his angels: \verse For I was an hungred, and ye gave me no meat: I was thirsty, and ye gave me no drink: \verse I was a stranger, and ye took me not in: naked, and ye clothed me not: sick, and in prison, and ye visited me not.

\verse Then shall they also answer him, saying, Lord, when saw we thee an hungred, or athirst, or a stranger, or naked, or sick, or in prison, and did not minister unto thee?  \verse Then shall he answer them, saying, Verily I say unto you, Inasmuch as ye did it not to one of the least of these, ye did it not to me.

\verse And these shall go away into everlasting punishment: but the righteous into life eternal.


\chapter
\verse And it came to pass, when Jesus had finished all these sayings, he said unto his disciples, \verse Ye know that after two days is the feast of the passover, and the Son of man is betrayed to be crucified.

\verse Then assembled together the chief priests, and the scribes, and the elders of the people, unto the palace of the high priest, who was called Caiaphas, \verse And consulted that they might take Jesus by subtilty, and kill him.

\verse But they said, Not on the feast day, lest there be an uproar among the people.

\verse Now when Jesus was in Bethany, in the house of Simon the leper, \verse There came unto him a woman having an alabaster box of very precious ointment, and poured it on his head, as he sat at meat.

\verse But when his disciples saw it, they had indignation, saying, To what purpose is this waste?  \verse For this ointment might have been sold for much, and given to the poor.

\verse When Jesus understood it, he said unto them, Why trouble ye the woman? for she hath wrought a good work upon me.

\verse For ye have the poor always with you; but me ye have not always.

\verse For in that she hath poured this ointment on my body, she did it for my burial.

\verse Verily I say unto you, Wheresoever this gospel shall be preached in the whole world, there shall also this, that this woman hath done, be told for a memorial of her.

\verse Then one of the twelve, called Judas Iscariot, went unto the chief priests, \verse And said unto them, What will ye give me, and I will deliver him unto you? And they covenanted with him for thirty pieces of silver.

\verse And from that time he sought opportunity to betray him.

\verse Now the first day of the feast of unleavened bread the disciples came to Jesus, saying unto him, Where wilt thou that we prepare for thee to eat the passover?  \verse And he said, Go into the city to such a man, and say unto him, The Master saith, My time is at hand; I will keep the passover at thy house with my disciples.

\verse And the disciples did as Jesus had appointed them; and they made ready the passover.

\verse Now when the even was come, he sat down with the twelve.

\verse And as they did eat, he said, Verily I say unto you, that one of you shall betray me.

\verse And they were exceeding sorrowful, and began every one of them to say unto him, Lord, is it I?  \verse And he answered and said, He that dippeth his hand with me in the dish, the same shall betray me.

\verse The Son of man goeth as it is written of him: but woe unto that man by whom the Son of man is betrayed! it had been good for that man if he had not been born.

\verse Then Judas, which betrayed him, answered and said, Master, is it I?  He said unto him, Thou hast said.

\verse And as they were eating, Jesus took bread, and blessed it, and brake it, and gave it to the disciples, and said, Take, eat; this is my body.

\verse And he took the cup, and gave thanks, and gave it to them, saying, Drink ye all of it; \verse For this is my blood of the new testament, which is shed for many for the remission of sins.

\verse But I say unto you, I will not drink henceforth of this fruit of the vine, until that day when I drink it new with you in my Father's kingdom.

\verse And when they had sung an hymn, they went out into the mount of Olives.

\verse Then saith Jesus unto them, All ye shall be offended because of me this night: for it is written, I will smite the shepherd, and the sheep of the flock shall be scattered abroad.

\verse But after I am risen again, I will go before you into Galilee.

\verse Peter answered and said unto him, Though all men shall be offended because of thee, yet will I never be offended.

\verse Jesus said unto him, Verily I say unto thee, That this night, before the cock crow, thou shalt deny me thrice.

\verse Peter said unto him, Though I should die with thee, yet will I not deny thee. Likewise also said all the disciples.

\verse Then cometh Jesus with them unto a place called Gethsemane, and saith unto the disciples, Sit ye here, while I go and pray yonder.

\verse And he took with him Peter and the two sons of Zebedee, and began to be sorrowful and very heavy.

\verse Then saith he unto them, My soul is exceeding sorrowful, even unto death: tarry ye here, and watch with me.

\verse And he went a little farther, and fell on his face, and prayed, saying, O my Father, if it be possible, let this cup pass from me: nevertheless not as I will, but as thou wilt.

\verse And he cometh unto the disciples, and findeth them asleep, and saith unto Peter, What, could ye not watch with me one hour?  \verse Watch and pray, that ye enter not into temptation: the spirit indeed is willing, but the flesh is weak.

\verse He went away again the second time, and prayed, saying, O my Father, if this cup may not pass away from me, except I drink it, thy will be done.

\verse And he came and found them asleep again: for their eyes were heavy.

\verse And he left them, and went away again, and prayed the third time, saying the same words.

\verse Then cometh he to his disciples, and saith unto them, Sleep on now, and take your rest: behold, the hour is at hand, and the Son of man is betrayed into the hands of sinners.

\verse Rise, let us be going: behold, he is at hand that doth betray me.

\verse And while he yet spake, lo, Judas, one of the twelve, came, and with him a great multitude with swords and staves, from the chief priests and elders of the people.

\verse Now he that betrayed him gave them a sign, saying, Whomsoever I shall kiss, that same is he: hold him fast.

\verse And forthwith he came to Jesus, and said, Hail, master; and kissed him.

\verse And Jesus said unto him, Friend, wherefore art thou come? Then came they, and laid hands on Jesus and took him.

\verse And, behold, one of them which were with Jesus stretched out his hand, and drew his sword, and struck a servant of the high priest's, and smote off his ear.

\verse Then said Jesus unto him, Put up again thy sword into his place: for all they that take the sword shall perish with the sword.

\verse Thinkest thou that I cannot now pray to my Father, and he shall presently give me more than twelve legions of angels?  \verse But how then shall the scriptures be fulfilled, that thus it must be?  \verse In that same hour said Jesus to the multitudes, Are ye come out as against a thief with swords and staves for to take me? I sat daily with you teaching in the temple, and ye laid no hold on me.

\verse But all this was done, that the scriptures of the prophets might be fulfilled. Then all the disciples forsook him, and fled.

\verse And they that had laid hold on Jesus led him away to Caiaphas the high priest, where the scribes and the elders were assembled.

\verse But Peter followed him afar off unto the high priest's palace, and went in, and sat with the servants, to see the end.

\verse Now the chief priests, and elders, and all the council, sought false witness against Jesus, to put him to death; \verse But found none: yea, though many false witnesses came, yet found they none. At the last came two false witnesses, \verse And said, This fellow said, I am able to destroy the temple of God, and to build it in three days.

\verse And the high priest arose, and said unto him, Answerest thou nothing? what is it which these witness against thee?  \verse But Jesus held his peace, And the high priest answered and said unto him, I adjure thee by the living God, that thou tell us whether thou be the Christ, the Son of God.

\verse Jesus saith unto him, Thou hast said: nevertheless I say unto you, Hereafter shall ye see the Son of man sitting on the right hand of power, and coming in the clouds of heaven.

\verse Then the high priest rent his clothes, saying, He hath spoken blasphemy; what further need have we of witnesses? behold, now ye have heard his blasphemy.

\verse What think ye? They answered and said, He is guilty of death.

\verse Then did they spit in his face, and buffeted him; and others smote him with the palms of their hands, \verse Saying, Prophesy unto us, thou Christ, Who is he that smote thee?  \verse Now Peter sat without in the palace: and a damsel came unto him, saying, Thou also wast with Jesus of Galilee.

\verse But he denied before them all, saying, I know not what thou sayest.

\verse And when he was gone out into the porch, another maid saw him, and said unto them that were there, This fellow was also with Jesus of Nazareth.

\verse And again he denied with an oath, I do not know the man.

\verse And after a while came unto him they that stood by, and said to Peter, Surely thou also art one of them; for thy speech bewrayeth thee.

\verse Then began he to curse and to swear, saying, I know not the man.  And immediately the cock crew.

\verse And Peter remembered the word of Jesus, which said unto him, Before the cock crow, thou shalt deny me thrice. And he went out, and wept bitterly.


\chapter
\verse When the morning was come, all the chief priests and elders of the people took counsel against Jesus to put him to death: \verse And when they had bound him, they led him away, and delivered him to Pontius Pilate the governor.

\verse Then Judas, which had betrayed him, when he saw that he was condemned, repented himself, and brought again the thirty pieces of silver to the chief priests and elders, \verse Saying, I have sinned in that I have betrayed the innocent blood. And they said, What is that to us? see thou to that.

\verse And he cast down the pieces of silver in the temple, and departed, and went and hanged himself.

\verse And the chief priests took the silver pieces, and said, It is not lawful for to put them into the treasury, because it is the price of blood.

\verse And they took counsel, and bought with them the potter's field, to bury strangers in.

\verse Wherefore that field was called, The field of blood, unto this day.

\verse Then was fulfilled that which was spoken by Jeremy the prophet, saying, And they took the thirty pieces of silver, the price of him that was valued, whom they of the children of Israel did value; \verse And gave them for the potter's field, as the Lord appointed me.

\verse And Jesus stood before the governor: and the governor asked him, saying, Art thou the King of the Jews? And Jesus said unto him, Thou sayest.

\verse And when he was accused of the chief priests and elders, he answered nothing.

\verse Then said Pilate unto him, Hearest thou not how many things they witness against thee?  \verse And he answered him to never a word; insomuch that the governor marvelled greatly.

\verse Now at that feast the governor was wont to release unto the people a prisoner, whom they would.

\verse And they had then a notable prisoner, called Barabbas.

\verse Therefore when they were gathered together, Pilate said unto them, Whom will ye that I release unto you? Barabbas, or Jesus which is called Christ?  \verse For he knew that for envy they had delivered him.

\verse When he was set down on the judgment seat, his wife sent unto him, saying, Have thou nothing to do with that just man: for I have suffered many things this day in a dream because of him.

\verse But the chief priests and elders persuaded the multitude that they should ask Barabbas, and destroy Jesus.

\verse The governor answered and said unto them, Whether of the twain will ye that I release unto you? They said, Barabbas.

\verse Pilate saith unto them, What shall I do then with Jesus which is called Christ? They all say unto him, Let him be crucified.

\verse And the governor said, Why, what evil hath he done? But they cried out the more, saying, Let him be crucified.

\verse When Pilate saw that he could prevail nothing, but that rather a tumult was made, he took water, and washed his hands before the multitude, saying, I am innocent of the blood of this just person: see ye to it.

\verse Then answered all the people, and said, His blood be on us, and on our children.

\verse Then released he Barabbas unto them: and when he had scourged Jesus, he delivered him to be crucified.

\verse Then the soldiers of the governor took Jesus into the common hall, and gathered unto him the whole band of soldiers.

\verse And they stripped him, and put on him a scarlet robe.

\verse And when they had platted a crown of thorns, they put it upon his head, and a reed in his right hand: and they bowed the knee before him, and mocked him, saying, Hail, King of the Jews!  \verse And they spit upon him, and took the reed, and smote him on the head.

\verse And after that they had mocked him, they took the robe off from him, and put his own raiment on him, and led him away to crucify him.

\verse And as they came out, they found a man of Cyrene, Simon by name: him they compelled to bear his cross.

\verse And when they were come unto a place called Golgotha, that is to say, a place of a skull, \verse They gave him vinegar to drink mingled with gall: and when he had tasted thereof, he would not drink.

\verse And they crucified him, and parted his garments, casting lots: that it might be fulfilled which was spoken by the prophet, They parted my garments among them, and upon my vesture did they cast lots.

\verse And sitting down they watched him there; \verse And set up over his head his accusation written, THIS IS JESUS THE KING OF THE JEWS.

\verse Then were there two thieves crucified with him, one on the right hand, and another on the left.

\verse And they that passed by reviled him, wagging their heads, \verse And saying, Thou that destroyest the temple, and buildest it in three days, save thyself. If thou be the Son of God, come down from the cross.

\verse Likewise also the chief priests mocking him, with the scribes and elders, said, \verse He saved others; himself he cannot save. If he be the King of Israel, let him now come down from the cross, and we will believe him.

\verse He trusted in God; let him deliver him now, if he will have him: for he said, I am the Son of God.

\verse The thieves also, which were crucified with him, cast the same in his teeth.

\verse Now from the sixth hour there was darkness over all the land unto the ninth hour.

\verse And about the ninth hour Jesus cried with a loud voice, saying, Eli, Eli, lama sabachthani? that is to say, My God, my God, why hast thou forsaken me?  \verse Some of them that stood there, when they heard that, said, This man calleth for Elias.

\verse And straightway one of them ran, and took a spunge, and filled it with vinegar, and put it on a reed, and gave him to drink.

\verse The rest said, Let be, let us see whether Elias will come to save him.

\verse Jesus, when he had cried again with a loud voice, yielded up the ghost.

\verse And, behold, the veil of the temple was rent in twain from the top to the bottom; and the earth did quake, and the rocks rent; \verse And the graves were opened; and many bodies of the saints which slept arose, \verse And came out of the graves after his resurrection, and went into the holy city, and appeared unto many.

\verse Now when the centurion, and they that were with him, watching Jesus, saw the earthquake, and those things that were done, they feared greatly, saying, Truly this was the Son of God.

\verse And many women were there beholding afar off, which followed Jesus from Galilee, ministering unto him: \verse Among which was Mary Magdalene, and Mary the mother of James and Joses, and the mother of Zebedees children.

\verse When the even was come, there came a rich man of Arimathaea, named Joseph, who also himself was Jesus' disciple: \verse He went to Pilate, and begged the body of Jesus. Then Pilate commanded the body to be delivered.

\verse And when Joseph had taken the body, he wrapped it in a clean linen cloth, \verse And laid it in his own new tomb, which he had hewn out in the rock: and he rolled a great stone to the door of the sepulchre, and departed.

\verse And there was Mary Magdalene, and the other Mary, sitting over against the sepulchre.

\verse Now the next day, that followed the day of the preparation, the chief priests and Pharisees came together unto Pilate, \verse Saying, Sir, we remember that that deceiver said, while he was yet alive, After three days I will rise again.

\verse Command therefore that the sepulchre be made sure until the third day, lest his disciples come by night, and steal him away, and say unto the people, He is risen from the dead: so the last error shall be worse than the first.

\verse Pilate said unto them, Ye have a watch: go your way, make it as sure as ye can.

\verse So they went, and made the sepulchre sure, sealing the stone, and setting a watch.


\chapter
\verse In the end of the sabbath, as it began to dawn toward the first day of the week, came Mary Magdalene and the other Mary to see the sepulchre.

\verse And, behold, there was a great earthquake: for the angel of the Lord descended from heaven, and came and rolled back the stone from the door, and sat upon it.

\verse His countenance was like lightning, and his raiment white as snow: \verse And for fear of him the keepers did shake, and became as dead men.

\verse And the angel answered and said unto the women, Fear not ye: for I know that ye seek Jesus, which was crucified.

\verse He is not here: for he is risen, as he said. Come, see the place where the Lord lay.

\verse And go quickly, and tell his disciples that he is risen from the dead; and, behold, he goeth before you into Galilee; there shall ye see him: lo, I have told you.

\verse And they departed quickly from the sepulchre with fear and great joy; and did run to bring his disciples word.

\verse And as they went to tell his disciples, behold, Jesus met them, saying, All hail. And they came and held him by the feet, and worshipped him.

\verse Then said Jesus unto them, Be not afraid: go tell my brethren that they go into Galilee, and there shall they see me.

\verse Now when they were going, behold, some of the watch came into the city, and shewed unto the chief priests all the things that were done.

\verse And when they were assembled with the elders, and had taken counsel, they gave large money unto the soldiers, \verse Saying, Say ye, His disciples came by night, and stole him away while we slept.

\verse And if this come to the governor's ears, we will persuade him, and secure you.

\verse So they took the money, and did as they were taught: and this saying is commonly reported among the Jews until this day.

\verse Then the eleven disciples went away into Galilee, into a mountain where Jesus had appointed them.

\verse And when they saw him, they worshipped him: but some doubted.

\verse And Jesus came and spake unto them, saying, All power is given unto me in heaven and in earth.

\verse Go ye therefore, and teach all nations, baptizing them in the name of the Father, and of the Son, and of the Holy Ghost: \verse Teaching them to observe all things whatsoever I have commanded you: and, lo, I am with you alway, even unto the end of the world. Amen.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "bible.tex"
%%% End: 
