\Chapter
\Verse Paul, a prisoner of Jesus Christ, and Timothy our brother, unto Philemon our dearly beloved, and fellowlabourer, \Verse And to our beloved Apphia, and Archippus our fellowsoldier, and to the church in thy house: \Verse Grace to you, and peace, from God our Father and the Lord Jesus Christ.

\Verse I thank my God, making mention of thee always in my prayers, \Verse Hearing of thy love and faith, which thou hast toward the Lord Jesus, and toward all saints; \Verse That the communication of thy faith may become effectual by the acknowledging of every good thing which is in you in Christ Jesus.

\Verse For we have great joy and consolation in thy love, because the bowels of the saints are refreshed by thee, brother.

\Verse Wherefore, though I might be much bold in Christ to enjoin thee that which is convenient, \Verse Yet for love's sake I rather beseech thee, being such an one as Paul the aged, and now also a prisoner of Jesus Christ.

\Verse I beseech thee for my son Onesimus, whom I have begotten in my bonds: \Verse Which in time past was to thee unprofitable, but now profitable to thee and to me: \Verse Whom I have sent again: thou therefore receive him, that is, mine own bowels: \Verse Whom I would have retained with me, that in thy stead he might have ministered unto me in the bonds of the gospel: \Verse But without thy mind would I do nothing; that thy benefit should not be as it were of necessity, but willingly.

\Verse For perhaps he therefore departed for a season, that thou shouldest receive him for ever; \Verse Not now as a servant, but above a servant, a brother beloved, specially to me, but how much more unto thee, both in the flesh, and in the Lord?  \Verse If thou count me therefore a partner, receive him as myself.

\Verse If he hath wronged thee, or oweth thee ought, put that on mine account; \Verse I Paul have written it with mine own hand, I will repay it: albeit I do not say to thee how thou owest unto me even thine own self besides.

\Verse Yea, brother, let me have joy of thee in the Lord: refresh my bowels in the Lord.

\Verse Having confidence in thy obedience I wrote unto thee, knowing that thou wilt also do more than I say.

\Verse But withal prepare me also a lodging: for I trust that through your prayers I shall be given unto you.

\Verse There salute thee Epaphras, my fellowprisoner in Christ Jesus; \Verse Marcus, Aristarchus, Demas, Lucas, my fellowlabourers.

\Verse The grace of our Lord Jesus Christ be with your spirit. Amen.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "bible.tex"
%%% End: 
