\book{Job}



\chapter
\verse There was a man in the land of Uz, whose name was Job; and that man was perfect and upright, and one that feared God, and eschewed evil.

\verse And there were born unto him seven sons and three daughters.

\verse His substance also was seven thousand sheep, and three thousand camels, and five hundred yoke of oxen, and five hundred she asses, and a very great household; so that this man was the greatest of all the men of the east.

\verse And his sons went and feasted in their houses, every one his day; and sent and called for their three sisters to eat and to drink with them.

\verse And it was so, when the days of their feasting were gone about, that Job sent and sanctified them, and rose up early in the morning, and offered burnt offerings according to the number of them all: for Job said, It may be that my sons have sinned, and cursed God in their hearts. Thus did Job continually.

\verse Now there was a day when the sons of God came to present themselves before the \LORD, and Satan came also among them.

\verse And the \LORD said unto Satan, Whence comest thou? Then Satan answered the \LORD, and said, From going to and fro in the earth, and from walking up and down in it.

\verse And the \LORD said unto Satan, Hast thou considered my servant Job, that there is none like him in the earth, a perfect and an upright man, one that feareth God, and escheweth evil?  \verse Then Satan answered the \LORD, and said, Doth Job fear God for nought?  \verse Hast not thou made an hedge about him, and about his house, and about all that he hath on every side? thou hast blessed the work of his hands, and his substance is increased in the land.

\verse But put forth thine hand now, and touch all that he hath, and he will curse thee to thy face.

\verse And the \LORD said unto Satan, Behold, all that he hath is in thy power; only upon himself put not forth thine hand. So Satan went forth from the presence of the \LORD.

\verse And there was a day when his sons and his daughters were eating and drinking wine in their eldest brother's house: \verse And there came a messenger unto Job, and said, The oxen were plowing, and the asses feeding beside them: \verse And the Sabeans fell upon them, and took them away; yea, they have slain the servants with the edge of the sword; and I only am escaped alone to tell thee.

\verse While he was yet speaking, there came also another, and said, The fire of God is fallen from heaven, and hath burned up the sheep, and the servants, and consumed them; and I only am escaped alone to tell thee.

\verse While he was yet speaking, there came also another, and said, The Chaldeans made out three bands, and fell upon the camels, and have carried them away, yea, and slain the servants with the edge of the sword; and I only am escaped alone to tell thee.

\verse While he was yet speaking, there came also another, and said, Thy sons and thy daughters were eating and drinking wine in their eldest brother's house: \verse And, behold, there came a great wind from the wilderness, and smote the four corners of the house, and it fell upon the young men, and they are dead; and I only am escaped alone to tell thee.

\verse Then Job arose, and rent his mantle, and shaved his head, and fell down upon the ground, and worshipped, \verse And said, Naked came I out of my mother's womb, and naked shall I return thither: the \LORD gave, and the \LORD hath taken away; blessed be the name of the \LORD.

\verse In all this Job sinned not, nor charged God foolishly.


\chapter
\verse Again there was a day when the sons of God came to present themselves before the \LORD, and Satan came also among them to present himself before the \LORD.

\verse And the \LORD said unto Satan, From whence comest thou? And Satan answered the \LORD, and said, From going to and fro in the earth, and from walking up and down in it.

\verse And the \LORD said unto Satan, Hast thou considered my servant Job, that there is none like him in the earth, a perfect and an upright man, one that feareth God, and escheweth evil? and still he holdeth fast his integrity, although thou movedst me against him, to destroy him without cause.

\verse And Satan answered the \LORD, and said, Skin for skin, yea, all that a man hath will he give for his life.

\verse But put forth thine hand now, and touch his bone and his flesh, and he will curse thee to thy face.

\verse And the \LORD said unto Satan, Behold, he is in thine hand; but save his life.

\verse So went Satan forth from the presence of the \LORD, and smote Job with sore boils from the sole of his foot unto his crown.

\verse And he took him a potsherd to scrape himself withal; and he sat down among the ashes.

\verse Then said his wife unto him, Dost thou still retain thine integrity?  curse God, and die.

\verse But he said unto her, Thou speakest as one of the foolish women speaketh. What? shall we receive good at the hand of God, and shall we not receive evil? In all this did not Job sin with his lips.

\verse Now when Job's three friends heard of all this evil that was come upon him, they came every one from his own place; Eliphaz the Temanite, and Bildad the Shuhite, and Zophar the Naamathite: for they had made an appointment together to come to mourn with him and to comfort him.

\verse And when they lifted up their eyes afar off, and knew him not, they lifted up their voice, and wept; and they rent every one his mantle, and sprinkled dust upon their heads toward heaven.

\verse So they sat down with him upon the ground seven days and seven nights, and none spake a word unto him: for they saw that his grief was very great.


\chapter
\verse After this opened Job his mouth, and cursed his day.

\verse And Job spake, and said, \verse Let the day perish wherein I was born, and the night in which it was said, There is a man child conceived.

\verse Let that day be darkness; let not God regard it from above, neither let the light shine upon it.

\verse Let darkness and the shadow of death stain it; let a cloud dwell upon it; let the blackness of the day terrify it.

\verse As for that night, let darkness seize upon it; let it not be joined unto the days of the year, let it not come into the number of the months.

\verse Lo, let that night be solitary, let no joyful voice come therein.

\verse Let them curse it that curse the day, who are ready to raise up their mourning.

\verse Let the stars of the twilight thereof be dark; let it look for light, but have none; neither let it see the dawning of the day: \verse Because it shut not up the doors of my mother's womb, nor hid sorrow from mine eyes.

\verse Why died I not from the womb? why did I not give up the ghost when I came out of the belly?  \verse Why did the knees prevent me? or why the breasts that I should suck?  \verse For now should I have lain still and been quiet, I should have slept: then had I been at rest, \verse With kings and counsellors of the earth, which build desolate places for themselves; \verse Or with princes that had gold, who filled their houses with silver: \verse Or as an hidden untimely birth I had not been; as infants which never saw light.

\verse There the wicked cease from troubling; and there the weary be at rest.

\verse There the prisoners rest together; they hear not the voice of the oppressor.

\verse The small and great are there; and the servant is free from his master.

\verse Wherefore is light given to him that is in misery, and life unto the bitter in soul; \verse Which long for death, but it cometh not; and dig for it more than for hid treasures; \verse Which rejoice exceedingly, and are glad, when they can find the grave?  \verse Why is light given to a man whose way is hid, and whom God hath hedged in?  \verse For my sighing cometh before I eat, and my roarings are poured out like the waters.

\verse For the thing which I greatly feared is come upon me, and that which I was afraid of is come unto me.

\verse I was not in safety, neither had I rest, neither was I quiet; yet trouble came.


\chapter
\verse Then Eliphaz the Temanite answered and said, \verse If we assay to commune with thee, wilt thou be grieved? but who can withhold himself from speaking?  \verse Behold, thou hast instructed many, and thou hast strengthened the weak hands.

\verse Thy words have upholden him that was falling, and thou hast strengthened the feeble knees.

\verse But now it is come upon thee, and thou faintest; it toucheth thee, and thou art troubled.

\verse Is not this thy fear, thy confidence, thy hope, and the uprightness of thy ways?  \verse Remember, I pray thee, who ever perished, being innocent? or where were the righteous cut off?  \verse Even as I have seen, they that plow iniquity, and sow wickedness, reap the same.

\verse By the blast of God they perish, and by the breath of his nostrils are they consumed.

\verse The roaring of the lion, and the voice of the fierce lion, and the teeth of the young lions, are broken.

\verse The old lion perisheth for lack of prey, and the stout lion's whelps are scattered abroad.

\verse Now a thing was secretly brought to me, and mine ear received a little thereof.

\verse In thoughts from the visions of the night, when deep sleep falleth on men, \verse Fear came upon me, and trembling, which made all my bones to shake.

\verse Then a spirit passed before my face; the hair of my flesh stood up: \verse It stood still, but I could not discern the form thereof: an image was before mine eyes, there was silence, and I heard a voice, saying, \verse Shall mortal man be more just than God? shall a man be more pure than his maker?  \verse Behold, he put no trust in his servants; and his angels he charged with folly: \verse How much less in them that dwell in houses of clay, whose foundation is in the dust, which are crushed before the moth?  \verse They are destroyed from morning to evening: they perish for ever without any regarding it.

\verse Doth not their excellency which is in them go away? they die, even without wisdom.


\chapter
\verse Call now, if there be any that will answer thee; and to which of the saints wilt thou turn?  \verse For wrath killeth the foolish man, and envy slayeth the silly one.

\verse I have seen the foolish taking root: but suddenly I cursed his habitation.

\verse His children are far from safety, and they are crushed in the gate, neither is there any to deliver them.

\verse Whose harvest the hungry eateth up, and taketh it even out of the thorns, and the robber swalloweth up their substance.

\verse Although affliction cometh not forth of the dust, neither doth trouble spring out of the ground; \verse Yet man is born unto trouble, as the sparks fly upward.

\verse I would seek unto God, and unto God would I commit my cause: \verse Which doeth great things and unsearchable; marvellous things without number: \verse Who giveth rain upon the earth, and sendeth waters upon the fields: \verse To set up on high those that be low; that those which mourn may be exalted to safety.

\verse He disappointeth the devices of the crafty, so that their hands cannot perform their enterprise.

\verse He taketh the wise in their own craftiness: and the counsel of the froward is carried headlong.

\verse They meet with darkness in the day time, and grope in the noonday as in the night.

\verse But he saveth the poor from the sword, from their mouth, and from the hand of the mighty.

\verse So the poor hath hope, and iniquity stoppeth her mouth.

\verse Behold, happy is the man whom God correcteth: therefore despise not thou the chastening of the Almighty: \verse For he maketh sore, and bindeth up: he woundeth, and his hands make whole.

\verse He shall deliver thee in six troubles: yea, in seven there shall no evil touch thee.

\verse In famine he shall redeem thee from death: and in war from the power of the sword.

\verse Thou shalt be hid from the scourge of the tongue: neither shalt thou be afraid of destruction when it cometh.

\verse At destruction and famine thou shalt laugh: neither shalt thou be afraid of the beasts of the earth.

\verse For thou shalt be in league with the stones of the field: and the beasts of the field shall be at peace with thee.

\verse And thou shalt know that thy tabernacle shall be in peace; and thou shalt visit thy habitation, and shalt not sin.

\verse Thou shalt know also that thy seed shall be great, and thine offspring as the grass of the earth.

\verse Thou shalt come to thy grave in a full age, like as a shock of corn cometh in in his season.

\verse Lo this, we have searched it, so it is; hear it, and know thou it for thy good.


\chapter
\verse But Job answered and said, \verse Oh that my grief were throughly weighed, and my calamity laid in the balances together!  \verse For now it would be heavier than the sand of the sea: therefore my words are swallowed up.

\verse For the arrows of the Almighty are within me, the poison whereof drinketh up my spirit: the terrors of God do set themselves in array against me.

\verse Doth the wild ass bray when he hath grass? or loweth the ox over his fodder?  \verse Can that which is unsavoury be eaten without salt? or is there any taste in the white of an egg?  \verse The things that my soul refused to touch are as my sorrowful meat.

\verse Oh that I might have my request; and that God would grant me the thing that I long for!  \verse Even that it would please God to destroy me; that he would let loose his hand, and cut me off!  \verse Then should I yet have comfort; yea, I would harden myself in sorrow: let him not spare; for I have not concealed the words of the Holy One.

\verse What is my strength, that I should hope? and what is mine end, that I should prolong my life?  \verse Is my strength the strength of stones? or is my flesh of brass?  \verse Is not my help in me? and is wisdom driven quite from me?  \verse To him that is afflicted pity should be shewed from his friend; but he forsaketh the fear of the Almighty.

\verse My brethren have dealt deceitfully as a brook, and as the stream of brooks they pass away; \verse Which are blackish by reason of the ice, and wherein the snow is hid: \verse What time they wax warm, they vanish: when it is hot, they are consumed out of their place.

\verse The paths of their way are turned aside; they go to nothing, and perish.

\verse The troops of Tema looked, the companies of Sheba waited for them.

\verse They were confounded because they had hoped; they came thither, and were ashamed.

\verse For now ye are nothing; ye see my casting down, and are afraid.

\verse Did I say, Bring unto me? or, Give a reward for me of your substance?  \verse Or, Deliver me from the enemy's hand? or, Redeem me from the hand of the mighty?  \verse Teach me, and I will hold my tongue: and cause me to understand wherein I have erred.

\verse How forcible are right words! but what doth your arguing reprove?  \verse Do ye imagine to reprove words, and the speeches of one that is desperate, which are as wind?  \verse Yea, ye overwhelm the fatherless, and ye dig a pit for your friend.

\verse Now therefore be content, look upon me; for it is evident unto you if I lie.

\verse Return, I pray you, let it not be iniquity; yea, return again, my righteousness is in it.

\verse Is there iniquity in my tongue? cannot my taste discern perverse things?  
\chapter
\verse Is there not an appointed time to man upon earth? are not his days also like the days of an hireling?  \verse As a servant earnestly desireth the shadow, and as an hireling looketh for the reward of his work: \verse So am I made to possess months of vanity, and wearisome nights are appointed to me.

\verse When I lie down, I say, When shall I arise, and the night be gone?  and I am full of tossings to and fro unto the dawning of the day.

\verse My flesh is clothed with worms and clods of dust; my skin is broken, and become loathsome.

\verse My days are swifter than a weaver's shuttle, and are spent without hope.

\verse O remember that my life is wind: mine eye shall no more see good.

\verse The eye of him that hath seen me shall see me no more: thine eyes are upon me, and I am not.

\verse As the cloud is consumed and vanisheth away: so he that goeth down to the grave shall come up no more.

\verse He shall return no more to his house, neither shall his place know him any more.

\verse Therefore I will not refrain my mouth; I will speak in the anguish of my spirit; I will complain in the bitterness of my soul.

\verse Am I a sea, or a whale, that thou settest a watch over me?  \verse When I say, My bed shall comfort me, my couch shall ease my complaints; \verse Then thou scarest me with dreams, and terrifiest me through visions: \verse So that my soul chooseth strangling, and death rather than my life.

\verse I loathe it; I would not live alway: let me alone; for my days are vanity.

\verse What is man, that thou shouldest magnify him? and that thou shouldest set thine heart upon him?  \verse And that thou shouldest visit him every morning, and try him every moment?  \verse How long wilt thou not depart from me, nor let me alone till I swallow down my spittle?  \verse I have sinned; what shall I do unto thee, O thou preserver of men?  why hast thou set me as a mark against thee, so that I am a burden to myself?  \verse And why dost thou not pardon my transgression, and take away my iniquity? for now shall I sleep in the dust; and thou shalt seek me in the morning, but I shall not be.


\chapter
\verse Then answered Bildad the Shuhite, and said, \verse How long wilt thou speak these things? and how long shall the words of thy mouth be like a strong wind?  \verse Doth God pervert judgment? or doth the Almighty pervert justice?  \verse If thy children have sinned against him, and he have cast them away for their transgression; \verse If thou wouldest seek unto God betimes, and make thy supplication to the Almighty; \verse If thou wert pure and upright; surely now he would awake for thee, and make the habitation of thy righteousness prosperous.

\verse Though thy beginning was small, yet thy latter end should greatly increase.

\verse For enquire, I pray thee, of the former age, and prepare thyself to the search of their fathers: \verse (For we are but of yesterday, and know nothing, because our days upon earth are a shadow:) \verse Shall not they teach thee, and tell thee, and utter words out of their heart?  \verse Can the rush grow up without mire? can the flag grow without water?  \verse Whilst it is yet in his greenness, and not cut down, it withereth before any other herb.

\verse So are the paths of all that forget God; and the hypocrite's hope shall perish: \verse Whose hope shall be cut off, and whose trust shall be a spider's web.

\verse He shall lean upon his house, but it shall not stand: he shall hold it fast, but it shall not endure.

\verse He is green before the sun, and his branch shooteth forth in his garden.

\verse His roots are wrapped about the heap, and seeth the place of stones.

\verse If he destroy him from his place, then it shall deny him, saying, I have not seen thee.

\verse Behold, this is the joy of his way, and out of the earth shall others grow.

\verse Behold, God will not cast away a perfect man, neither will he help the evil doers: \verse Till he fill thy mouth with laughing, and thy lips with rejoicing.

\verse They that hate thee shall be clothed with shame; and the dwelling place of the wicked shall come to nought.


\chapter
\verse Then Job answered and said, \verse I know it is so of a truth: but how should man be just with God?  \verse If he will contend with him, he cannot answer him one of a thousand.

\verse He is wise in heart, and mighty in strength: who hath hardened himself against him, and hath prospered?  \verse Which removeth the mountains, and they know not: which overturneth them in his anger.

\verse Which shaketh the earth out of her place, and the pillars thereof tremble.

\verse Which commandeth the sun, and it riseth not; and sealeth up the stars.

\verse Which alone spreadeth out the heavens, and treadeth upon the waves of the sea.

\verse Which maketh Arcturus, Orion, and Pleiades, and the chambers of the south.

\verse Which doeth great things past finding out; yea, and wonders without number.

\verse Lo, he goeth by me, and I see him not: he passeth on also, but I perceive him not.

\verse Behold, he taketh away, who can hinder him? who will say unto him, What doest thou?  \verse If God will not withdraw his anger, the proud helpers do stoop under him.

\verse How much less shall I answer him, and choose out my words to reason with him?  \verse Whom, though I were righteous, yet would I not answer, but I would make supplication to my judge.

\verse If I had called, and he had answered me; yet would I not believe that he had hearkened unto my voice.

\verse For he breaketh me with a tempest, and multiplieth my wounds without cause.

\verse He will not suffer me to take my breath, but filleth me with bitterness.

\verse If I speak of strength, lo, he is strong: and if of judgment, who shall set me a time to plead?  \verse If I justify myself, mine own mouth shall condemn me: if I say, I am perfect, it shall also prove me perverse.

\verse Though I were perfect, yet would I not know my soul: I would despise my life.

\verse This is one thing, therefore I said it, He destroyeth the perfect and the wicked.

\verse If the scourge slay suddenly, he will laugh at the trial of the innocent.

\verse The earth is given into the hand of the wicked: he covereth the faces of the judges thereof; if not, where, and who is he?  \verse Now my days are swifter than a post: they flee away, they see no good.

\verse They are passed away as the swift ships: as the eagle that hasteth to the prey.

\verse If I say, I will forget my complaint, I will leave off my heaviness, and comfort myself: \verse I am afraid of all my sorrows, I know that thou wilt not hold me innocent.

\verse If I be wicked, why then labour I in vain?  \verse If I wash myself with snow water, and make my hands never so clean; \verse Yet shalt thou plunge me in the ditch, and mine own clothes shall abhor me.

\verse For he is not a man, as I am, that I should answer him, and we should come together in judgment.

\verse Neither is there any daysman betwixt us, that might lay his hand upon us both.

\verse Let him take his rod away from me, and let not his fear terrify me: \verse Then would I speak, and not fear him; but it is not so with me.


\chapter
\verse My soul is weary of my life; I will leave my complaint upon myself; I will speak in the bitterness of my soul.

\verse I will say unto God, Do not condemn me; shew me wherefore thou contendest with me.

\verse Is it good unto thee that thou shouldest oppress, that thou shouldest despise the work of thine hands, and shine upon the counsel of the wicked?  \verse Hast thou eyes of flesh? or seest thou as man seeth?  \verse Are thy days as the days of man? are thy years as man's days, \verse That thou enquirest after mine iniquity, and searchest after my sin?  \verse Thou knowest that I am not wicked; and there is none that can deliver out of thine hand.

\verse Thine hands have made me and fashioned me together round about; yet thou dost destroy me.

\verse Remember, I beseech thee, that thou hast made me as the clay; and wilt thou bring me into dust again?  \verse Hast thou not poured me out as milk, and curdled me like cheese?  \verse Thou hast clothed me with skin and flesh, and hast fenced me with bones and sinews.

\verse Thou hast granted me life and favour, and thy visitation hath preserved my spirit.

\verse And these things hast thou hid in thine heart: I know that this is with thee.

\verse If I sin, then thou markest me, and thou wilt not acquit me from mine iniquity.

\verse If I be wicked, woe unto me; and if I be righteous, yet will I not lift up my head. I am full of confusion; therefore see thou mine affliction; \verse For it increaseth. Thou huntest me as a fierce lion: and again thou shewest thyself marvellous upon me.

\verse Thou renewest thy witnesses against me, and increasest thine indignation upon me; changes and war are against me.

\verse Wherefore then hast thou brought me forth out of the womb? Oh that I had given up the ghost, and no eye had seen me!  \verse I should have been as though I had not been; I should have been carried from the womb to the grave.

\verse Are not my days few? cease then, and let me alone, that I may take comfort a little, \verse Before I go whence I shall not return, even to the land of darkness and the shadow of death; \verse A land of darkness, as darkness itself; and of the shadow of death, without any order, and where the light is as darkness.


\chapter
\verse Then answered Zophar the Naamathite, and said, \verse Should not the multitude of words be answered? and should a man full of talk be justified?  \verse Should thy lies make men hold their peace? and when thou mockest, shall no man make thee ashamed?  \verse For thou hast said, My doctrine is pure, and I am clean in thine eyes.

\verse But oh that God would speak, and open his lips against thee; \verse And that he would shew thee the secrets of wisdom, that they are double to that which is! Know therefore that God exacteth of thee less than thine iniquity deserveth.

\verse Canst thou by searching find out God? canst thou find out the Almighty unto perfection?  \verse It is as high as heaven; what canst thou do? deeper than hell; what canst thou know?  \verse The measure thereof is longer than the earth, and broader than the sea.

\verse If he cut off, and shut up, or gather together, then who can hinder him?  \verse For he knoweth vain men: he seeth wickedness also; will he not then consider it?  \verse For vain men would be wise, though man be born like a wild ass's colt.

\verse If thou prepare thine heart, and stretch out thine hands toward him; \verse If iniquity be in thine hand, put it far away, and let not wickedness dwell in thy tabernacles.

\verse For then shalt thou lift up thy face without spot; yea, thou shalt be stedfast, and shalt not fear: \verse Because thou shalt forget thy misery, and remember it as waters that pass away: \verse And thine age shall be clearer than the noonday: thou shalt shine forth, thou shalt be as the morning.

\verse And thou shalt be secure, because there is hope; yea, thou shalt dig about thee, and thou shalt take thy rest in safety.

\verse Also thou shalt lie down, and none shall make thee afraid; yea, many shall make suit unto thee.

\verse But the eyes of the wicked shall fail, and they shall not escape, and their hope shall be as the giving up of the ghost.


\chapter
\verse And Job answered and said, \verse No doubt but ye are the people, and wisdom shall die with you.

\verse But I have understanding as well as you; I am not inferior to you: yea, who knoweth not such things as these?  \verse I am as one mocked of his neighbour, who calleth upon God, and he answereth him: the just upright man is laughed to scorn.

\verse He that is ready to slip with his feet is as a lamp despised in the thought of him that is at ease.

\verse The tabernacles of robbers prosper, and they that provoke God are secure; into whose hand God bringeth abundantly.

\verse But ask now the beasts, and they shall teach thee; and the fowls of the air, and they shall tell thee: \verse Or speak to the earth, and it shall teach thee: and the fishes of the sea shall declare unto thee.

\verse Who knoweth not in all these that the hand of the \LORD hath wrought this?  \verse In whose hand is the soul of every living thing, and the breath of all mankind.

\verse Doth not the ear try words? and the mouth taste his meat?  \verse With the ancient is wisdom; and in length of days understanding.

\verse With him is wisdom and strength, he hath counsel and understanding.

\verse Behold, he breaketh down, and it cannot be built again: he shutteth up a man, and there can be no opening.

\verse Behold, he withholdeth the waters, and they dry up: also he sendeth them out, and they overturn the earth.

\verse With him is strength and wisdom: the deceived and the deceiver are his.

\verse He leadeth counsellors away spoiled, and maketh the judges fools.

\verse He looseth the bond of kings, and girdeth their loins with a girdle.

\verse He leadeth princes away spoiled, and overthroweth the mighty.

\verse He removeth away the speech of the trusty, and taketh away the understanding of the aged.

\verse He poureth contempt upon princes, and weakeneth the strength of the mighty.

\verse He discovereth deep things out of darkness, and bringeth out to light the shadow of death.

\verse He increaseth the nations, and destroyeth them: he enlargeth the nations, and straiteneth them again.

\verse He taketh away the heart of the chief of the people of the earth, and causeth them to wander in a wilderness where there is no way.

\verse They grope in the dark without light, and he maketh them to stagger like a drunken man.


\chapter
\verse Lo, mine eye hath seen all this, mine ear hath heard and understood it.

\verse What ye know, the same do I know also: I am not inferior unto you.

\verse Surely I would speak to the Almighty, and I desire to reason with God.

\verse But ye are forgers of lies, ye are all physicians of no value.

\verse O that ye would altogether hold your peace! and it should be your wisdom.

\verse Hear now my reasoning, and hearken to the pleadings of my lips.

\verse Will ye speak wickedly for God? and talk deceitfully for him?  \verse Will ye accept his person? will ye contend for God?  \verse Is it good that he should search you out? or as one man mocketh another, do ye so mock him?  \verse He will surely reprove you, if ye do secretly accept persons.

\verse Shall not his excellency make you afraid? and his dread fall upon you?  \verse Your remembrances are like unto ashes, your bodies to bodies of clay.

\verse Hold your peace, let me alone, that I may speak, and let come on me what will.

\verse Wherefore do I take my flesh in my teeth, and put my life in mine hand?  \verse Though he slay me, yet will I trust in him: but I will maintain mine own ways before him.

\verse He also shall be my salvation: for an hypocrite shall not come before him.

\verse Hear diligently my speech, and my declaration with your ears.

\verse Behold now, I have ordered my cause; I know that I shall be justified.

\verse Who is he that will plead with me? for now, if I hold my tongue, I shall give up the ghost.

\verse Only do not two things unto me: then will I not hide myself from thee.

\verse Withdraw thine hand far from me: and let not thy dread make me afraid.

\verse Then call thou, and I will answer: or let me speak, and answer thou me.

\verse How many are mine iniquities and sins? make me to know my transgression and my sin.

\verse Wherefore hidest thou thy face, and holdest me for thine enemy?  \verse Wilt thou break a leaf driven to and fro? and wilt thou pursue the dry stubble?  \verse For thou writest bitter things against me, and makest me to possess the iniquities of my youth.

\verse Thou puttest my feet also in the stocks, and lookest narrowly unto all my paths; thou settest a print upon the heels of my feet.

\verse And he, as a rotten thing, consumeth, as a garment that is moth eaten.


\chapter
\verse Man that is born of a woman is of few days and full of trouble.

\verse He cometh forth like a flower, and is cut down: he fleeth also as a shadow, and continueth not.

\verse And doth thou open thine eyes upon such an one, and bringest me into judgment with thee?  \verse Who can bring a clean thing out of an unclean? not one.

\verse Seeing his days are determined, the number of his months are with thee, thou hast appointed his bounds that he cannot pass; \verse Turn from him, that he may rest, till he shall accomplish, as an hireling, his day.

\verse For there is hope of a tree, if it be cut down, that it will sprout again, and that the tender branch thereof will not cease.

\verse Though the root thereof wax old in the earth, and the stock thereof die in the ground; \verse Yet through the scent of water it will bud, and bring forth boughs like a plant.

\verse But man dieth, and wasteth away: yea, man giveth up the ghost, and where is he?  \verse As the waters fail from the sea, and the flood decayeth and drieth up: \verse So man lieth down, and riseth not: till the heavens be no more, they shall not awake, nor be raised out of their sleep.

\verse O that thou wouldest hide me in the grave, that thou wouldest keep me secret, until thy wrath be past, that thou wouldest appoint me a set time, and remember me!  \verse If a man die, shall he live again?  all the days of my appointed time will I wait, till my change come.

\verse Thou shalt call, and I will answer thee: thou wilt have a desire to the work of thine hands.

\verse For now thou numberest my steps: dost thou not watch over my sin?  \verse My transgression is sealed up in a bag, and thou sewest up mine iniquity.

\verse And surely the mountains falling cometh to nought, and the rock is removed out of his place.

\verse The waters wear the stones: thou washest away the things which grow out of the dust of the earth; and thou destroyest the hope of man.

\verse Thou prevailest for ever against him, and he passeth: thou changest his countenance, and sendest him away.

\verse His sons come to honour, and he knoweth it not; and they are brought low, but he perceiveth it not of them.

\verse But his flesh upon him shall have pain, and his soul within him shall mourn.


\chapter
\verse Then answered Eliphaz the Temanite, and said, \verse Should a wise man utter vain knowledge, and fill his belly with the east wind?  \verse Should he reason with unprofitable talk? or with speeches wherewith he can do no good?  \verse Yea, thou castest off fear, and restrainest prayer before God.

\verse For thy mouth uttereth thine iniquity, and thou choosest the tongue of the crafty.

\verse Thine own mouth condemneth thee, and not I: yea, thine own lips testify against thee.

\verse Art thou the first man that was born? or wast thou made before the hills?  \verse Hast thou heard the secret of God? and dost thou restrain wisdom to thyself?  \verse What knowest thou, that we know not?  what understandest thou, which is not in us?  \verse With us are both the grayheaded and very aged men, much elder than thy father.

\verse Are the consolations of God small with thee? is there any secret thing with thee?  \verse Why doth thine heart carry thee away? and what do thy eyes wink at, \verse That thou turnest thy spirit against God, and lettest such words go out of thy mouth?  \verse What is man, that he should be clean? and he which is born of a woman, that he should be righteous?  \verse Behold, he putteth no trust in his saints; yea, the heavens are not clean in his sight.

\verse How much more abominable and filthy is man, which drinketh iniquity like water?  \verse I will shew thee, hear me; and that which I have seen I will declare; \verse Which wise men have told from their fathers, and have not hid it: \verse Unto whom alone the earth was given, and no stranger passed among them.

\verse The wicked man travaileth with pain all his days, and the number of years is hidden to the oppressor.

\verse A dreadful sound is in his ears: in prosperity the destroyer shall come upon him.

\verse He believeth not that he shall return out of darkness, and he is waited for of the sword.

\verse He wandereth abroad for bread, saying, Where is it? he knoweth that the day of darkness is ready at his hand.

\verse Trouble and anguish shall make him afraid; they shall prevail against him, as a king ready to the battle.

\verse For he stretcheth out his hand against God, and strengtheneth himself against the Almighty.

\verse He runneth upon him, even on his neck, upon the thick bosses of his bucklers: \verse Because he covereth his face with his fatness, and maketh collops of fat on his flanks.

\verse And he dwelleth in desolate cities, and in houses which no man inhabiteth, which are ready to become heaps.

\verse He shall not be rich, neither shall his substance continue, neither shall he prolong the perfection thereof upon the earth.

\verse He shall not depart out of darkness; the flame shall dry up his branches, and by the breath of his mouth shall he go away.

\verse Let not him that is deceived trust in vanity: for vanity shall be his recompence.

\verse It shall be accomplished before his time, and his branch shall not be green.

\verse He shall shake off his unripe grape as the vine, and shall cast off his flower as the olive.

\verse For the congregation of hypocrites shall be desolate, and fire shall consume the tabernacles of bribery.

\verse They conceive mischief, and bring forth vanity, and their belly prepareth deceit.


\chapter
\verse Then Job answered and said, \verse I have heard many such things: miserable comforters are ye all.

\verse Shall vain words have an end? or what emboldeneth thee that thou answerest?  \verse I also could speak as ye do: if your soul were in my soul's stead, I could heap up words against you, and shake mine head at you.

\verse But I would strengthen you with my mouth, and the moving of my lips should asswage your grief.

\verse Though I speak, my grief is not asswaged: and though I forbear, what am I eased?  \verse But now he hath made me weary: thou hast made desolate all my company.

\verse And thou hast filled me with wrinkles, which is a witness against me: and my leanness rising up in me beareth witness to my face.

\verse He teareth me in his wrath, who hateth me: he gnasheth upon me with his teeth; mine enemy sharpeneth his eyes upon me.

\verse They have gaped upon me with their mouth; they have smitten me upon the cheek reproachfully; they have gathered themselves together against me.

\verse God hath delivered me to the ungodly, and turned me over into the hands of the wicked.

\verse I was at ease, but he hath broken me asunder: he hath also taken me by my neck, and shaken me to pieces, and set me up for his mark.

\verse His archers compass me round about, he cleaveth my reins asunder, and doth not spare; he poureth out my gall upon the ground.

\verse He breaketh me with breach upon breach, he runneth upon me like a giant.

\verse I have sewed sackcloth upon my skin, and defiled my horn in the dust.

\verse My face is foul with weeping, and on my eyelids is the shadow of death; \verse Not for any injustice in mine hands: also my prayer is pure.

\verse O earth, cover not thou my blood, and let my cry have no place.

\verse Also now, behold, my witness is in heaven, and my record is on high.

\verse My friends scorn me: but mine eye poureth out tears unto God.

\verse O that one might plead for a man with God, as a man pleadeth for his neighbour!  \verse When a few years are come, then I shall go the way whence I shall not return.


\chapter
\verse My breath is corrupt, my days are extinct, the graves are ready for me.

\verse Are there not mockers with me? and doth not mine eye continue in their provocation?  \verse Lay down now, put me in a surety with thee; who is he that will strike hands with me?  \verse For thou hast hid their heart from understanding: therefore shalt thou not exalt them.

\verse He that speaketh flattery to his friends, even the eyes of his children shall fail.

\verse He hath made me also a byword of the people; and aforetime I was as a tabret.

\verse Mine eye also is dim by reason of sorrow, and all my members are as a shadow.

\verse Upright men shall be astonied at this, and the innocent shall stir up himself against the hypocrite.

\verse The righteous also shall hold on his way, and he that hath clean hands shall be stronger and stronger.

\verse But as for you all, do ye return, and come now: for I cannot find one wise man among you.

\verse My days are past, my purposes are broken off, even the thoughts of my heart.

\verse They change the night into day: the light is short because of darkness.

\verse If I wait, the grave is mine house: I have made my bed in the darkness.

\verse I have said to corruption, Thou art my father: to the worm, Thou art my mother, and my sister.

\verse And where is now my hope? as for my hope, who shall see it?  \verse They shall go down to the bars of the pit, when our rest together is in the dust.


\chapter
\verse Then answered Bildad the Shuhite, and said, \verse How long will it be ere ye make an end of words? mark, and afterwards we will speak.

\verse Wherefore are we counted as beasts, and reputed vile in your sight?  \verse He teareth himself in his anger: shall the earth be forsaken for thee? and shall the rock be removed out of his place?  \verse Yea, the light of the wicked shall be put out, and the spark of his fire shall not shine.

\verse The light shall be dark in his tabernacle, and his candle shall be put out with him.

\verse The steps of his strength shall be straitened, and his own counsel shall cast him down.

\verse For he is cast into a net by his own feet, and he walketh upon a snare.

\verse The gin shall take him by the heel, and the robber shall prevail against him.

\verse The snare is laid for him in the ground, and a trap for him in the way.

\verse Terrors shall make him afraid on every side, and shall drive him to his feet.

\verse His strength shall be hungerbitten, and destruction shall be ready at his side.

\verse It shall devour the strength of his skin: even the firstborn of death shall devour his strength.

\verse His confidence shall be rooted out of his tabernacle, and it shall bring him to the king of terrors.

\verse It shall dwell in his tabernacle, because it is none of his: brimstone shall be scattered upon his habitation.

\verse His roots shall be dried up beneath, and above shall his branch be cut off.

\verse His remembrance shall perish from the earth, and he shall have no name in the street.

\verse He shall be driven from light into darkness, and chased out of the world.

\verse He shall neither have son nor nephew among his people, nor any remaining in his dwellings.

\verse They that come after him shall be astonied at his day, as they that went before were affrighted.

\verse Surely such are the dwellings of the wicked, and this is the place of him that knoweth not God.


\chapter
\verse Then Job answered and said, \verse How long will ye vex my soul, and break me in pieces with words?  \verse These ten times have ye reproached me: ye are not ashamed that ye make yourselves strange to me.

\verse And be it indeed that I have erred, mine error remaineth with myself.

\verse If indeed ye will magnify yourselves against me, and plead against me my reproach: \verse Know now that God hath overthrown me, and hath compassed me with his net.

\verse Behold, I cry out of wrong, but I am not heard: I cry aloud, but there is no judgment.

\verse He hath fenced up my way that I cannot pass, and he hath set darkness in my paths.

\verse He hath stripped me of my glory, and taken the crown from my head.

\verse He hath destroyed me on every side, and I am gone: and mine hope hath he removed like a tree.

\verse He hath also kindled his wrath against me, and he counteth me unto him as one of his enemies.

\verse His troops come together, and raise up their way against me, and encamp round about my tabernacle.

\verse He hath put my brethren far from me, and mine acquaintance are verily estranged from me.

\verse My kinsfolk have failed, and my familiar friends have forgotten me.

\verse They that dwell in mine house, and my maids, count me for a stranger: I am an alien in their sight.

\verse I called my servant, and he gave me no answer; I intreated him with my mouth.

\verse My breath is strange to my wife, though I intreated for the children's sake of mine own body.

\verse Yea, young children despised me; I arose, and they spake against me.

\verse All my inward friends abhorred me: and they whom I loved are turned against me.

\verse My bone cleaveth to my skin and to my flesh, and I am escaped with the skin of my teeth.

\verse Have pity upon me, have pity upon me, O ye my friends; for the hand of God hath touched me.

\verse Why do ye persecute me as God, and are not satisfied with my flesh?  \verse Oh that my words were now written! oh that they were printed in a book!  \verse That they were graven with an iron pen and lead in the rock for ever!  \verse For I know that my redeemer liveth, and that he shall stand at the latter day upon the earth: \verse And though after my skin worms destroy this body, yet in my flesh shall I see God: \verse Whom I shall see for myself, and mine eyes shall behold, and not another; though my reins be consumed within me.

\verse But ye should say, Why persecute we him, seeing the root of the matter is found in me?  \verse Be ye afraid of the sword: for wrath bringeth the punishments of the sword, that ye may know there is a judgment.


\chapter
\verse Then answered Zophar the Naamathite, and said, \verse Therefore do my thoughts cause me to answer, and for this I make haste.

\verse I have heard the check of my reproach, and the spirit of my understanding causeth me to answer.

\verse Knowest thou not this of old, since man was placed upon earth, \verse That the triumphing of the wicked is short, and the joy of the hypocrite but for a moment?  \verse Though his excellency mount up to the heavens, and his head reach unto the clouds; \verse Yet he shall perish for ever like his own dung: they which have seen him shall say, Where is he?  \verse He shall fly away as a dream, and shall not be found: yea, he shall be chased away as a vision of the night.

\verse The eye also which saw him shall see him no more; neither shall his place any more behold him.

\verse His children shall seek to please the poor, and his hands shall restore their goods.

\verse His bones are full of the sin of his youth, which shall lie down with him in the dust.

\verse Though wickedness be sweet in his mouth, though he hide it under his tongue; \verse Though he spare it, and forsake it not; but keep it still within his mouth: \verse Yet his meat in his bowels is turned, it is the gall of asps within him.

\verse He hath swallowed down riches, and he shall vomit them up again: God shall cast them out of his belly.

\verse He shall suck the poison of asps: the viper's tongue shall slay him.

\verse He shall not see the rivers, the floods, the brooks of honey and butter.

\verse That which he laboured for shall he restore, and shall not swallow it down: according to his substance shall the restitution be, and he shall not rejoice therein.

\verse Because he hath oppressed and hath forsaken the poor; because he hath violently taken away an house which he builded not; \verse Surely he shall not feel quietness in his belly, he shall not save of that which he desired.

\verse There shall none of his meat be left; therefore shall no man look for his goods.

\verse In the fulness of his sufficiency he shall be in straits: every hand of the wicked shall come upon him.

\verse When he is about to fill his belly, God shall cast the fury of his wrath upon him, and shall rain it upon him while he is eating.

\verse He shall flee from the iron weapon, and the bow of steel shall strike him through.

\verse It is drawn, and cometh out of the body; yea, the glittering sword cometh out of his gall: terrors are upon him.

\verse All darkness shall be hid in his secret places: a fire not blown shall consume him; it shall go ill with him that is left in his tabernacle.

\verse The heaven shall reveal his iniquity; and the earth shall rise up against him.

\verse The increase of his house shall depart, and his goods shall flow away in the day of his wrath.

\verse This is the portion of a wicked man from God, and the heritage appointed unto him by God.


\chapter
\verse But Job answered and said, \verse Hear diligently my speech, and let this be your consolations.

\verse Suffer me that I may speak; and after that I have spoken, mock on.

\verse As for me, is my complaint to man? and if it were so, why should not my spirit be troubled?  \verse Mark me, and be astonished, and lay your hand upon your mouth.

\verse Even when I remember I am afraid, and trembling taketh hold on my flesh.

\verse Wherefore do the wicked live, become old, yea, are mighty in power?  \verse Their seed is established in their sight with them, and their offspring before their eyes.

\verse Their houses are safe from fear, neither is the rod of God upon them.

\verse Their bull gendereth, and faileth not; their cow calveth, and casteth not her calf.

\verse They send forth their little ones like a flock, and their children dance.

\verse They take the timbrel and harp, and rejoice at the sound of the organ.

\verse They spend their days in wealth, and in a moment go down to the grave.

\verse Therefore they say unto God, Depart from us; for we desire not the knowledge of thy ways.

\verse What is the Almighty, that we should serve him? and what profit should we have, if we pray unto him?  \verse Lo, their good is not in their hand: the counsel of the wicked is far from me.

\verse How oft is the candle of the wicked put out! and how oft cometh their destruction upon them! God distributeth sorrows in his anger.

\verse They are as stubble before the wind, and as chaff that the storm carrieth away.

\verse God layeth up his iniquity for his children: he rewardeth him, and he shall know it.

\verse His eyes shall see his destruction, and he shall drink of the wrath of the Almighty.

\verse For what pleasure hath he in his house after him, when the number of his months is cut off in the midst?  \verse Shall any teach God knowledge? seeing he judgeth those that are high.

\verse One dieth in his full strength, being wholly at ease and quiet.

\verse His breasts are full of milk, and his bones are moistened with marrow.

\verse And another dieth in the bitterness of his soul, and never eateth with pleasure.

\verse They shall lie down alike in the dust, and the worms shall cover them.

\verse Behold, I know your thoughts, and the devices which ye wrongfully imagine against me.

\verse For ye say, Where is the house of the prince? and where are the dwelling places of the wicked?  \verse Have ye not asked them that go by the way? and do ye not know their tokens, \verse That the wicked is reserved to the day of destruction? they shall be brought forth to the day of wrath.

\verse Who shall declare his way to his face? and who shall repay him what he hath done?  \verse Yet shall he be brought to the grave, and shall remain in the tomb.

\verse The clods of the valley shall be sweet unto him, and every man shall draw after him, as there are innumerable before him.

\verse How then comfort ye me in vain, seeing in your answers there remaineth falsehood?  
\chapter
\verse Then Eliphaz the Temanite answered and said, \verse Can a man be profitable unto God, as he that is wise may be profitable unto himself?  \verse Is it any pleasure to the Almighty, that thou art righteous? or is it gain to him, that thou makest thy ways perfect?  \verse Will he reprove thee for fear of thee? will he enter with thee into judgment?  \verse Is not thy wickedness great? and thine iniquities infinite?  \verse For thou hast taken a pledge from thy brother for nought, and stripped the naked of their clothing.

\verse Thou hast not given water to the weary to drink, and thou hast withholden bread from the hungry.

\verse But as for the mighty man, he had the earth; and the honourable man dwelt in it.

\verse Thou hast sent widows away empty, and the arms of the fatherless have been broken.

\verse Therefore snares are round about thee, and sudden fear troubleth thee; \verse Or darkness, that thou canst not see; and abundance of waters cover thee.

\verse Is not God in the height of heaven? and behold the height of the stars, how high they are!  \verse And thou sayest, How doth God know?  can he judge through the dark cloud?  \verse Thick clouds are a covering to him, that he seeth not; and he walketh in the circuit of heaven.

\verse Hast thou marked the old way which wicked men have trodden?  \verse Which were cut down out of time, whose foundation was overflown with a flood: \verse Which said unto God, Depart from us: and what can the Almighty do for them?  \verse Yet he filled their houses with good things: but the counsel of the wicked is far from me.

\verse The righteous see it, and are glad: and the innocent laugh them to scorn.

\verse Whereas our substance is not cut down, but the remnant of them the fire consumeth.

\verse Acquaint now thyself with him, and be at peace: thereby good shall come unto thee.

\verse Receive, I pray thee, the law from his mouth, and lay up his words in thine heart.

\verse If thou return to the Almighty, thou shalt be built up, thou shalt put away iniquity far from thy tabernacles.

\verse Then shalt thou lay up gold as dust, and the gold of Ophir as the stones of the brooks.

\verse Yea, the Almighty shall be thy defence, and thou shalt have plenty of silver.

\verse For then shalt thou have thy delight in the Almighty, and shalt lift up thy face unto God.

\verse Thou shalt make thy prayer unto him, and he shall hear thee, and thou shalt pay thy vows.

\verse Thou shalt also decree a thing, and it shall be established unto thee: and the light shall shine upon thy ways.

\verse When men are cast down, then thou shalt say, There is lifting up; and he shall save the humble person.

\verse He shall deliver the island of the innocent: and it is delivered by the pureness of thine hands.


\chapter
\verse Then Job answered and said, \verse Even to day is my complaint bitter: my stroke is heavier than my groaning.

\verse Oh that I knew where I might find him! that I might come even to his seat!  \verse I would order my cause before him, and fill my mouth with arguments.

\verse I would know the words which he would answer me, and understand what he would say unto me.

\verse Will he plead against me with his great power? No; but he would put strength in me.

\verse There the righteous might dispute with him; so should I be delivered for ever from my judge.

\verse Behold, I go forward, but he is not there; and backward, but I cannot perceive him: \verse On the left hand, where he doth work, but I cannot behold him: he hideth himself on the right hand, that I cannot see him: \verse But he knoweth the way that I take: when he hath tried me, I shall come forth as gold.

\verse My foot hath held his steps, his way have I kept, and not declined.

\verse Neither have I gone back from the commandment of his lips; I have esteemed the words of his mouth more than my necessary food.

\verse But he is in one mind, and who can turn him? and what his soul desireth, even that he doeth.

\verse For he performeth the thing that is appointed for me: and many such things are with him.

\verse Therefore am I troubled at his presence: when I consider, I am afraid of him.

\verse For God maketh my heart soft, and the Almighty troubleth me: \verse Because I was not cut off before the darkness, neither hath he covered the darkness from my face.


\chapter
\verse Why, seeing times are not hidden from the Almighty, do they that know him not see his days?  \verse Some remove the landmarks; they violently take away flocks, and feed thereof.

\verse They drive away the ass of the fatherless, they take the widow's ox for a pledge.

\verse They turn the needy out of the way: the poor of the earth hide themselves together.

\verse Behold, as wild asses in the desert, go they forth to their work; rising betimes for a prey: the wilderness yieldeth food for them and for their children.

\verse They reap every one his corn in the field: and they gather the vintage of the wicked.

\verse They cause the naked to lodge without clothing, that they have no covering in the cold.

\verse They are wet with the showers of the mountains, and embrace the rock for want of a shelter.

\verse They pluck the fatherless from the breast, and take a pledge of the poor.

\verse They cause him to go naked without clothing, and they take away the sheaf from the hungry; \verse Which make oil within their walls, and tread their winepresses, and suffer thirst.

\verse Men groan from out of the city, and the soul of the wounded crieth out: yet God layeth not folly to them.

\verse They are of those that rebel against the light; they know not the ways thereof, nor abide in the paths thereof.

\verse The murderer rising with the light killeth the poor and needy, and in the night is as a thief.

\verse The eye also of the adulterer waiteth for the twilight, saying, No eye shall see me: and disguiseth his face.

\verse In the dark they dig through houses, which they had marked for themselves in the daytime: they know not the light.

\verse For the morning is to them even as the shadow of death: if one know them, they are in the terrors of the shadow of death.

\verse He is swift as the waters; their portion is cursed in the earth: he beholdeth not the way of the vineyards.

\verse Drought and heat consume the snow waters: so doth the grave those which have sinned.

\verse The womb shall forget him; the worm shall feed sweetly on him; he shall be no more remembered; and wickedness shall be broken as a tree.

\verse He evil entreateth the barren that beareth not: and doeth not good to the widow.

\verse He draweth also the mighty with his power: he riseth up, and no man is sure of life.

\verse Though it be given him to be in safety, whereon he resteth; yet his eyes are upon their ways.

\verse They are exalted for a little while, but are gone and brought low; they are taken out of the way as all other, and cut off as the tops of the ears of corn.

\verse And if it be not so now, who will make me a liar, and make my speech nothing worth?  
\chapter
\verse Then answered Bildad the Shuhite, and said, \verse Dominion and fear are with him, he maketh peace in his high places.

\verse Is there any number of his armies? and upon whom doth not his light arise?  \verse How then can man be justified with God? or how can he be clean that is born of a woman?  \verse Behold even to the moon, and it shineth not; yea, the stars are not pure in his sight.

\verse How much less man, that is a worm? and the son of man, which is a worm?  
\chapter
\verse But Job answered and said, \verse How hast thou helped him that is without power? how savest thou the arm that hath no strength?  \verse How hast thou counselled him that hath no wisdom? and how hast thou plentifully declared the thing as it is?  \verse To whom hast thou uttered words? and whose spirit came from thee?  \verse Dead things are formed from under the waters, and the inhabitants thereof.

\verse Hell is naked before him, and destruction hath no covering.

\verse He stretcheth out the north over the empty place, and hangeth the earth upon nothing.

\verse He bindeth up the waters in his thick clouds; and the cloud is not rent under them.

\verse He holdeth back the face of his throne, and spreadeth his cloud upon it.

\verse He hath compassed the waters with bounds, until the day and night come to an end.

\verse The pillars of heaven tremble and are astonished at his reproof.

\verse He divideth the sea with his power, and by his understanding he smiteth through the proud.

\verse By his spirit he hath garnished the heavens; his hand hath formed the crooked serpent.

\verse Lo, these are parts of his ways: but how little a portion is heard of him? but the thunder of his power who can understand?  
\chapter
\verse Moreover Job continued his parable, and said, \verse As God liveth, who hath taken away my judgment; and the Almighty, who hath vexed my soul; \verse All the while my breath is in me, and the spirit of God is in my nostrils; \verse My lips shall not speak wickedness, nor my tongue utter deceit.

\verse God forbid that I should justify you: till I die I will not remove mine integrity from me.

\verse My righteousness I hold fast, and will not let it go: my heart shall not reproach me so long as I live.

\verse Let mine enemy be as the wicked, and he that riseth up against me as the unrighteous.

\verse For what is the hope of the hypocrite, though he hath gained, when God taketh away his soul?  \verse Will God hear his cry when trouble cometh upon him?  \verse Will he delight himself in the Almighty? will he always call upon God?  \verse I will teach you by the hand of God: that which is with the Almighty will I not conceal.

\verse Behold, all ye yourselves have seen it; why then are ye thus altogether vain?  \verse This is the portion of a wicked man with God, and the heritage of oppressors, which they shall receive of the Almighty.

\verse If his children be multiplied, it is for the sword: and his offspring shall not be satisfied with bread.

\verse Those that remain of him shall be buried in death: and his widows shall not weep.

\verse Though he heap up silver as the dust, and prepare raiment as the clay; \verse He may prepare it, but the just shall put it on, and the innocent shall divide the silver.

\verse He buildeth his house as a moth, and as a booth that the keeper maketh.

\verse The rich man shall lie down, but he shall not be gathered: he openeth his eyes, and he is not.

\verse Terrors take hold on him as waters, a tempest stealeth him away in the night.

\verse The east wind carrieth him away, and he departeth: and as a storm hurleth him out of his place.

\verse For God shall cast upon him, and not spare: he would fain flee out of his hand.

\verse Men shall clap their hands at him, and shall hiss him out of his place.


\chapter
\verse Surely there is a vein for the silver, and a place for gold where they fine it.

\verse Iron is taken out of the earth, and brass is molten out of the stone.

\verse He setteth an end to darkness, and searcheth out all perfection: the stones of darkness, and the shadow of death.

\verse The flood breaketh out from the inhabitant; even the waters forgotten of the foot: they are dried up, they are gone away from men.

\verse As for the earth, out of it cometh bread: and under it is turned up as it were fire.

\verse The stones of it are the place of sapphires: and it hath dust of gold.

\verse There is a path which no fowl knoweth, and which the vulture's eye hath not seen: \verse The lion's whelps have not trodden it, nor the fierce lion passed by it.

\verse He putteth forth his hand upon the rock; he overturneth the mountains by the roots.

\verse He cutteth out rivers among the rocks; and his eye seeth every precious thing.

\verse He bindeth the floods from overflowing; and the thing that is hid bringeth he forth to light.

\verse But where shall wisdom be found? and where is the place of understanding?  \verse Man knoweth not the price thereof; neither is it found in the land of the living.

\verse The depth saith, It is not in me: and the sea saith, It is not with me.

\verse It cannot be gotten for gold, neither shall silver be weighed for the price thereof.

\verse It cannot be valued with the gold of Ophir, with the precious onyx, or the sapphire.

\verse The gold and the crystal cannot equal it: and the exchange of it shall not be for jewels of fine gold.

\verse No mention shall be made of coral, or of pearls: for the price of wisdom is above rubies.

\verse The topaz of Ethiopia shall not equal it, neither shall it be valued with pure gold.

\verse Whence then cometh wisdom? and where is the place of understanding?  \verse Seeing it is hid from the eyes of all living, and kept close from the fowls of the air.

\verse Destruction and death say, We have heard the fame thereof with our ears.

\verse God understandeth the way thereof, and he knoweth the place thereof.

\verse For he looketh to the ends of the earth, and seeth under the whole heaven; \verse To make the weight for the winds; and he weigheth the waters by measure.

\verse When he made a decree for the rain, and a way for the lightning of the thunder: \verse Then did he see it, and declare it; he prepared it, yea, and searched it out.

\verse And unto man he said, Behold, the fear of the \LORD, that is wisdom; and to depart from evil is understanding.


\chapter
\verse Moreover Job continued his parable, and said, \verse Oh that I were as in months past, as in the days when God preserved me; \verse When his candle shined upon my head, and when by his light I walked through darkness; \verse As I was in the days of my youth, when the secret of God was upon my tabernacle; \verse When the Almighty was yet with me, when my children were about me; \verse When I washed my steps with butter, and the rock poured me out rivers of oil; \verse When I went out to the gate through the city, when I prepared my seat in the street!  \verse The young men saw me, and hid themselves: and the aged arose, and stood up.

\verse The princes refrained talking, and laid their hand on their mouth.

\verse The nobles held their peace, and their tongue cleaved to the roof of their mouth.

\verse When the ear heard me, then it blessed me; and when the eye saw me, it gave witness to me: \verse Because I delivered the poor that cried, and the fatherless, and him that had none to help him.

\verse The blessing of him that was ready to perish came upon me: and I caused the widow's heart to sing for joy.

\verse I put on righteousness, and it clothed me: my judgment was as a robe and a diadem.

\verse I was eyes to the blind, and feet was I to the lame.

\verse I was a father to the poor: and the cause which I knew not I searched out.

\verse And I brake the jaws of the wicked, and plucked the spoil out of his teeth.

\verse Then I said, I shall die in my nest, and I shall multiply my days as the sand.

\verse My root was spread out by the waters, and the dew lay all night upon my branch.

\verse My glory was fresh in me, and my bow was renewed in my hand.

\verse Unto me men gave ear, and waited, and kept silence at my counsel.

\verse After my words they spake not again; and my speech dropped upon them.

\verse And they waited for me as for the rain; and they opened their mouth wide as for the latter rain.

\verse If I laughed on them, they believed it not; and the light of my countenance they cast not down.

\verse I chose out their way, and sat chief, and dwelt as a king in the army, as one that comforteth the mourners.


\chapter
\verse But now they that are younger than I have me in derision, whose fathers I would have disdained to have set with the dogs of my flock.

\verse Yea, whereto might the strength of their hands profit me, in whom old age was perished?  \verse For want and famine they were solitary; fleeing into the wilderness in former time desolate and waste.

\verse Who cut up mallows by the bushes, and juniper roots for their meat.

\verse They were driven forth from among men, (they cried after them as after a thief;) \verse To dwell in the cliffs of the valleys, in caves of the earth, and in the rocks.

\verse Among the bushes they brayed; under the nettles they were gathered together.

\verse They were children of fools, yea, children of base men: they were viler than the earth.

\verse And now am I their song, yea, I am their byword.

\verse They abhor me, they flee far from me, and spare not to spit in my face.

\verse Because he hath loosed my cord, and afflicted me, they have also let loose the bridle before me.

\verse Upon my right hand rise the youth; they push away my feet, and they raise up against me the ways of their destruction.

\verse They mar my path, they set forward my calamity, they have no helper.

\verse They came upon me as a wide breaking in of waters: in the desolation they rolled themselves upon me.

\verse Terrors are turned upon me: they pursue my soul as the wind: and my welfare passeth away as a cloud.

\verse And now my soul is poured out upon me; the days of affliction have taken hold upon me.

\verse My bones are pierced in me in the night season: and my sinews take no rest.

\verse By the great force of my disease is my garment changed: it bindeth me about as the collar of my coat.

\verse He hath cast me into the mire, and I am become like dust and ashes.

\verse I cry unto thee, and thou dost not hear me: I stand up, and thou regardest me not.

\verse Thou art become cruel to me: with thy strong hand thou opposest thyself against me.

\verse Thou liftest me up to the wind; thou causest me to ride upon it, and dissolvest my substance.

\verse For I know that thou wilt bring me to death, and to the house appointed for all living.

\verse Howbeit he will not stretch out his hand to the grave, though they cry in his destruction.

\verse Did not I weep for him that was in trouble? was not my soul grieved for the poor?  \verse When I looked for good, then evil came unto me: and when I waited for light, there came darkness.

\verse My bowels boiled, and rested not: the days of affliction prevented me.

\verse I went mourning without the sun: I stood up, and I cried in the congregation.

\verse I am a brother to dragons, and a companion to owls.

\verse My skin is black upon me, and my bones are burned with heat.

\verse My harp also is turned to mourning, and my organ into the voice of them that weep.


\chapter
\verse I made a covenant with mine eyes; why then should I think upon a maid?  \verse For what portion of God is there from above? and what inheritance of the Almighty from on high?  \verse Is not destruction to the wicked? and a strange punishment to the workers of iniquity?  \verse Doth not he see my ways, and count all my steps?  \verse If I have walked with vanity, or if my foot hath hasted to deceit; \verse Let me be weighed in an even balance that God may know mine integrity.

\verse If my step hath turned out of the way, and mine heart walked after mine eyes, and if any blot hath cleaved to mine hands; \verse Then let me sow, and let another eat; yea, let my offspring be rooted out.

\verse If mine heart have been deceived by a woman, or if I have laid wait at my neighbour's door; \verse Then let my wife grind unto another, and let others bow down upon her.

\verse For this is an heinous crime; yea, it is an iniquity to be punished by the judges.

\verse For it is a fire that consumeth to destruction, and would root out all mine increase.

\verse If I did despise the cause of my manservant or of my maidservant, when they contended with me; \verse What then shall I do when God riseth up? and when he visiteth, what shall I answer him?  \verse Did not he that made me in the womb make him? and did not one fashion us in the womb?  \verse If I have withheld the poor from their desire, or have caused the eyes of the widow to fail; \verse Or have eaten my morsel myself alone, and the fatherless hath not eaten thereof; \verse (For from my youth he was brought up with me, as with a father, and I have guided her from my mother's womb;) \verse If I have seen any perish for want of clothing, or any poor without covering; \verse If his loins have not blessed me, and if he were not warmed with the fleece of my sheep; \verse If I have lifted up my hand against the fatherless, when I saw my help in the gate: \verse Then let mine arm fall from my shoulder blade, and mine arm be broken from the bone.

\verse For destruction from God was a terror to me, and by reason of his highness I could not endure.

\verse If I have made gold my hope, or have said to the fine gold, Thou art my confidence; \verse If I rejoice because my wealth was great, and because mine hand had gotten much; \verse If I beheld the sun when it shined, or the moon walking in brightness; \verse And my heart hath been secretly enticed, or my mouth hath kissed my hand: \verse This also were an iniquity to be punished by the judge: for I should have denied the God that is above.

\verse If I rejoice at the destruction of him that hated me, or lifted up myself when evil found him: \verse Neither have I suffered my mouth to sin by wishing a curse to his soul.

\verse If the men of my tabernacle said not, Oh that we had of his flesh!  we cannot be satisfied.

\verse The stranger did not lodge in the street: but I opened my doors to the traveller.

\verse If I covered my transgressions as Adam, by hiding mine iniquity in my bosom: \verse Did I fear a great multitude, or did the contempt of families terrify me, that I kept silence, and went not out of the door?  \verse Oh that one would hear me! behold, my desire is, that the Almighty would answer me, and that mine adversary had written a book.

\verse Surely I would take it upon my shoulder, and bind it as a crown to me.

\verse I would declare unto him the number of my steps; as a prince would I go near unto him.

\verse If my land cry against me, or that the furrows likewise thereof complain; \verse If I have eaten the fruits thereof without money, or have caused the owners thereof to lose their life: \verse Let thistles grow instead of wheat, and cockle instead of barley.

The words of Job are ended.


\chapter
\verse So these three men ceased to answer Job, because he was righteous in his own eyes.

\verse Then was kindled the wrath of Elihu the son of Barachel the Buzite, of the kindred of Ram: against Job was his wrath kindled, because he justified himself rather than God.

\verse Also against his three friends was his wrath kindled, because they had found no answer, and yet had condemned Job.

\verse Now Elihu had waited till Job had spoken, because they were elder than he.

\verse When Elihu saw that there was no answer in the mouth of these three men, then his wrath was kindled.

\verse And Elihu the son of Barachel the Buzite answered and said, I am young, and ye are very old; wherefore I was afraid, and durst not shew you mine opinion.

\verse I said, Days should speak, and multitude of years should teach wisdom.

\verse But there is a spirit in man: and the inspiration of the Almighty giveth them understanding.

\verse Great men are not always wise: neither do the aged understand judgment.

\verse Therefore I said, Hearken to me; I also will shew mine opinion.

\verse Behold, I waited for your words; I gave ear to your reasons, whilst ye searched out what to say.

\verse Yea, I attended unto you, and, behold, there was none of you that convinced Job, or that answered his words: \verse Lest ye should say, We have found out wisdom: God thrusteth him down, not man.

\verse Now he hath not directed his words against me: neither will I answer him with your speeches.

\verse They were amazed, they answered no more: they left off speaking.

\verse When I had waited, (for they spake not, but stood still, and answered no more;) \verse I said, I will answer also my part, I also will shew mine opinion.

\verse For I am full of matter, the spirit within me constraineth me.

\verse Behold, my belly is as wine which hath no vent; it is ready to burst like new bottles.

\verse I will speak, that I may be refreshed: I will open my lips and answer.

\verse Let me not, I pray you, accept any man's person, neither let me give flattering titles unto man.

\verse For I know not to give flattering titles; in so doing my maker would soon take me away.


\chapter
\verse Wherefore, Job, I pray thee, hear my speeches, and hearken to all my words.

\verse Behold, now I have opened my mouth, my tongue hath spoken in my mouth.

\verse My words shall be of the uprightness of my heart: and my lips shall utter knowledge clearly.

\verse The spirit of God hath made me, and the breath of the Almighty hath given me life.

\verse If thou canst answer me, set thy words in order before me, stand up.

\verse Behold, I am according to thy wish in God's stead: I also am formed out of the clay.

\verse Behold, my terror shall not make thee afraid, neither shall my hand be heavy upon thee.

\verse Surely thou hast spoken in mine hearing, and I have heard the voice of thy words, saying, \verse I am clean without transgression, I am innocent; neither is there iniquity in me.

\verse Behold, he findeth occasions against me, he counteth me for his enemy, \verse He putteth my feet in the stocks, he marketh all my paths.

\verse Behold, in this thou art not just: I will answer thee, that God is greater than man.

\verse Why dost thou strive against him? for he giveth not account of any of his matters.

\verse For God speaketh once, yea twice, yet man perceiveth it not.

\verse In a dream, in a vision of the night, when deep sleep falleth upon men, in slumberings upon the bed; \verse Then he openeth the ears of men, and sealeth their instruction, \verse That he may withdraw man from his purpose, and hide pride from man.

\verse He keepeth back his soul from the pit, and his life from perishing by the sword.

\verse He is chastened also with pain upon his bed, and the multitude of his bones with strong pain: \verse So that his life abhorreth bread, and his soul dainty meat.

\verse His flesh is consumed away, that it cannot be seen; and his bones that were not seen stick out.

\verse Yea, his soul draweth near unto the grave, and his life to the destroyers.

\verse If there be a messenger with him, an interpreter, one among a thousand, to shew unto man his uprightness: \verse Then he is gracious unto him, and saith, Deliver him from going down to the pit: I have found a ransom.

\verse His flesh shall be fresher than a child's: he shall return to the days of his youth: \verse He shall pray unto God, and he will be favourable unto him: and he shall see his face with joy: for he will render unto man his righteousness.

\verse He looketh upon men, and if any say, I have sinned, and perverted that which was right, and it profited me not; \verse He will deliver his soul from going into the pit, and his life shall see the light.

\verse Lo, all these things worketh God oftentimes with man, \verse To bring back his soul from the pit, to be enlightened with the light of the living.

\verse Mark well, O Job, hearken unto me: hold thy peace, and I will speak.

\verse If thou hast anything to say, answer me: speak, for I desire to justify thee.

\verse If not, hearken unto me: hold thy peace, and I shall teach thee wisdom.


\chapter
\verse Furthermore Elihu answered and said, \verse Hear my words, O ye wise men; and give ear unto me, ye that have knowledge.

\verse For the ear trieth words, as the mouth tasteth meat.

\verse Let us choose to us judgment: let us know among ourselves what is good.

\verse For Job hath said, I am righteous: and God hath taken away my judgment.

\verse Should I lie against my right? my wound is incurable without transgression.

\verse What man is like Job, who drinketh up scorning like water?  \verse Which goeth in company with the workers of iniquity, and walketh with wicked men.

\verse For he hath said, It profiteth a man nothing that he should delight himself with God.

\verse Therefore hearken unto me ye men of understanding: far be it from God, that he should do wickedness; and from the Almighty, that he should commit iniquity.

\verse For the work of a man shall he render unto him, and cause every man to find according to his ways.

\verse Yea, surely God will not do wickedly, neither will the Almighty pervert judgment.

\verse Who hath given him a charge over the earth? or who hath disposed the whole world?  \verse If he set his heart upon man, if he gather unto himself his spirit and his breath; \verse All flesh shall perish together, and man shall turn again unto dust.

\verse If now thou hast understanding, hear this: hearken to the voice of my words.

\verse Shall even he that hateth right govern? and wilt thou condemn him that is most just?  \verse Is it fit to say to a king, Thou art wicked? and to princes, Ye are ungodly?  \verse How much less to him that accepteth not the persons of princes, nor regardeth the rich more than the poor? for they all are the work of his hands.

\verse In a moment shall they die, and the people shall be troubled at midnight, and pass away: and the mighty shall be taken away without hand.

\verse For his eyes are upon the ways of man, and he seeth all his goings.

\verse There is no darkness, nor shadow of death, where the workers of iniquity may hide themselves.

\verse For he will not lay upon man more than right; that he should enter into judgment with God.

\verse He shall break in pieces mighty men without number, and set others in their stead.

\verse Therefore he knoweth their works, and he overturneth them in the night, so that they are destroyed.

\verse He striketh them as wicked men in the open sight of others; \verse Because they turned back from him, and would not consider any of his ways: \verse So that they cause the cry of the poor to come unto him, and he heareth the cry of the afflicted.

\verse When he giveth quietness, who then can make trouble? and when he hideth his face, who then can behold him? whether it be done against a nation, or against a man only: \verse That the hypocrite reign not, lest the people be ensnared.

\verse Surely it is meet to be said unto God, I have borne chastisement, I will not offend any more: \verse That which I see not teach thou me: if I have done iniquity, I will do no more.

\verse Should it be according to thy mind? he will recompense it, whether thou refuse, or whether thou choose; and not I: therefore speak what thou knowest.

\verse Let men of understanding tell me, and let a wise man hearken unto me.

\verse Job hath spoken without knowledge, and his words were without wisdom.

\verse My desire is that Job may be tried unto the end because of his answers for wicked men.

\verse For he addeth rebellion unto his sin, he clappeth his hands among us, and multiplieth his words against God.


\chapter
\verse Elihu spake moreover, and said, \verse Thinkest thou this to be right, that thou saidst, My righteousness is more than God's?  \verse For thou saidst, What advantage will it be unto thee? and, What profit shall I have, if I be cleansed from my sin?  \verse I will answer thee, and thy companions with thee.

\verse Look unto the heavens, and see; and behold the clouds which are higher than thou.

\verse If thou sinnest, what doest thou against him? or if thy transgressions be multiplied, what doest thou unto him?  \verse If thou be righteous, what givest thou him? or what receiveth he of thine hand?  \verse Thy wickedness may hurt a man as thou art; and thy righteousness may profit the son of man.

\verse By reason of the multitude of oppressions they make the oppressed to cry: they cry out by reason of the arm of the mighty.

\verse But none saith, Where is God my maker, who giveth songs in the night; \verse Who teacheth us more than the beasts of the earth, and maketh us wiser than the fowls of heaven?  \verse There they cry, but none giveth answer, because of the pride of evil men.

\verse Surely God will not hear vanity, neither will the Almighty regard it.

\verse Although thou sayest thou shalt not see him, yet judgment is before him; therefore trust thou in him.

\verse But now, because it is not so, he hath visited in his anger; yet he knoweth it not in great extremity: \verse Therefore doth Job open his mouth in vain; he multiplieth words without knowledge.


\chapter
\verse Elihu also proceeded, and said, \verse Suffer me a little, and I will shew thee that I have yet to speak on God's behalf.

\verse I will fetch my knowledge from afar, and will ascribe righteousness to my Maker.

\verse For truly my words shall not be false: he that is perfect in knowledge is with thee.

\verse Behold, God is mighty, and despiseth not any: he is mighty in strength and wisdom.

\verse He preserveth not the life of the wicked: but giveth right to the poor.

\verse He withdraweth not his eyes from the righteous: but with kings are they on the throne; yea, he doth establish them for ever, and they are exalted.

\verse And if they be bound in fetters, and be holden in cords of affliction; \verse Then he sheweth them their work, and their transgressions that they have exceeded.

\verse He openeth also their ear to discipline, and commandeth that they return from iniquity.

\verse If they obey and serve him, they shall spend their days in prosperity, and their years in pleasures.

\verse But if they obey not, they shall perish by the sword, and they shall die without knowledge.

\verse But the hypocrites in heart heap up wrath: they cry not when he bindeth them.

\verse They die in youth, and their life is among the unclean.

\verse He delivereth the poor in his affliction, and openeth their ears in oppression.

\verse Even so would he have removed thee out of the strait into a broad place, where there is no straitness; and that which should be set on thy table should be full of fatness.

\verse But thou hast fulfilled the judgment of the wicked: judgment and justice take hold on thee.

\verse Because there is wrath, beware lest he take thee away with his stroke: then a great ransom cannot deliver thee.

\verse Will he esteem thy riches? no, not gold, nor all the forces of strength.

\verse Desire not the night, when people are cut off in their place.

\verse Take heed, regard not iniquity: for this hast thou chosen rather than affliction.

\verse Behold, God exalteth by his power: who teacheth like him?  \verse Who hath enjoined him his way? or who can say, Thou hast wrought iniquity?  \verse Remember that thou magnify his work, which men behold.

\verse Every man may see it; man may behold it afar off.

\verse Behold, God is great, and we know him not, neither can the number of his years be searched out.

\verse For he maketh small the drops of water: they pour down rain according to the vapour thereof: \verse Which the clouds do drop and distil upon man abundantly.

\verse Also can any understand the spreadings of the clouds, or the noise of his tabernacle?  \verse Behold, he spreadeth his light upon it, and covereth the bottom of the sea.

\verse For by them judgeth he the people; he giveth meat in abundance.

\verse With clouds he covereth the light; and commandeth it not to shine by the cloud that cometh betwixt.

\verse The noise thereof sheweth concerning it, the cattle also concerning the vapour.


\chapter
\verse At this also my heart trembleth, and is moved out of his place.

\verse Hear attentively the noise of his voice, and the sound that goeth out of his mouth.

\verse He directeth it under the whole heaven, and his lightning unto the ends of the earth.

\verse After it a voice roareth: he thundereth with the voice of his excellency; and he will not stay them when his voice is heard.

\verse God thundereth marvellously with his voice; great things doeth he, which we cannot comprehend.

\verse For he saith to the snow, Be thou on the earth; likewise to the small rain, and to the great rain of his strength.

\verse He sealeth up the hand of every man; that all men may know his work.

\verse Then the beasts go into dens, and remain in their places.

\verse Out of the south cometh the whirlwind: and cold out of the north.

\verse By the breath of God frost is given: and the breadth of the waters is straitened.

\verse Also by watering he wearieth the thick cloud: he scattereth his bright cloud: \verse And it is turned round about by his counsels: that they may do whatsoever he commandeth them upon the face of the world in the earth.

\verse He causeth it to come, whether for correction, or for his land, or for mercy.

\verse Hearken unto this, O Job: stand still, and consider the wondrous works of God.

\verse Dost thou know when God disposed them, and caused the light of his cloud to shine?  \verse Dost thou know the balancings of the clouds, the wondrous works of him which is perfect in knowledge?  \verse How thy garments are warm, when he quieteth the earth by the south wind?  \verse Hast thou with him spread out the sky, which is strong, and as a molten looking glass?  \verse Teach us what we shall say unto him; for we cannot order our speech by reason of darkness.

\verse Shall it be told him that I speak? if a man speak, surely he shall be swallowed up.

\verse And now men see not the bright light which is in the clouds: but the wind passeth, and cleanseth them.

\verse Fair weather cometh out of the north: with God is terrible majesty.

\verse Touching the Almighty, we cannot find him out: he is excellent in power, and in judgment, and in plenty of justice: he will not afflict.

\verse Men do therefore fear him: he respecteth not any that are wise of heart.


\chapter
\verse Then the \LORD answered Job out of the whirlwind, and said, \verse Who is this that darkeneth counsel by words without knowledge?  \verse Gird up now thy loins like a man; for I will demand of thee, and answer thou me.

\verse Where wast thou when I laid the foundations of the earth?  declare, if thou hast understanding.

\verse Who hath laid the measures thereof, if thou knowest? or who hath stretched the line upon it?  \verse Whereupon are the foundations thereof fastened? or who laid the corner stone thereof; \verse When the morning stars sang together, and all the sons of God shouted for joy?  \verse Or who shut up the sea with doors, when it brake forth, as if it had issued out of the womb?  \verse When I made the cloud the garment thereof, and thick darkness a swaddlingband for it, \verse And brake up for it my decreed place, and set bars and doors, \verse And said, Hitherto shalt thou come, but no further: and here shall thy proud waves be stayed?  \verse Hast thou commanded the morning since thy days; and caused the dayspring to know his place; \verse That it might take hold of the ends of the earth, that the wicked might be shaken out of it?  \verse It is turned as clay to the seal; and they stand as a garment.

\verse And from the wicked their light is withholden, and the high arm shall be broken.

\verse Hast thou entered into the springs of the sea? or hast thou walked in the search of the depth?  \verse Have the gates of death been opened unto thee? or hast thou seen the doors of the shadow of death?  \verse Hast thou perceived the breadth of the earth? declare if thou knowest it all.

\verse Where is the way where light dwelleth? and as for darkness, where is the place thereof, \verse That thou shouldest take it to the bound thereof, and that thou shouldest know the paths to the house thereof?  \verse Knowest thou it, because thou wast then born? or because the number of thy days is great?  \verse Hast thou entered into the treasures of the snow? or hast thou seen the treasures of the hail, \verse Which I have reserved against the time of trouble, against the day of battle and war?  \verse By what way is the light parted, which scattereth the east wind upon the earth?  \verse Who hath divided a watercourse for the overflowing of waters, or a way for the lightning of thunder; \verse To cause it to rain on the earth, where no man is; on the wilderness, wherein there is no man; \verse To satisfy the desolate and waste ground; and to cause the bud of the tender herb to spring forth?  \verse Hath the rain a father? or who hath begotten the drops of dew?  \verse Out of whose womb came the ice? and the hoary frost of heaven, who hath gendered it?  \verse The waters are hid as with a stone, and the face of the deep is frozen.

\verse Canst thou bind the sweet influences of Pleiades, or loose the bands of Orion?  \verse Canst thou bring forth Mazzaroth in his season?  or canst thou guide Arcturus with his sons?  \verse Knowest thou the ordinances of heaven? canst thou set the dominion thereof in the earth?  \verse Canst thou lift up thy voice to the clouds, that abundance of waters may cover thee?  \verse Canst thou send lightnings, that they may go and say unto thee, Here we are?  \verse Who hath put wisdom in the inward parts? or who hath given understanding to the heart?  \verse Who can number the clouds in wisdom? or who can stay the bottles of heaven, \verse When the dust groweth into hardness, and the clods cleave fast together?  \verse Wilt thou hunt the prey for the lion? or fill the appetite of the young lions, \verse When they couch in their dens, and abide in the covert to lie in wait?  \verse Who provideth for the raven his food? when his young ones cry unto God, they wander for lack of meat.


\chapter
\verse Knowest thou the time when the wild goats of the rock bring forth? or canst thou mark when the hinds do calve?  \verse Canst thou number the months that they fulfil? or knowest thou the time when they bring forth?  \verse They bow themselves, they bring forth their young ones, they cast out their sorrows.

\verse Their young ones are in good liking, they grow up with corn; they go forth, and return not unto them.

\verse Who hath sent out the wild ass free? or who hath loosed the bands of the wild ass?  \verse Whose house I have made the wilderness, and the barren land his dwellings.

\verse He scorneth the multitude of the city, neither regardeth he the crying of the driver.

\verse The range of the mountains is his pasture, and he searcheth after every green thing.

\verse Will the unicorn be willing to serve thee, or abide by thy crib?  \verse Canst thou bind the unicorn with his band in the furrow? or will he harrow the valleys after thee?  \verse Wilt thou trust him, because his strength is great? or wilt thou leave thy labour to him?  \verse Wilt thou believe him, that he will bring home thy seed, and gather it into thy barn?  \verse Gavest thou the goodly wings unto the peacocks?  or wings and feathers unto the ostrich?  \verse Which leaveth her eggs in the earth, and warmeth them in dust, \verse And forgetteth that the foot may crush them, or that the wild beast may break them.

\verse She is hardened against her young ones, as though they were not her's: her labour is in vain without fear; \verse Because God hath deprived her of wisdom, neither hath he imparted to her understanding.

\verse What time she lifteth up herself on high, she scorneth the horse and his rider.

\verse Hast thou given the horse strength? hast thou clothed his neck with thunder?  \verse Canst thou make him afraid as a grasshopper? the glory of his nostrils is terrible.

\verse He paweth in the valley, and rejoiceth in his strength: he goeth on to meet the armed men.

\verse He mocketh at fear, and is not affrighted; neither turneth he back from the sword.

\verse The quiver rattleth against him, the glittering spear and the shield.

\verse He swalloweth the ground with fierceness and rage: neither believeth he that it is the sound of the trumpet.

\verse He saith among the trumpets, Ha, ha; and he smelleth the battle afar off, the thunder of the captains, and the shouting.

\verse Doth the hawk fly by thy wisdom, and stretch her wings toward the south?  \verse Doth the eagle mount up at thy command, and make her nest on high?  \verse She dwelleth and abideth on the rock, upon the crag of the rock, and the strong place.

\verse From thence she seeketh the prey, and her eyes behold afar off.

\verse Her young ones also suck up blood: and where the slain are, there is she.


\chapter
\verse Moreover the \LORD answered Job, and said, \verse Shall he that contendeth with the Almighty instruct him? he that reproveth God, let him answer it.

\verse Then Job answered the \LORD, and said, \verse Behold, I am vile; what shall I answer thee? I will lay mine hand upon my mouth.

\verse Once have I spoken; but I will not answer: yea, twice; but I will proceed no further.

\verse Then answered the \LORD unto Job out of the whirlwind, and said, \verse Gird up thy loins now like a man: I will demand of thee, and declare thou unto me.

\verse Wilt thou also disannul my judgment? wilt thou condemn me, that thou mayest be righteous?  \verse Hast thou an arm like God? or canst thou thunder with a voice like him?  \verse Deck thyself now with majesty and excellency; and array thyself with glory and beauty.

\verse Cast abroad the rage of thy wrath: and behold every one that is proud, and abase him.

\verse Look on every one that is proud, and bring him low; and tread down the wicked in their place.

\verse Hide them in the dust together; and bind their faces in secret.

\verse Then will I also confess unto thee that thine own right hand can save thee.

\verse Behold now behemoth, which I made with thee; he eateth grass as an ox.

\verse Lo now, his strength is in his loins, and his force is in the navel of his belly.

\verse He moveth his tail like a cedar: the sinews of his stones are wrapped together.

\verse His bones are as strong pieces of brass; his bones are like bars of iron.

\verse He is the chief of the ways of God: he that made him can make his sword to approach unto him.

\verse Surely the mountains bring him forth food, where all the beasts of the field play.

\verse He lieth under the shady trees, in the covert of the reed, and fens.

\verse The shady trees cover him with their shadow; the willows of the brook compass him about.

\verse Behold, he drinketh up a river, and hasteth not: he trusteth that he can draw up Jordan into his mouth.

\verse He taketh it with his eyes: his nose pierceth through snares.


\chapter
\verse Canst thou draw out leviathan with an hook? or his tongue with a cord which thou lettest down?  \verse Canst thou put an hook into his nose? or bore his jaw through with a thorn?  \verse Will he make many supplications unto thee? will he speak soft words unto thee?  \verse Will he make a covenant with thee? wilt thou take him for a servant for ever?  \verse Wilt thou play with him as with a bird? or wilt thou bind him for thy maidens?  \verse Shall the companions make a banquet of him? shall they part him among the merchants?  \verse Canst thou fill his skin with barbed irons? or his head with fish spears?  \verse Lay thine hand upon him, remember the battle, do no more.

\verse Behold, the hope of him is in vain: shall not one be cast down even at the sight of him?  \verse None is so fierce that dare stir him up: who then is able to stand before me?  \verse Who hath prevented me, that I should repay him? whatsoever is under the whole heaven is mine.

\verse I will not conceal his parts, nor his power, nor his comely proportion.

\verse Who can discover the face of his garment? or who can come to him with his double bridle?  \verse Who can open the doors of his face? his teeth are terrible round about.

\verse His scales are his pride, shut up together as with a close seal.

\verse One is so near to another, that no air can come between them.

\verse They are joined one to another, they stick together, that they cannot be sundered.

\verse By his neesings a light doth shine, and his eyes are like the eyelids of the morning.

\verse Out of his mouth go burning lamps, and sparks of fire leap out.

\verse Out of his nostrils goeth smoke, as out of a seething pot or caldron.

\verse His breath kindleth coals, and a flame goeth out of his mouth.

\verse In his neck remaineth strength, and sorrow is turned into joy before him.

\verse The flakes of his flesh are joined together: they are firm in themselves; they cannot be moved.

\verse His heart is as firm as a stone; yea, as hard as a piece of the nether millstone.

\verse When he raiseth up himself, the mighty are afraid: by reason of breakings they purify themselves.

\verse The sword of him that layeth at him cannot hold: the spear, the dart, nor the habergeon.

\verse He esteemeth iron as straw, and brass as rotten wood.

\verse The arrow cannot make him flee: slingstones are turned with him into stubble.

\verse Darts are counted as stubble: he laugheth at the shaking of a spear.

\verse Sharp stones are under him: he spreadeth sharp pointed things upon the mire.

\verse He maketh the deep to boil like a pot: he maketh the sea like a pot of ointment.

\verse He maketh a path to shine after him; one would think the deep to be hoary.

\verse Upon earth there is not his like, who is made without fear.

\verse He beholdeth all high things: he is a king over all the children of pride.


\chapter
\verse Then Job answered the \LORD, and said, \verse I know that thou canst do every thing, and that no thought can be withholden from thee.

\verse Who is he that hideth counsel without knowledge? therefore have I uttered that I understood not; things too wonderful for me, which I knew not.

\verse Hear, I beseech thee, and I will speak: I will demand of thee, and declare thou unto me.

\verse I have heard of thee by the hearing of the ear: but now mine eye seeth thee.

\verse Wherefore I abhor myself, and repent in dust and ashes.

\verse And it was so, that after the \LORD had spoken these words unto Job, the \LORD said to Eliphaz the Temanite, My wrath is kindled against thee, and against thy two friends: for ye have not spoken of me the thing that is right, as my servant Job hath.

\verse Therefore take unto you now seven bullocks and seven rams, and go to my servant Job, and offer up for yourselves a burnt offering; and my servant Job shall pray for you: for him will I accept: lest I deal with you after your folly, in that ye have not spoken of me the thing which is right, like my servant Job.

\verse So Eliphaz the Temanite and Bildad the Shuhite and Zophar the Naamathite went, and did according as the \LORD commanded them: the \LORD also accepted Job.

\verse And the \LORD turned the captivity of Job, when he prayed for his friends: also the \LORD gave Job twice as much as he had before.

\verse Then came there unto him all his brethren, and all his sisters, and all they that had been of his acquaintance before, and did eat bread with him in his house: and they bemoaned him, and comforted him over all the evil that the \LORD had brought upon him: every man also gave him a piece of money, and every one an earring of gold.

\verse So the \LORD blessed the latter end of Job more than his beginning: for he had fourteen thousand sheep, and six thousand camels, and a thousand yoke of oxen, and a thousand she asses.

\verse He had also seven sons and three daughters.

\verse And he called the name of the first, Jemima; and the name of the second, Kezia; and the name of the third, Kerenhappuch.

\verse And in all the land were no women found so fair as the daughters of Job: and their father gave them inheritance among their brethren.

\verse After this lived Job an hundred and forty years, and saw his sons, and his sons' sons, even four generations.

\verse So Job died, being old and full of days.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "bible.tex"
%%% End: 
