\book{Obadiah}


1:1 The vision of Obadiah. Thus saith the Lord GOD concerning Edom;
We have heard a rumour from the \LORD, and an ambassador is sent among
the heathen, Arise ye, and let us rise up against her in battle.

1:2 Behold, I have made thee small among the heathen: thou art greatly
despised.

1:3 The pride of thine heart hath deceived thee, thou that dwellest in
the clefts of the rock, whose habitation is high; that saith in his
heart, Who shall bring me down to the ground?  1:4 Though thou exalt
thyself as the eagle, and though thou set thy nest among the stars,
thence will I bring thee down, saith the \LORD.

1:5 If thieves came to thee, if robbers by night, (how art thou cut
off!)  would they not have stolen till they had enough? if the
grapegatherers came to thee, would they not leave some grapes?  1:6
How are the things of Esau searched out! how are his hidden things
sought up!  1:7 All the men of thy confederacy have brought thee even
to the border: the men that were at peace with thee have deceived
thee, and prevailed against thee; that they eat thy bread have laid a
wound under thee: there is none understanding in him.

1:8 Shall I not in that day, saith the \LORD, even destroy the wise men
out of Edom, and understanding out of the mount of Esau?  1:9 And thy
mighty men, O Teman, shall be dismayed, to the end that every one of
the mount of Esau may be cut off by slaughter.

1:10 For thy violence against thy brother Jacob shame shall cover
thee, and thou shalt be cut off for ever.

1:11 In the day that thou stoodest on the other side, in the day that
the strangers carried away captive his forces, and foreigners entered
into his gates, and cast lots upon Jerusalem, even thou wast as one of
them.

1:12 But thou shouldest not have looked on the day of thy brother in
the day that he became a stranger; neither shouldest thou have
rejoiced over the children of Judah in the day of their destruction;
neither shouldest thou have spoken proudly in the day of distress.

1:13 Thou shouldest not have entered into the gate of my people in the
day of their calamity; yea, thou shouldest not have looked on their
affliction in the day of their calamity, nor have laid hands on their
substance in the day of their calamity; 1:14 Neither shouldest thou
have stood in the crossway, to cut off those of his that did escape;
neither shouldest thou have delivered up those of his that did remain
in the day of distress.

1:15 For the day of the \LORD is near upon all the heathen: as thou
hast done, it shall be done unto thee: thy reward shall return upon
thine own head.

1:16 For as ye have drunk upon my holy mountain, so shall all the
heathen drink continually, yea, they shall drink, and they shall
swallow down, and they shall be as though they had not been.

1:17 But upon mount Zion shall be deliverance, and there shall be
holiness; and the house of Jacob shall possess their possessions.

1:18 And the house of Jacob shall be a fire, and the house of Joseph a
flame, and the house of Esau for stubble, and they shall kindle in
them, and devour them; and there shall not be any remaining of the
house of Esau; for the \LORD hath spoken it.

1:19 And they of the south shall possess the mount of Esau; and they
of the plain the Philistines: and they shall possess the fields of
Ephraim, and the fields of Samaria: and Benjamin shall possess Gilead.

1:20 And the captivity of this host of the children of Israel shall
possess that of the Canaanites, even unto Zarephath; and the captivity
of Jerusalem, which is in Sepharad, shall possess the cities of the
south.

1:21 And saviours shall come up on mount Zion to judge the mount of
Esau; and the kingdom shall be the \LORD's.




Jonah


1:1 Now the word of the \LORD came unto Jonah the son of Amittai,
saying, 1:2 Arise, go to Nineveh, that great city, and cry against it;
for their wickedness is come up before me.

1:3 But Jonah rose up to flee unto Tarshish from the presence of the
\LORD, and went down to Joppa; and he found a ship going to Tarshish:
so he paid the fare thereof, and went down into it, to go with them
unto Tarshish from the presence of the \LORD.

1:4 But the \LORD sent out a great wind into the sea, and there was a
mighty tempest in the sea, so that the ship was like to be broken.

1:5 Then the mariners were afraid, and cried every man unto his god,
and cast forth the wares that were in the ship into the sea, to
lighten it of them. But Jonah was gone down into the sides of the
ship; and he lay, and was fast asleep.

1:6 So the shipmaster came to him, and said unto him, What meanest
thou, O sleeper? arise, call upon thy God, if so be that God will
think upon us, that we perish not.

1:7 And they said every one to his fellow, Come, and let us cast lots,
that we may know for whose cause this evil is upon us. So they cast
lots, and the lot fell upon Jonah.

1:8 Then said they unto him, Tell us, we pray thee, for whose cause
this evil is upon us; What is thine occupation? and whence comest
thou? what is thy country? and of what people art thou?  1:9 And he
said unto them, I am an Hebrew; and I fear the \LORD, the God of
heaven, which hath made the sea and the dry land.

1:10 Then were the men exceedingly afraid, and said unto him. Why hast
thou done this? For the men knew that he fled from the presence of the
\LORD, because he had told them.

1:11 Then said they unto him, What shall we do unto thee, that the sea
may be calm unto us? for the sea wrought, and was tempestuous.

1:12 And he said unto them, Take me up, and cast me forth into the
sea; so shall the sea be calm unto you: for I know that for my sake
this great tempest is upon you.

1:13 Nevertheless the men rowed hard to bring it to the land; but they
could not: for the sea wrought, and was tempestuous against them.

1:14 Wherefore they cried unto the \LORD, and said, We beseech thee, O
\LORD, we beseech thee, let us not perish for this man's life, and lay
not upon us innocent blood: for thou, O \LORD, hast done as it pleased
thee.

1:15 So they look up Jonah, and cast him forth into the sea: and the
sea ceased from her raging.

1:16 Then the men feared the \LORD exceedingly, and offered a sacrifice
unto the \LORD, and made vows.

1:17 Now the \LORD had prepared a great fish to swallow up Jonah. And
Jonah was in the belly of the fish three days and three nights.

2:1 Then Jonah prayed unto the \LORD his God out of the fish's belly,
2:2 And said, I cried by reason of mine affliction unto the \LORD, and
he heard me; out of the belly of hell cried I, and thou heardest my
voice.

2:3 For thou hadst cast me into the deep, in the midst of the seas;
and the floods compassed me about: all thy billows and thy waves
passed over me.

2:4 Then I said, I am cast out of thy sight; yet I will look again
toward thy holy temple.

2:5 The waters compassed me about, even to the soul: the depth closed
me round about, the weeds were wrapped about my head.

2:6 I went down to the bottoms of the mountains; the earth with her
bars was about me for ever: yet hast thou brought up my life from
corruption, O \LORD my God.

2:7 When my soul fainted within me I remembered the \LORD: and my
prayer came in unto thee, into thine holy temple.

2:8 They that observe lying vanities forsake their own mercy.

2:9 But I will sacrifice unto thee with the voice of thanksgiving; I
will pay that that I have vowed. Salvation is of the \LORD.

2:10 And the \LORD spake unto the fish, and it vomited out Jonah upon
the dry land.

3:1 And the word of the \LORD came unto Jonah the second time, saying,
3:2 Arise, go unto Nineveh, that great city, and preach unto it the
preaching that I bid thee.

3:3 So Jonah arose, and went unto Nineveh, according to the word of
the \LORD. Now Nineveh was an exceeding great city of three days'
journey.

3:4 And Jonah began to enter into the city a day's journey, and he
cried, and said, Yet forty days, and Nineveh shall be overthrown.

3:5 So the people of Nineveh believed God, and proclaimed a fast, and
put on sackcloth, from the greatest of them even to the least of them.

3:6 For word came unto the king of Nineveh, and he arose from his
throne, and he laid his robe from him, and covered him with sackcloth,
and sat in ashes.

3:7 And he caused it to be proclaimed and published through Nineveh by
the decree of the king and his nobles, saying, Let neither man nor
beast, herd nor flock, taste any thing: let them not feed, nor drink
water: 3:8 But let man and beast be covered with sackcloth, and cry
mightily unto God: yea, let them turn every one from his evil way, and
from the violence that is in their hands.

3:9 Who can tell if God will turn and repent, and turn away from his
fierce anger, that we perish not?  3:10 And God saw their works, that
they turned from their evil way; and God repented of the evil, that he
had said that he would do unto them; and he did it not.

4:1 But it displeased Jonah exceedingly, and he was very angry.

4:2 And he prayed unto the \LORD, and said, I pray thee, O \LORD, was
not this my saying, when I was yet in my country? Therefore I fled
before unto Tarshish: for I knew that thou art a gracious God, and
merciful, slow to anger, and of great kindness, and repentest thee of
the evil.

4:3 Therefore now, O \LORD, take, I beseech thee, my life from me; for
it is better for me to die than to live.

4:4 Then said the \LORD, Doest thou well to be angry?  4:5 So Jonah
went out of the city, and sat on the east side of the city, and there
made him a booth, and sat under it in the shadow, till he might see
what would become of the city.

4:6 And the \LORD God prepared a gourd, and made it to come up over
Jonah, that it might be a shadow over his head, to deliver him from
his grief. So Jonah was exceeding glad of the gourd.

4:7 But God prepared a worm when the morning rose the next day, and it
smote the gourd that it withered.

4:8 And it came to pass, when the sun did arise, that God prepared a
vehement east wind; and the sun beat upon the head of Jonah, that he
fainted, and wished in himself to die, and said, It is better for me
to die than to live.

4:9 And God said to Jonah, Doest thou well to be angry for the gourd?
And he said, I do well to be angry, even unto death.

4:10 Then said the \LORD, Thou hast had pity on the gourd, for the
which thou hast not laboured, neither madest it grow; which came up in
a night, and perished in a night: 4:11 And should not I spare Nineveh,
that great city, wherein are more then sixscore thousand persons that
cannot discern between their right hand and their left hand; and also
much cattle?

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "bible.tex"
%%% End: 
